### 2016-08-25 docs/article.Rmd
articolo

---
title: "R package qtdshift: A functional programming approach to shifted probability ditributions"
author: "Veronica"
date: "August 25, 2016"
output: html_document
---

## Introduction

This paper aims to be an implementation of the methodology used by Andrea Spano article, "R package qdist: A functional programming approach to truncated probability ditributions", which provides an alternative method for computing from truncated probability distribution using a functional programming approach. I was inspired by his work to implement the same methodology to shifted probability distributions.

## Shifted probability function in R

Why shifted distributions? Shift is distribution trasformation used daily in many statistical fields. R does not provide a specific package for shifted distribution models, but it provides package `stats` which contains a wide set of functions for standard distribution models. I refer to: probability, density, quantile and random number generator functions, that share a stable naming convention both for the names of the functions and in the first argument of these functions. In particular:

* probability distribution functions: `p<dist>(q)` with `q` vector of quantiles;
* density functions: `d<dist>(x)` with `x` vector of quantiles;
* quantile functions: `q<dist>(p)` with `p` vector of probabilities;
* random number generation: `r<dist>(n)` with `n` number of observations.

where `<dist>` indicates the conventional name of distribution family.

Let us see an example.  
Considering the Normal distribution (`norm`), we have: `pnorm(q)`, `dnorm(x)`, `qnorm(p)` and `rnorm(n)`.

So, we use to write:

```{r dnorm}
dnorm(x = 6:10, mean = 8, sd = 1)
```

to get density values at 6:10 from a normal distribution with parameters `mean = 8` and `sd = 1`.

Going back to shifted distribution, we could easily write a shifted version for `dnorm()`, as:

```{r}
sdnorm <- function(x, mean = 0, sd = 1, shift = 0, ...){
            x_shifted <- x - shift
            density <- dnorm(x = x_shifted, mean = mean, sd = sd, ...)
            return(density)
}
```

So, we write:

```{r}
sdnorm(x = 8:12, mean = 10, sd = 1, shift = 1)
```

to get density values at 6:10 from a normal distribution with parameters `mean = 8` and `sd = 1`, shifted of 1 unit.

At first view, this approach seems easy and cool but thinking about it I realized that I would need to write a different function for each probability distribution I would like to work with. So I decided to use functional programming approach to develop four general functions which compute shifted distribution models for any distribution chosen:

* `pshift()` for shifted probability distributions;
* `dshift()` for shifted density functions;
* `qshift()` for shifted quantile functions;
* `rshift()` for shifted random numbers generator functions.

These functions are included in `qtdshift` package available at [https://github.com/veronicagiro/qshift](https://github.com/veronicagiro/qshift).

They share the same logic: they take as input a probability distribution as a character string and return the equivalent shifted distribution as a function object.

```{r, eval = FALSE}
devtools::install_github("veronicagiro/qshift", build=FALSE)
require(qtdshift)
spnorm <- pshift("norm")
sdnorm <- dshift("norm")
sqnorm <- qshift("norm")
srnorm <- rshift("norm")
```

<!--
esempi
-->

### 2016-08-26 docs/article.Rmd
articolo ok

@@
 ---
-title: "R package qtdshift: A functional programming approach to shifted probability ditributions"
+title: 'R package qtdshift: A functional programming approach to shifted probability
+  ditributions'
 author: "Veronica"
@@
 
-This paper aims to be an implementation of the methodology used by Andrea Spano article, "R package qdist: A functional programming approach to truncated probability ditributions", which provides an alternative method for computing from truncated probability distribution using a functional programming approach.
+This paper aims to be an implementation of the methodology used by Andrea Spanò article, "R package qdist: A functional programming approach to truncated probability ditributions", which provides an alternative method for computing from truncated probability distribution using a functional programming approach.
 I was inspired by his work to implement the same methodology to shifted probability distributions.
 
-## Shifted probability function in R
+## Shifted probability functions in R
 
 Why shifted distributions?
-Shift is distribution trasformation used daily in many statistical fields.
+Shift is a distribution trasformation used daily in many statistical fields.
 R does not provide a specific package for shifted distribution models, but it provides package `stats` which contains a wide set of functions for standard distribution models.
@@
 
-to get density values at 6:10 from a normal distribution with parameters `mean = 8` and `sd = 1`.
+to get density values at 6:10 from a Normal distribution with parameters `mean = 8` and `sd = 1`.
 
@@
 
-```{r}
-sdnorm <- function(x, mean = 0, sd = 1, shift = 0, ...){
+```{r sdnorm_definition}
+shift_dnorm <- function(x, mean = 0, sd = 1, shift = 0, ...){
             x_shifted <- x - shift
@@
 
-So, we write:
+So we replicate the previous example but shifting the Normal distribution of 1 unit:
 
-```{r}
-sdnorm(x = 8:12, mean = 10, sd = 1, shift = 1)
+```{r sdnorm_example_1}
+shift_dnorm(x = 6:10, mean = 8, sd = 1, shift = 1)
 ```
-
-to get density values at 6:10 from a normal distribution with parameters `mean = 8` and `sd = 1`, shifted of 1 unit.
 
@@
 
+* `dshift()` for shifted density functions;
 * `pshift()` for shifted probability distributions;
-* `dshift()` for shifted density functions;
 * `qshift()` for shifted quantile functions;
@@
 
-They share the same logic: they take as input a probability distribution as a character string and return the equivalent shifted distribution as a function object.
-
-```{r, eval = FALSE}
+```{r github, eval = FALSE}
 devtools::install_github("veronicagiro/qshift", build=FALSE)
-require(qtdshift)
-spnorm <- pshift("norm")
-sdnorm <- dshift("norm")
-sqnorm <- qshift("norm")
-srnorm <- rshift("norm")
 ```
 
-<!--
-esempi
--->
+They share the same logic: they take as input a probability distribution name, as a character string, and return the equivalent shifted distribution as a function object.
+
+```{r qtdshift, message=FALSE}
+require(qtdshift)
+sdnorm <- dshift("norm") # density function
+spnorm <- pshift("norm") # probability function
+sqnorm <- qshift("norm") # quantile function
+srnorm <- rshift("norm") # random generation function
+```
+
+We see in detail the computational process.   
+Let us consider `dshift()`. As we seen, it takes as input the distribution name (`norm`) and proceeds as follows:
+
+* gets the corresponding function `dnorm()`;
+* uses `dnorm()` to create a function, say `density()` that computes shifted probability for normal distributions;
+* modify the formals of `density()` so that it has the same formals as `dnorm()` plus `shift`, corresponding to the shift value;
+* returns `density()`.
+
+So, the function `sdnorm()` has the same formals of the function `dnorm()`, plus the parameter `shift` which indicates the shift value to be applied to distribution; by default it is set as 0.
+
+```{r args}
+args(sdnorm)
+args(dnorm)
+```
+
+Package `qtdshift` is based on two key concepts being part of the functional nature of the R programming language:
+
+* the environment of a function can be used as a placeholder for other objects;
+* function formals can be manipulated.
+
+Theoretical references in chapters [Environments](http://www.quantide.com/website/ramarro-chapter-03/) and [Functions](http://www.quantide.com/website/ramarro-chapter-04/) of Andrea Spano weeb Book [Ramarro](http://www.quantide.com/website/ramarro-r-for-developers/), available on Quantide web site.
+
+### Examples
+
+Let us consider the Normal distribution and the previously defined functions `sdnorm()`, `spnorm()`, `sqnorm()` and `srnorm()`.
+
+* Density function, `sdnorm()`.  
+    This function has a double use:
+
+    * generate probability values from a non-shifted normal distribution by leaving parameter `shift` set to its default (`shift = 0`)
+    * generate probability values from any shifted normal distribution by setting `shift` value
+
+    ```{r sdnorm_example}
+x <- seq(4, 14, len = 100)
+not_shifted <- sdnorm(x = x, mean = 8, sd = 1.5)
+positive_shifted <- sdnorm(x = x, mean = 8, sd = 1.5, shift = 1)
+negative_shifted <- sdnorm(x = x, mean = 8, sd = 1.5, shift = -1)
+    ```
+    We can visualize these results by plotting them:
+
+    ```{r sdnorm_plot, message=FALSE}
+require(ggplot2)
+df <- data.frame(x=x, 
+                 sdnorm_type = factor(c(rep("not_shifted", times=length(x)),
+                    rep("positive_shifted", times=length(x)),
+                    rep("negative_shifted", times=length(x))),
+                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
+                  value = c(not_shifted, positive_shifted, negative_shifted))
+
+pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=sdnorm_type)) +
+  geom_line(size=1) +
+  scale_color_manual(values = c("not_shifted" = "cyan3", 
+                                "positive_shifted" = "#89FF99", 
+                                "negative_shifted" = "#FF89B4")) +
+  xlab("Value") + ylab("Density") +
+  ggtitle("Density plot for shifted and non-shifted normal distribution") +
+  theme(legend.title=element_blank(), legend.text= element_text(size=12),
+  plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+    ```
+
+* Probability function, `spnorm()`:
+
+    ```{r spnorm_example}
+not_shifted <- spnorm(q = x, mean = 8, sd = 1.5)
+positive_shifted <- spnorm(q = x, mean = 8, sd = 1.5, shift = 1)
+negative_shifted <- spnorm(q = x, mean = 8, sd = 1.5, shift = -1)
+    ```
+    We can visualize these results by plotting them:
+
+    ```{r spnorm_plot}
+df <- data.frame(x=x, 
+                 spnorm_type = factor(c(rep("not_shifted", times=length(x)),
+                    rep("positive_shifted", times=length(x)),
+                    rep("negative_shifted", times=length(x))),
+                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
+                  value = c(not_shifted, positive_shifted, negative_shifted))
+
+pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=spnorm_type)) +
+  geom_line(size=1) +
+  scale_color_manual(values = c("not_shifted" = "cyan3", 
+                                "positive_shifted" = "#89FF99", 
+                                "negative_shifted" = "#FF89B4")) +
+  xlab("Quantile") + ylab("Probability") +
+  ggtitle("Probability plot for shifted and non-shifted normal distribution") +
+  theme(legend.title=element_blank(), legend.text= element_text(size=12),
+  plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+    ```
+
+* Quantile function, `sqnorm()`:
+
+    ```{r sqnorm_example}
+    x <- c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1)
+not_shifted <- sqnorm(p = x, mean = 8, sd = 1.5)
+positive_shifted <- sqnorm(p = x, mean = 8, sd = 1.5, shift = 1)
+negative_shifted <- sqnorm(p = x, mean = 8, sd = 1.5, shift = -1)
+    ```
+    We can visualize these results by plotting them:
+
+    ```{r sqnorm_plot}
+df <- data.frame(x=x, 
+                 sqnorm_type = factor(c(rep("not_shifted", times=length(x)),
+                    rep("positive_shifted", times=length(x)),
+                    rep("negative_shifted", times=length(x))),
+                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
+                  value = c(not_shifted, positive_shifted, negative_shifted))
+
+pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=sqnorm_type)) +
+  geom_line(size=1) +
+  scale_color_manual(values = c("not_shifted" = "cyan3", 
+                                "positive_shifted" = "#89FF99", 
+                                "negative_shifted" = "#FF89B4")) +
+  ylab("Quantile") + xlab("Probability") +
+  ggtitle("Quantile plot for shifted and non-shifted normal distribution") +
+  theme(legend.title=element_blank(), legend.text= element_text(size=12),
+  plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+    ```
+
+* Random generator function, `srnorm()`:
+
+    ```{r srnorm_example}
+    x <- 10000
+not_shifted <- srnorm(n = x, mean = 8, sd = 1.5, set_seed = T)
+positive_shifted <- srnorm(n = x, mean = 8, sd = 1.5, shift = 3, set_seed = T)
+negative_shifted <- srnorm(n = x, mean = 8, sd = 1.5, shift = -3, set_seed = T)
+    ```
+    We can visualize these results by plotting them:
+
+    ```{r srnorm_plot}
+df <- data.frame(srnorm_type = factor(c(rep("not_shifted", times=x),
+                    rep("positive_shifted", times=x),
+                    rep("negative_shifted", times=x)),
+                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
+                  value = c(not_shifted, positive_shifted, negative_shifted))
+
+pl <- pl <- ggplot(data = df, mapping = aes(x=value, group=srnorm_type, fill=srnorm_type)) +
+  geom_density(alpha=0.6) +
+    scale_fill_manual(values = c("not_shifted" = "cyan3", 
+                                "positive_shifted" = "#89FF99", 
+                                "negative_shifted" = "#FF89B4")) +
+  ggtitle("Density plot for shifted and non-shifted normal distribution") +
+  theme(legend.title=element_blank(), legend.text= element_text(size=12),
+  plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+    ```
+
+As we know, Normal distribution support goes from $-\infty$ to $+\infty$. What happens considerig a distribution with a limited support?   
+Let us consider the Uniform distribution.
+
+First of all we have to define its probability functions:
+
+```{r qtdshift_unif}
+sdunif <- dshift("unif")
+spunif <- pshift("unif")
+squnif <- qshift("unif")
+srunif <- rshift("unif")
+```
+
+Now the newly created functions are ready for computing from shifted probability models.  
+We limited Uniform support between 6 and 12.
+
+* Density function, `sdunif()`:
+
+    ```{r sdunif_example}
+x <- seq(4, 14, len = 100)
+not_shifted <- sdunif(x = x, min = 6, max = 12)
+positive_shifted <- sdunif(x = x, min = 6, max = 12, shift = 1)
+negative_shifted <- sdunif(x = x, min = 6, max = 12, shift = -1)
+    ```
+    We can visualize these results by plotting them:
+
+    ```{r sdunif_plot}
+df <- data.frame(x=x, 
+                 sdunif_type = factor(c(rep("not_shifted", times=length(x)),
+                    rep("positive_shifted", times=length(x)),
+                    rep("negative_shifted", times=length(x))),
+                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
+                  value = c(not_shifted, positive_shifted, negative_shifted))
+
+pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=sdunif_type)) +
+  geom_line(size=1) +
+  scale_color_manual(values = c("not_shifted" = "cyan3", 
+                                "positive_shifted" = "#89FF99", 
+                                "negative_shifted" = "#FF89B4")) +
+  xlab("Value") + ylab("Density") +
+  ggtitle("Density plot for shifted and non-shifted uniform distribution") +
+  theme(legend.title=element_blank(), legend.text= element_text(size=12),
+  plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+    ```
+
+* Probability function, `spunif()`:
+
+    ```{r spunif_example}
+not_shifted <- spunif(q = x, min = 6, max = 12)
+positive_shifted <- spunif(q = x, min = 6, max = 12, shift = 1)
+negative_shifted <- spunif(q = x, min = 6, max = 12, shift = -1)
+    ```
+    We can visualize these results by plotting them:
+
+    ```{r spunif_plot}
+df <- data.frame(x=x, 
+                 spunif_type = factor(c(rep("not_shifted", times=length(x)),
+                    rep("positive_shifted", times=length(x)),
+                    rep("negative_shifted", times=length(x))),
+                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
+                  value = c(not_shifted, positive_shifted, negative_shifted))
+
+pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=spunif_type)) +
+  geom_line(size=1) +
+  scale_color_manual(values = c("not_shifted" = "cyan3", 
+                                "positive_shifted" = "#89FF99", 
+                                "negative_shifted" = "#FF89B4")) +
+  xlab("Quantile") + ylab("Probability") +
+  ggtitle("Probability plot for shifted and non-shifted uniform distribution") +
+  theme(legend.title=element_blank(), legend.text= element_text(size=12),
+  plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+    ```
+
+* Quantile function, `squnif()`:
+
+    ```{r squnif_example}
+    x <- c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1)
+not_shifted <- squnif(p = x, min = 6, max = 12)
+positive_shifted <- squnif(p = x, min = 6, max = 12, shift = 1)
+negative_shifted <- squnif(p = x, min = 6, max = 12, shift = -1)
+    ```
+    We can visualize these results by plotting them:
+
+    ```{r squnif_plot}
+df <- data.frame(x=x, 
+                 squnif_type = factor(c(rep("not_shifted", times=length(x)),
+                    rep("positive_shifted", times=length(x)),
+                    rep("negative_shifted", times=length(x))),
+                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
+                  value = c(not_shifted, positive_shifted, negative_shifted))
+
+pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=squnif_type)) +
+  geom_line(size=1) +
+  scale_color_manual(values = c("not_shifted" = "cyan3",  
+                                "positive_shifted" = "#89FF99", 
+                                "negative_shifted" = "#FF89B4")) +
+  ylab("Quantile") + xlab("Probability") +
+  ggtitle("Quantile plot for shifted and non-shifted uniform distribution") +
+  theme(legend.title=element_blank(), legend.text= element_text(size=12),
+  plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+    ```
+
+* Random generator function, `srunif()`:
+
+    ```{r srunif_example}
+    x <- 10000
+not_shifted <- srunif(n = x, min = 6, max = 12, set_seed = T)
+positive_shifted <- srunif(n = x, min = 6, max = 12, shift = 3, set_seed = T)
+negative_shifted <- srunif(n = x, min = 6, max = 12, shift = -3, set_seed = T)
+    ```
+    We can visualize these results by plotting them:
+
+    ```{r srunif_plot}
+df <- data.frame(srunif_type = factor(c(rep("not_shifted", times=x),
+                    rep("positive_shifted", times=x),
+                    rep("negative_shifted", times=x)),
+                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
+                  value = c(not_shifted, positive_shifted, negative_shifted))
+
+pl <- pl <- ggplot(data = df, mapping = aes(x=value, group=srunif_type, fill=srunif_type)) +
+  geom_density(alpha=0.6) +
+    scale_color_manual(values = c("not_shifted" = "cyan3", 
+                                "positive_shifted" = "#89FF99", 
+                                "negative_shifted" = "#FF89B4")) +
+  xlab("Value") + ylab("Density") +
+  ggtitle("Density plot for shifted and non-shifted uniform distribution") +
+  theme(legend.title=element_blank(), legend.text= element_text(size=12),
+  plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+
+    ```
+
+As we see, Uniform support changes according to the transformation (the shift) applied to the distribution.
+
+## Extending the computation
+
+In his article, Andrea Spanò use the newly defined probability distribution functions as a base for defining a function for maximum likelihood estimate for the truncated normal distribution.
+I demonstrate that it is possible also for shifted normal distribution.
+Let us see how.
+
+We start from the previously defined fucntion `sdnorm()`, that will be used within an estimator function:
+
+```{r max_likelihood_est}
+ltnorm <- function(x, shift = 0) {
+  theta <- c(mean(x), sd(x))
+  ll <- function(theta, x, shift = 0) {
+    mean <- theta[1]
+    sd <- theta[2]
+    ld <- sdnorm(x = x, mean = mean, sd = sd, shift=shift, log=TRUE)
+    -sum(ld)
+    } 
+  optim(par = theta, fn = ll, x = x, shift=shift)[["par"]]
+}
+```
+
+```{r max_likelihood_est_example}
+# generate a random sample from a shifted normal distribution
+x <- srnorm(n = 1000, mean = 5, sd = 2, shift = 1)
+# maximum likelihood estimate for the shifted normal distribution
+ltnorm(x = x, shift  = 1)
+```
+
+## Bibliography
+
+Pace L., Salvan A. . _Introduzione alla Statistica_. CEDAM. 2010
+
+Spano S. .  _R package qdist: A functional programming approach to truncated probability ditributions_. url
+
+Wickham H. . _Advanced r_. http://adv-r.had.co.nz/. Accessed: 2016-08-24.

### 2016-08-26 R/dshift.R

@@ -65,7 +65,8 @@ dshift <- function(dist){
       dargs <- intersect_args(x = dargs, y = call)
 
       # method for computing density values for shifted distributions
-      dargs$x <- eval(dargs$x) - shift
+      dargs$x <- eval(dargs$x)
+      dargs$x <- dargs$x - shift
       density <- do.call("ddist", as.list(dargs))
 
       # returns density values for shifted distributions

### 2016-08-26 R/pshift.R

@@ -64,8 +64,9 @@ pshift <- function(dist){
     # as a result, the whole string gets all unique arguments belonging to distribution function and pdist
     pargs <- intersect_args(x = pargs, y = call)
 
-    # method for computing distribution values for shifted distributions
-    pargs$q <- eval(pargs$q) - shift
+    # method for computing distribution values for shifted distribution
+    pargs$q <- eval(pargs$q)
+    pargs$q <- pargs$q - shift
     probability <-  do.call("pdist", as.list(pargs))
 
     # returns distribution values for shifted distributions

### 2016-08-26 R/qshift.R

@@ -50,19 +50,10 @@ qshift <- function (dist){
 
     qdist=paste("q", dist, sep = "")
 
-    # uncomment if use solution 1
-    # gets distribution function
-    #pdist=paste("p", dist, sep = "")
-
     # gets quantile function
     # gets argument of quantile function
     qdist <-  get(qdist, mode = "function")
     qargs <-  formals(qdist)
-
-    # uncomment if use solution 1
-    # gets argument of distribution function
-    # pdist <-  get(pdist, mode = "function")
-    # pargs <- formals(pdist)
 
     # Output function starts here
     quantile <- function() {
@@ -73,19 +64,7 @@ qshift <- function (dist){
       # as a result, the whole string gets all unique arguments belonging to quantile function and qdist
       qargs <- intersect_args(x = qargs, y = call)
 
-      # uncomment if use solution 1
-      #pargs <- c(pargs[!is.element(names(pargs), names(call))], call[is.element(names(call), names(pargs))])
-      #pargs <- intersect_args(x = pargs, y = call)
-
-      # method for computing quantile values for shifted distributions
-      # Solution 1
-      #quantile_1 <- do.call("qdist", as.list(qargs))
-      #pargs$q <- quantile_1 + shift
-      #probability <- do.call("pdist", as.list(pargs))
-      #qargs$p <- probability
-      #quantile <- do.call("qdist", as.list(qargs))
-
-      # Solution 2 (correct?)
+      # method for computing quantiles values for shifted distributions
        quantile <- do.call("qdist", as.list(qargs))
        quantile <- quantile + shift
 

### 2016-08-26 R/rshift.R

@@ -61,14 +61,6 @@ rshift <- function(dist){
   # gets argument of random generation function
   rargs <-  formals(rdist)
 
-  # qdist <- paste("q", dist, sep = "")
-  # qdist <-  get(qdist, mode = "function")
-  # qargs <-  formals(qdist)
-  #
-  # pdist <- paste("p", dist, sep = "")
-  # pdist <-  get(pdist, mode = "function")
-  # pargs <-  formals(pdist)
-
   random <- function(...){
 
     # gets random generation arguments
@@ -76,12 +68,6 @@ rshift <- function(dist){
 
     # as a result, the whole string gets all unique arguments belonging to random generation function and rdist
     rargs <- intersect_args(x = rargs, y = call)
-    # pargs <- intersect_args(x = pargs, y = call)
-    # qargs <- intersect_args(x = qargs, y = call)
-
-    # qargs$p <- runif(n)
-    # random_1 <- do.call(qdist, as.list(qargs))
-    # random <- random + shift
 
     # Define starting point for random computation or not
     if(set_seed){

### 2016-08-31 docs/article.Rmd
article

@@
 ---
-title: 'R package qtdshift: A functional programming approach to shifted probability
-  ditributions'
+title: 'R package qtdshift: A functional programming approach to shifted probability distributions'
 author: "Veronica"
 date: "August 25, 2016"
-output: html_document
+output: pdf_document
 ---
@@
 
-This paper aims to be an implementation of the methodology used by Andrea Spanò article, "R package qdist: A functional programming approach to truncated probability ditributions", which provides an alternative method for computing from truncated probability distribution using a functional programming approach.
-I was inspired by his work to implement the same methodology to shifted probability distributions.
+Shifted distributions are used daily in many statistical fields.   
+A shifted distribution is a distribution to which is added a constant, $Y = X + a$, where $X$ and $Y$ are random variables and $a$ is a constant value.
+
+The aim of this paper is to describe an R package, `qdtshift` which generates shifted distribution models for any distribution chosen.
+The packege was realized using a functional programming approach and it is ispired to Andrea Spano article "R package qdist: A functional programming approach to truncated probability ditributions", which provides an alternative method for computing from truncated probability distribution using a functional programming approach.
 
@@
 
-Why shifted distributions?
-Shift is a distribution trasformation used daily in many statistical fields.
-R does not provide a specific package for shifted distribution models, but it provides package `stats` which contains a wide set of functions for standard distribution models.
-I refer to: probability, density, quantile and random number generator functions, that share a stable naming convention both for the names of the functions and in the first argument of these functions.
+R does not provide a specific package for shifted distribution models, but it provides package `stats` which contains a wide set of functions for standard distribution models: probability, density, quantile and random number generator functions.
+All them share a stable naming convention both for the names of the functions and in the first argument of these functions.
 In particular:
@@
     ```{r sqnorm_example}
-    x <- c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1)
+x <- (1:999)/1000
 not_shifted <- sqnorm(p = x, mean = 8, sd = 1.5)
@@
 
-Spano S. .  _R package qdist: A functional programming approach to truncated probability ditributions_. url
+Spano A. .  _R package qdist: A functional programming approach to truncated probability ditributions_. url
 

### 2016-08-31 docs/article.Rmd
article

@@
 Shifted distributions are used daily in many statistical fields.   
-A shifted distribution is a distribution to which is added a constant, $Y = X + a$, where $X$ and $Y$ are random variables and $a$ is a constant value.
-
+A shifted distribution is a distribution with a constant added, $Y = X + a$, where $X$ and $Y$ are random variables and $a$ is a constant value.  
 The aim of this paper is to describe an R package, `qdtshift` which generates shifted distribution models for any distribution chosen.
-The packege was realized using a functional programming approach and it is ispired to Andrea Spano article "R package qdist: A functional programming approach to truncated probability ditributions", which provides an alternative method for computing from truncated probability distribution using a functional programming approach.
+The package was realized using a functional programming approach and inspires its computational method to Andrea Spanò article "R package qdist: A functional programming approach to truncated probability ditributions", which provides an alternative method for computing from truncated probability distribution using a functional programming approach.
 
@@
 
+* density functions: `d<dist>(x)` with `x` vector of quantiles;
 * probability distribution functions: `p<dist>(q)` with `q` vector of quantiles;
-* density functions: `d<dist>(x)` with `x` vector of quantiles;
 * quantile functions: `q<dist>(p)` with `p` vector of probabilities;
@@
 
-At first view, this approach seems easy and cool but thinking about it I realized that I would need to write a different function for each probability distribution I would like to work with.
-So I decided to use functional programming approach to develop four general functions which compute shifted distribution models for any distribution chosen:
+At first view, this approach seems easy and cool but thinking about it, we realize that we would need to write a different function for each probability distribution we would like to work with.
+Functional programming approach could be the solution, as Andrea Spanò demonstrates in his article (see "R package qdist: A functional programming approach to truncated probability ditributions" for more details).
+
+`qtdshift` package includes four general functions which compute shifted distribution models for any distribution chosen:
 
@@
 
-These functions are included in `qtdshift` package available at [https://github.com/veronicagiro/qshift](https://github.com/veronicagiro/qshift).
+`qtdshift` package is available at [https://github.com/veronicagiro/qshift](https://github.com/veronicagiro/qshift).
 
@@
 
-They share the same logic: they take as input a probability distribution name, as a character string, and return the equivalent shifted distribution as a function object.
+The four functions share the same logic: they take as input a probability distribution name, as a character string, and return the equivalent shifted distribution as a function object.
 
@@
 
-We see in detail the computational process.   
-Let us consider `dshift()`. As we seen, it takes as input the distribution name (`norm`) and proceeds as follows:
+Talking about the computational process, we can say that package `qtdshift` is based on two key concepts being part of the functional nature of the R programming language:
+
+* the environment of a function can be used as a placeholder for other objects;
+* function formals can be manipulated.
+
+For example, let us consider `dshift()`. As we seen, it takes as input the distribution name (`norm`) and proceeds as follows:
 
@@
 
-Package `qtdshift` is based on two key concepts being part of the functional nature of the R programming language:
-
-* the environment of a function can be used as a placeholder for other objects;
-* function formals can be manipulated.
-
-Theoretical references in chapters [Environments](http://www.quantide.com/website/ramarro-chapter-03/) and [Functions](http://www.quantide.com/website/ramarro-chapter-04/) of Andrea Spano weeb Book [Ramarro](http://www.quantide.com/website/ramarro-r-for-developers/), available on Quantide web site.
+Other references are in chapters [Environments](http://www.quantide.com/website/ramarro-chapter-03/) and [Functions](http://www.quantide.com/website/ramarro-chapter-04/) of Andrea Spanò weeb Book [Ramarro](http://www.quantide.com/website/ramarro-r-for-developers/), available on Quantide web site.
 
@@
     ```{r squnif_example}
-    x <- c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1)
+x <- (1:999)/1000
 not_shifted <- squnif(p = x, min = 6, max = 12)
@@
 
-In his article, Andrea Spanò use the newly defined probability distribution functions as a base for defining a function for maximum likelihood estimate for the truncated normal distribution.
-I demonstrate that it is possible also for shifted normal distribution.
-Let us see how.
+Further implementations are possible starting from the definition of a shifted distribution.  
+Suppose we want to define a function for maximum likelihood estimate for the shifted weibull distribution.
 
-We start from the previously defined fucntion `sdnorm()`, that will be used within an estimator function:
+Firstly let us define the shifted density function for the weibull distribution:
+
+```{r sdweibull_def}
+sdweibull <- dshift("weibull")
+```
+
+and, subsequently, by using `sdweibull()` within an estimator function:
 
 ```{r max_likelihood_est}
-ltnorm <- function(x, shift = 0) {
-  theta <- c(mean(x), sd(x))
-  ll <- function(theta, x, shift = 0) {
-    mean <- theta[1]
-    sd <- theta[2]
-    ld <- sdnorm(x = x, mean = mean, sd = sd, shift=shift, log=TRUE)
+ltweibull <- function(x){
+  # starting points for parameters (scale, shape) and shift
+  shift <- min(x) - 0.01
+  x1 <- x - shift
+  shape <- (sd(x1)/mean(x1))^(-1.086)
+  scale <- mean(x1)/gamma(1+1/shape)
+  # parameters vector definition
+  theta <- c(shape, scale, shift)
+  # likelihood function
+  ll <- function(theta, x){
+    shape <- theta[1]
+    scale <- theta[2]
+    shift <- theta[3]
+    ld <- sdweibull(x = x, shape = shape, scale = scale, shift = shift, log = TRUE)
     -sum(ld)
   } 
-  optim(par = theta, fn = ll, x = x, shift=shift)[["par"]]
+  # maximum likelihood estimation
+  optim(par = theta, fn = ll, x = x)[["par"]]
 }
@@
 
+Let us see how it works on a random generated sample from a shifted weibull distribution:
+
 ```{r max_likelihood_est_example}
-# generate a random sample from a shifted normal distribution
-x <- srnorm(n = 1000, mean = 5, sd = 2, shift = 1)
-# maximum likelihood estimate for the shifted normal distribution
-ltnorm(x = x, shift  = 1)
+# generate a random sample from a shifted weibull distribution
+srweibull <- rshift("weibull")
+x <- srweibull(n = 100000, shape = 1, scale = 5, shift = 1)
+
+# maximum likelihood estimate for the shifted weibull distribution
+ltweibull(x = x)
 ```
@@
 
-Spano A. .  _R package qdist: A functional programming approach to truncated probability ditributions_. url
+Spanò A. .  _R package qdist: A functional programming approach to truncated probability ditributions_. url
 

### 2016-08-31 docs/article.Rmd
article

@@
 ---
-title: 'R package qtdshift: A functional programming approach to shifted probability distributions'
+title: 'R package qshift: A functional programming approach to shifted probability distributions'
 author: "Veronica"
 date: "August 25, 2016"
-output: pdf_document
+output:
+  html_document:
+    theme: null
+    highlight: null
 ---
@@
 Shifted distributions are used daily in many statistical fields.  
-A shifted distribution is a distribution with a constant added, $Y = X + a$, where $X$ and $Y$ are random variables and $a$ is a constant value.  
-The aim of this paper is to describe an R package, `qdtshift` which generates shifted distribution models for any distribution chosen.
+Given a random variable $X$ with distribution function $F(x)$, a shifted random variable $Y$ from $X$ is a random variable with a constant, $k$, added, defined as $Y = X+k$, with distribution function $F(x-k)$.
+$Y$ support changes according to shift value.
+
+The aim of this paper is to describe an R package, `qshift` which generates shifted distribution functions for any distribution chosen.
 The package was realized using a functional programming approach and inspires its computational method to Andrea Spanò article "R package qdist: A functional programming approach to truncated probability ditributions", which provides an alternative method for computing from truncated probability distribution using a functional programming approach.
@@
 
-R does not provide a specific package for shifted distribution models, but it provides package `stats` which contains a wide set of functions for standard distribution models: probability, density, quantile and random number generator functions.
+R does not provide a specific package for shifted distribution functions, but it provides package `stats` which contains a wide set of functions for standard distribution functions: probability, density, quantile and random number generator functions.
 All them share a stable naming convention both for the names of the functions and in the first argument of these functions.
@@
 At first view, this approach seems easy and cool but thinking about it, we realize that we would need to write a different function for each probability distribution we would like to work with.
-Functional programming approach could be the solution, as Andrea Spanò demonstrates in his article (see "R package qdist: A functional programming approach to truncated probability ditributions" for more details).
+Functional programming approach could be a solution, as Andrea Spanò demonstrates in his article (see "R package qdist: A functional programming approach to truncated probability ditributions" for more details).
 
-`qtdshift` package includes four general functions which compute shifted distribution models for any distribution chosen:
+`qshift` package includes four general functions which compute shifted distribution models for any distribution chosen:
 
@@
 
-`qtdshift` package is available at [https://github.com/veronicagiro/qshift](https://github.com/veronicagiro/qshift).
+`qshift` package is available at [https://github.com/veronicagiro/qshift](https://github.com/veronicagiro/qshift).
 
@@
 
-```{r qtdshift, message=FALSE}
-require(qtdshift)
+```{r qshift, message=FALSE}
+require(qshift)
 sdnorm <- dshift("norm") # density function
@@
 
-Talking about the computational process, we can say that package `qtdshift` is based on two key concepts being part of the functional nature of the R programming language:
+Talking about the computational process, package `qshift` is based on two key concepts being part of the functional nature of the R programming language:
 
@@
 
-### Examples
+## Examples
 
@@
   ggtitle("Density plot for shifted and non-shifted normal distribution") +
-  theme(legend.title=element_blank(), legend.text= element_text(size=12),
+  theme(legend.title=element_blank(), legend.text= element_text(size=10),
   plot.title = element_text(size=14), axis.title=element_text(size=12))
@@
   ggtitle("Probability plot for shifted and non-shifted normal distribution") +
-  theme(legend.title=element_blank(), legend.text= element_text(size=12),
+  theme(legend.title=element_blank(), legend.text= element_text(size=10),
   plot.title = element_text(size=14), axis.title=element_text(size=12))
@@
   ggtitle("Quantile plot for shifted and non-shifted normal distribution") +
-  theme(legend.title=element_blank(), legend.text= element_text(size=12),
+  theme(legend.title=element_blank(), legend.text= element_text(size=10),
   plot.title = element_text(size=14), axis.title=element_text(size=12))
@@
     ```
+
+As expected, the Normal shifted distribution is a Normal distribution with changed mean. The change on the mean is equal to the shift.
 
@@
 
-```{r qtdshift_unif}
+```{r qshift_unif}
 sdunif <- dshift("unif")
@@
 
-Now the newly created functions are ready for computing from shifted probability models.  
 We limited Uniform support between 6 and 12.
@@
 pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=sdunif_type)) +
-  geom_line(size=1) +
+  geom_line(size=1, alpha =0.8) +
   scale_color_manual(values = c("not_shifted" = "cyan3", 
@@
   ggtitle("Density plot for shifted and non-shifted uniform distribution") +
-  theme(legend.title=element_blank(), legend.text= element_text(size=12),
+  theme(legend.title=element_blank(), legend.text= element_text(size=10),
   plot.title = element_text(size=14), axis.title=element_text(size=12))
@@
 pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=spunif_type)) +
-  geom_line(size=1) +
+  geom_line(size=1, alpha =0.8) +
   scale_color_manual(values = c("not_shifted" = "cyan3", 
@@
   ggtitle("Probability plot for shifted and non-shifted uniform distribution") +
-  theme(legend.title=element_blank(), legend.text= element_text(size=12),
+  theme(legend.title=element_blank(), legend.text= element_text(size=10),
   plot.title = element_text(size=14), axis.title=element_text(size=12))
@@
 pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=squnif_type)) +
-  geom_line(size=1) +
+  geom_line(size=1, alpha =0.8) +
   scale_color_manual(values = c("not_shifted" = "cyan3",  
@@
   ggtitle("Quantile plot for shifted and non-shifted uniform distribution") +
-  theme(legend.title=element_blank(), legend.text= element_text(size=12),
+  theme(legend.title=element_blank(), legend.text= element_text(size=10),
   plot.title = element_text(size=14), axis.title=element_text(size=12))

### 2016-09-01 docs/article.Rmd
articolo definitivo (senza link all'articolo di Andrea)

@@
 ---
-title: 'R package qshift: A functional programming approach to shifted probability distributions'
-author: "Veronica"
-date: "August 25, 2016"
+title: 'R package qshift: A functional programming approach to shifted probability
+  distributions'
 output:
+  pdf_document: default
   html_document:
+    highlight: null
     theme: null
-    highlight: null
 ---
@@
 The aim of this paper is to describe an R package, `qshift` which generates shifted distribution functions for any distribution chosen.
-The package was realized using a functional programming approach and inspires its computational method to Andrea Spanò article "R package qdist: A functional programming approach to truncated probability ditributions", which provides an alternative method for computing from truncated probability distribution using a functional programming approach.
+The package was realized using a functional programming approach and inspires its computational method to Andrea Spanò article "R package `qdist`: A functional programming approach to truncated probability distributions", which provides an alternative method for computing from truncated probability distributions using a functional programming approach.
 
@@
 
-R does not provide a specific package for shifted distribution functions, but it provides package `stats` which contains a wide set of functions for standard distribution functions: probability, density, quantile and random number generator functions.
+R does not provide a specific package for shifted distribution functions, but it provides package `stats` which contains a wide set of functions for standard distributions: probability, density, quantile and random number generation functions.
 All them share a stable naming convention both for the names of the functions and in the first argument of these functions.
@@
 Let us see an example.  
-Considering the Normal distribution (`norm`), we have: `pnorm(q)`, `dnorm(x)`, `qnorm(p)` and `rnorm(n)`.
+Considering the Normal distribution (`norm`), we have: `dnorm(x)`, `pnorm(q)`, `qnorm(p)` and `rnorm(n)`.
 
@@
 At first view, this approach seems easy and cool but thinking about it, we realize that we would need to write a different function for each probability distribution we would like to work with.
-Functional programming approach could be a solution, as Andrea Spanò demonstrates in his article (see "R package qdist: A functional programming approach to truncated probability ditributions" for more details).
+Functional programming approach could be a solution, as Andrea Spanò demonstrates in his article (see "R package `qdist`: A functional programming approach to truncated probability distributions" for more details).
 
-`qshift` package includes four general functions which compute shifted distribution models for any distribution chosen:
+`qshift` package includes four general functions which compute shifted distribution functions for any distribution chosen:
 
@@
 * `qshift()` for shifted quantile functions;
-* `rshift()` for shifted random numbers generator functions.
+* `rshift()` for shifted random numbers generation functions.
 
@@
 ```{r github, eval = FALSE}
-devtools::install_github("veronicagiro/qshift", build=FALSE)
+devtools::install_github("veronicagiro/qshift")
 ```
@@
 * gets the corresponding function `dnorm()`;
-* uses `dnorm()` to create a function, say `density()` that computes shifted probability for normal distributions;
+* uses `dnorm()` to create a function, say `density()` that computes shifted density for normal distributions;
 * modify the formals of `density()` so that it has the same formals as `dnorm()` plus `shift`, corresponding to the shift value;
@@
   xlab("Value") + ylab("Density") +
-  ggtitle("Density plot for shifted and non-shifted normal distribution") +
+  ggtitle("Density plot for shifted and not-shifted Normal distributions") +
   theme(legend.title=element_blank(), legend.text= element_text(size=10),
@@
   xlab("Quantile") + ylab("Probability") +
-  ggtitle("Probability plot for shifted and non-shifted normal distribution") +
+  ggtitle("Probability plot for shifted and not-shifted Normal distributions") +
   theme(legend.title=element_blank(), legend.text= element_text(size=10),
@@
   ylab("Quantile") + xlab("Probability") +
-  ggtitle("Quantile plot for shifted and non-shifted normal distribution") +
+  ggtitle("Quantile plot for shifted and not-shifted Normal distributions") +
   theme(legend.title=element_blank(), legend.text= element_text(size=10),
@@
 
-* Random generator function, `srnorm()`:
+* Random generation function, `srnorm()`:
 
     ```{r srnorm_example}
-    x <- 10000
+    x <- 100000
 not_shifted <- srnorm(n = x, mean = 8, sd = 1.5, set_seed = T)
@@
     ```
+    `srnorm()` and all functions generated by `rshift()` have a further parameter, `set_seed`, which aim is to fix the seed in random number generation, when set as `TRUE`.
+By default it is set as `FALSE`.
+In this case it is set as `TRUE` in order to fix the seed for visualizing the differences between random generated distributions.
     We can visualize these results by plotting them:
@@
                                 "negative_shifted" = "#FF89B4")) +
-  ggtitle("Density plot for shifted and non-shifted normal distribution") +
+  ggtitle("Density plot for shifted and not-shifted Normal distributions") +
   theme(legend.title=element_blank(), legend.text= element_text(size=10),
@@
 
-As expected, the Normal shifted distribution is a Normal distribution with changed mean. The change on the mean is equal to the shift.
+As expected, the Normal shifted distribution is a Normal distribution with changed mean. The change in the mean is equal to the shift.
 
@@
   xlab("Value") + ylab("Density") +
-  ggtitle("Density plot for shifted and non-shifted uniform distribution") +
+  ggtitle("Density plot for shifted and not-shifted Uniform distributions") +
   theme(legend.title=element_blank(), legend.text= element_text(size=10),
@@
   xlab("Quantile") + ylab("Probability") +
-  ggtitle("Probability plot for shifted and non-shifted uniform distribution") +
+  ggtitle("Probability plot for shifted and not-shifted Uniform distributions") +
   theme(legend.title=element_blank(), legend.text= element_text(size=10),
@@
   ylab("Quantile") + xlab("Probability") +
-  ggtitle("Quantile plot for shifted and non-shifted uniform distribution") +
+  ggtitle("Quantile plot for shifted and not-shifted Uniform distributions") +
   theme(legend.title=element_blank(), legend.text= element_text(size=10),
@@
 
-* Random generator function, `srunif()`:
+* Random generation function, `srunif()`:
 
@@
 Further implementations are possible starting from the definition of a shifted distribution.  
-Suppose we want to define a function for maximum likelihood estimate for the shifted weibull distribution.
+Suppose we want to define a function for maximum likelihood estimate for the shifted Weibull distribution.
 
-Firstly let us define the shifted density function for the weibull distribution:
+Firstly let us define the shifted density function for the Weibull distribution:
 
@@
 ltweibull <- function(x){
-  # starting points for parameters (scale, shape) and shift
+  # starting values for parameters (scale, shape) and shift
   shift <- min(x) - 0.01
@@
 
-Let us see how it works on a random generated sample from a shifted weibull distribution:
+Let us see how it works on a random generated sample from a shifted Weibull distribution:
 

### 2016-09-06 R/rshift.R
vignette ok

@@ -30,7 +30,7 @@
 #' @examples
 #'
 #' # Binomial Distribution
-#' rb <- rshift("binom", set_seed =F)
+#' rb <- rshift("binom")
 #' set.seed(1)
 #' zero_shift <- rb(n = 10000, size = 6, prob = 0.5, shift = 0)
 #' zero_shift

### 2016-09-06 vignettes/qshift.R

@@ -1,0 +1,271 @@
+## ---- echo=TRUE, message=FALSE-------------------------------------------
+knitr::opts_chunk$set(fig.width = 7, fig.height = 7)
+
+## ----dnorm---------------------------------------------------------------
+dnorm(x = 6:10, mean = 8, sd = 1)
+
+## ----sdnorm_definition---------------------------------------------------
+shift_dnorm <- function(x, mean = 0, sd = 1, shift = 0, ...){
+            x_shifted <- x - shift
+            density <- dnorm(x = x_shifted, mean = mean, sd = sd, ...)
+            return(density)
+}
+
+## ----sdnorm_example_1----------------------------------------------------
+shift_dnorm(x = 6:10, mean = 8, sd = 1, shift = 1)
+
+## ----qshift, message=FALSE-----------------------------------------------
+require(qshift)
+sdnorm <- dshift("norm") # density function
+spnorm <- pshift("norm") # probability function
+sqnorm <- qshift("norm") # quantile function
+srnorm <- rshift("norm") # random generation function
+
+## ----args----------------------------------------------------------------
+args(sdnorm)
+args(dnorm)
+
+## ----sdnorm_example------------------------------------------------------
+x <- seq(4, 14, len = 100)
+not_shifted <- sdnorm(x = x, mean = 8, sd = 1.5)
+positive_shifted <- sdnorm(x = x, mean = 8, sd = 1.5, shift = 1)
+negative_shifted <- sdnorm(x = x, mean = 8, sd = 1.5, shift = -1)
+
+## ----sdnorm_plot, message=FALSE------------------------------------------
+require(ggplot2)
+df <- data.frame(x=x, 
+             sdnorm_type = factor(c(rep("not_shifted", times=length(x)),
+                rep("positive_shifted", times=length(x)),
+                rep("negative_shifted", times=length(x))),
+                levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
+              value = c(not_shifted, positive_shifted, negative_shifted))
+
+pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=sdnorm_type)) +
+  geom_line(size=1) +
+  scale_color_manual(values = c("not_shifted" = "cyan3", 
+                            "positive_shifted" = "#89FF99", 
+                            "negative_shifted" = "#FF89B4")) +
+  xlab("Value") + ylab("Density") +
+  ggtitle("Density plot for shifted and not-shifted Normal distributions") +
[…]

### 2016-09-12 docs/prove-qdist+qshift.R
prove x articolo distribuzioni shiftate + troncate

@@ -1,0 +1,931 @@
+require(qdist)
+require(qshift)
+require(ggplot2)
+###################################################
+# definire tutti le funzioni per le distribuzioni #
+###################################################
+
+# Normale
+
+dtnorm <- dtruncate("norm")
+ptnorm <- ptruncate("norm")
+qtnorm <- qtruncate("norm")
+rtnorm <- rtruncate("norm")
+
+dsnorm <- dshift("norm")
+psnorm <- pshift("norm")
+qsnorm <- qshift("norm")
+rsnorm <- rshift("norm")
+
+############# d ################
+
+#############################
+# prima tronco e poi shifto #
+#############################
+a <- dnorm(x = 1:10, mean = 5, sd = 3)  
+dtnorm <- dtruncate("norm")
+b <- dtnorm(x = 1:10, mean = 5, sd = 3, L = 2, U = 8)
+
+dtsnorm <- dshift("tnorm")
+
+c <- dtsnorm(x = 1:10, mean = 5, sd = 3, shift=2, L = 2, U = 8)
+
+df <- data.frame(a,b,c,d =1:10, stringsAsFactors = F)
+
+pl <- ggplot() + 
+  geom_line(data = df, mapping = aes(y=a, x=d), col="blue")+
+  geom_line(data = df, mapping = aes(y=b, x=d), col="red")+
+  geom_line(data = df, mapping = aes(y=c, x=d), col="green")
+
+pl
+#############################
+# prima shifto e poi tronco #
+#############################
+
+dsnorm <- dshift("norm")
+b <- dsnorm(x = 1:10, mean = 5, sd = 3, shift=2)
+
+dtsnorm <- dtruncate("snorm")
+c <- dtsnorm(x = 1:10, mean = 5, sd = 3, shift=2, L = 2, U = 8)
+
+df <- data.frame(a,b,c,d =1:10, stringsAsFactors = F)
+
+pl <- ggplot() + 
+  geom_line(data = df, mapping = aes(y=a, x=d), col="blue") +
+  geom_line(data = df, mapping = aes(y=b, x=d), col="red") +
+  geom_line(data = df, mapping = aes(y=c, x=d), col="green")
+
+pl
+
+
+############# p ################
+
+#############################
+# prima tronco e poi shifto #
+#############################
+a <- pnorm(q = 1:10, mean = 5, sd = 3)  
[…]

### 2016-09-12 docs/article-shift+trunc.Rmd
articolo distr shiftate e troncate

---
title: "distr shiftate + troncate"
author: "Veronica"
date: "September 12, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 2016-09-12 docs/prove-qdist+qshift.R
test

@@ -22,7 +22,7 @@ rsnorm <- rshift("norm")
 #############################
 # prima tronco e poi shifto #
 #############################
-a <- dnorm(x = 1:10, mean = 5, sd = 3)  
+a <- dnorm(x = 1:10, mean = 5, sd = 3)
 dtnorm <- dtruncate("norm")
 b <- dtnorm(x = 1:10, mean = 5, sd = 3, L = 2, U = 8)
 
@@ -32,7 +32,7 @@ c <- dtsnorm(x = 1:10, mean = 5, sd = 3, shift=2, L = 2, U = 8)
 
 df <- data.frame(a,b,c,d =1:10, stringsAsFactors = F)
 
-pl <- ggplot() + 
+pl <- ggplot() +
   geom_line(data = df, mapping = aes(y=a, x=d), col="blue")+
   geom_line(data = df, mapping = aes(y=b, x=d), col="red")+
   geom_line(data = df, mapping = aes(y=c, x=d), col="green")
@@ -50,7 +50,7 @@ c <- dtsnorm(x = 1:10, mean = 5, sd = 3, shift=2, L = 2, U = 8)
 
 df <- data.frame(a,b,c,d =1:10, stringsAsFactors = F)
 
-pl <- ggplot() + 
+pl <- ggplot() +
   geom_line(data = df, mapping = aes(y=a, x=d), col="blue") +
   geom_line(data = df, mapping = aes(y=b, x=d), col="red") +
   geom_line(data = df, mapping = aes(y=c, x=d), col="green")
@@ -63,7 +63,7 @@ pl
 #############################
 # prima tronco e poi shifto #
 #############################
-a <- pnorm(q = 1:10, mean = 5, sd = 3)  
+a <- pnorm(q = 1:10, mean = 5, sd = 3)
 ptnorm <- ptruncate("norm")
 b <- ptnorm(q = 1:10, mean = 5, sd = 3, L = 2, U = 8)
 
@@ -73,7 +73,7 @@ c <- ptsnorm(q = 1:10, mean = 5, sd = 3, shift=2, L = 2, U = 8)
 
 df <- data.frame(a,b,c,d =1:10, stringsAsFactors = F)
 
-pl <- ggplot() + 
+pl <- ggplot() +
   geom_line(data = df, mapping = aes(y=a, x=d), col="blue")+
   geom_line(data = df, mapping = aes(y=b, x=d), col="red")+
   geom_line(data = df, mapping = aes(y=c, x=d), col="green")
[…]

### 2016-09-14 docs/article-shift+trunc.Rmd
articolo distr tronc+shift

@@
 ```
+
+R programming functional nature allows us programmer to implement and manipulate functions as we need.
+Functional programming is a style of programming that emphasizes the evaluation of expressions rather than the execution of commands.
+It has a great power and lots of properties that combined together can allows us to produce results with the most efficient computational method.
+In this article we will consider the functional factory method proposed by Andrea Spanò in "R package qdist: A functional programming approach to truncated probability ditributions", and used by Veronica Giro in "R package qshift: A functional programming approach to shifted probability distributions", that produces functions for truncated and shifted distributions, respectively.
+[`qdist`](https://github.com/andreaspano/qdist) and [`qshift`](https://github.com/veronicagiro/qshift) packages are available on Github.
+We will demonstrate that it is possible to produce functions for truncated and shifted or shifted and truncated distributions only combining the two packages.
+
+# Functional programming as a new approach for computing distribution functions
+
+R provides probability, density, quantile and random number generator functions in `stats` package, using a stable naming convention both for the names of the functions and in the first argument of these functions:
+
+* probability distribution functions p<dist>(q) with q vector of quantiles;
+* density functions d<dist>(x) with x vector of quantiles;
+* quantile functions q<dist>(p) with p vector of probabilities;
+* random number generation r<dist>(n) with n number of observations.
+
+where <dist> indicates the distribution family.
+
+As a result we have: pnorm(q), dnorm(x), qnorm(p) and rnorm(n) for normal distribution and similarly for all other distributions.
+
+Starting from this well structured syntax, Andrea Spanò proposed a functional programming method that computes truncated distributions functions for each distribution chosen without changing the programming style and syntax of previous functions.
+
+The methodology is based in two steps:
+
+* define the required functions suitable for computing from specific truncated probability models by using the tools available within its package:
+
+```{r}
+#devtools::install_github("andreaspano/qdist", build=FALSE)
+require(qdist)
+tpnorm <- ptruncate("norm") 
+tdnorm <- dtruncate("norm")
+tqnorm <- qtruncate("norm")
+trnorm <- rtruncate("norm")
+```
+    These functions take as input a probability distribution as a character string, in this case the normal distribution, and
+return the equivalent truncated distribution as a function object.
+
+* use the newly created functions for computing from truncated probability models.
+
+```{r}
+q <- seq(6, 14, len = 100)
+ptnorm(q, mean = 10, sd = 1, L = 9, U = 11)
+```
+    we use function tpnorm() to compute the distribution function of a normal distribution with mean 10 and standard deviation 1, truncated in 9 and 11.
+
+The newly generated function `ptnorm()` has the same formals as the original `pnorm()`, plus two extra parameters: `L` and `U`, respectively for lower and upper truncation thresholds set by default to `-Inf` and `+Inf`:
+
+```{r}
+args(pnorm)
+args(tpnorm)
+```
+
+This article will explain an alternative computational method based on function factory and on two R key concepts:
+
+• the environment of a function can be used as a placeholder for other objects;
+• function formals can be manipulated.

### 2016-09-14 docs/article-shift+trunc.Rmd
article shift+trunc

@@
 ```
+
+# Introduction
 
@@
 
-Starting from this well structured syntax, Andrea Spanò proposed a functional programming method that computes truncated distributions functions for each distribution chosen without changing the programming style and syntax of previous functions.
+Starting from this well structured syntax, Andrea Spanò proposed, in `qdist`, a functional programming method that computes truncated distributions functions for each distribution chosen without changing the programming style and syntax of previous functions.
 
-The methodology is based in two steps:
+The methodology is based on two R key concepts:
 
-* define the required functions suitable for computing from specific truncated probability models by using the tools available within its package:
+* the environment of a function can be used as a placeholder for other objects;
+* function formals can be manipulated.
+
+and develops in two steps:
+
+* define the required functions suitable for computing from specific truncated probability models by using the functions available within its package.
+These functions take as input a probability distribution as a character string, in this case the normal distribution, and return the equivalent truncated distribution as a function object.
 
@@
 require(qdist)
-tpnorm <- ptruncate("norm") 
-tdnorm <- dtruncate("norm")
-tqnorm <- qtruncate("norm")
-trnorm <- rtruncate("norm")
+tpnorm <- ptruncate("norm") # probability function 
+tdnorm <- dtruncate("norm") # density function
+tqnorm <- qtruncate("norm") # quantile function
+trnorm <- rtruncate("norm") # random generation function
 ```
-    These functions take as input a probability distribution as a character string, in this case the normal distribution, and
-return the equivalent truncated distribution as a function object.
 
-* use the newly created functions for computing from truncated probability models.
+* use the newly created functions (`tpnorm()`, `tdnorm()`, `tqnorm()`, `trnorm()`) for computing from truncated probability models.
 
@@
 ```
-    we use function tpnorm() to compute the distribution function of a normal distribution with mean 10 and standard deviation 1, truncated in 9 and 11.
+    In this example we use function `tpnorm()` to compute the distribution function of a normal distribution with mean 10 and standard deviation 1, truncated in 9 and 11.
 
@@
 
-This article will explain an alternative computational method based on function factory and on two R key concepts:
+`qshift` package share the same logic and the same methodology. The difference is that the functions are designed for shifted distributions.
 
-• the environment of a function can be used as a placeholder for other objects;
-• function formals can be manipulated.
+Also in this case, the methodology develops in two steps:
+
+* define the required functions suitable for computing from specific shifted probability models by using the tools available within its package.
+These functions take as input a probability distribution as a character string, in this case the normal distribution, and return the equivalent truncated distribution as a function object.
+
+```{r}
+require(qshift)
+dsnorm <- dshift("norm") # density function
+psnorm <- pshift("norm") # probability function
+qsnorm <- qshift("norm") # quantile function
+rsnorm <- rshift("norm") # random generation function
+```
+
+* se the newly created functions (`tsnorm()`, `tsnorm()`, `tsnorm()`, `tsnorm()`) for computing from truncated probability models.
+```{r}
+x <- seq(4, 14, len = 100)
+positive_shifted <- psnorm(q = x, mean = 8, sd = 1.5, shift = 1)
+```
+
+The newly generated function `psnorm()` has the same formals as the original `pnorm()`, plus one extra parameter: `shift`, set by default to 0:
+
+```{r}
+args(pnorm)
+args(tpnorm)
+```
+
+# Distribution function for truncated and shifted or for shifted and truncated distributions
+
+Functional programming approach used to built the packages is the key that allows to obtain distribution functions for truncated and shifted or for shifted and truncated distributions.
+
+Let us consider the previously defined functions for truncated and shifted distributions functions.
+
+Suppose we want to generate the probability functions for truncated shifted normal distributions.
+
+As `pnorm()` is the function that computes the distribution function for a normal distributions, named `norm`, `ptnorm()`, defined as
+
+```{r}
+ptnorm <- ptruncate("norm")
+```
+
+is the function that computes the distribution function for a truncated normal distributions, named `tnorm`.
+
+So for computing the distribution function of a shifted trucated normal distribution we have to pass `tnorm` distribution as `pshift()` argument:
+
+```{r}
+pstnorm <- pshift("tnorm")
+```
+
+So `pstnorm()` is the function that computes the distribution function of `stnorm` shifted truncated normal distribution.
+
+```{r}
+args(pstnorm)
+```
+
+Looking at the function arguments we see that the sintax style is that of `pnorm()` function plus `L` and `U` arguments for truncated distribution and `shift` argument for shifted distribution.
+
+The inverse procedure are necessary to generate the distribution function for a truncated shifted distribution.
+
+Starting from `pshift` we define:
+
+```{r}
+psnorm <- pshift("norm")
+```
+
+`psnorm()` is the distrinìbution function of `snorm` shifted normal distribution.
+
+So for computing the distribution function of a trucated shifted normal distribution we have to pass `snorm` distribution as `ptruncate()` argument:
+
+```{r}
+ptsnorm <- ptruncate("snorm")
+```
+
+So `ptsnorm()` is the function that computes the distribution function of `stnorm` truncated shifted normal distribution.

### 2016-09-14 docs/article-shift+trunc.Rmd
articolo distr shift + trunc

@@
 
+For more details and examples about `qdist` and `qshift` see articles "R package qdist: A functional programming approach to truncated probability ditributions" and "R package qshift: A functional programming approach to shifted probability distributions" available on [Quantide blog](http://www.quantide.com/blog/).
+
 # Distribution function for truncated and shifted or for shifted and truncated distributions
 
-Functional programming approach used to built the packages is the key that allows to obtain distribution functions for truncated and shifted or for shifted and truncated distributions.
+In the previous paragraph we saw how `qdist` and `qshift` packages works and that they are built with a functional programming approach.
+The functional programming approach whereby the packages are built is the key that allows to obtain distribution functions for truncated and shifted or for shifted and truncated distributions.
+Let us see how!
 
-Let us consider the previously defined functions for truncated and shifted distributions functions.
-
-Suppose we want to generate the probability functions for truncated shifted normal distributions.
-
-As `pnorm()` is the function that computes the distribution function for a normal distributions, named `norm`, `ptnorm()`, defined as
-
+Suppose we want to generate the probability functions for a truncated shifted normal distribution.  
+As `pnorm()` is the function that computes the distribution function for a normal distributions, conventionally named `norm`, `ptnorm()`, defined as: 
 ```{r}
@@
 
-So `pstnorm()` is the function that computes the distribution function of `stnorm` shifted truncated normal distribution.
+So `pstnorm()` is the function that computes the distribution function of `stnorm`, the shifted truncated normal distribution.
 
@@
 
-The inverse procedure are necessary to generate the distribution function for a truncated shifted distribution.
+The inverse procedure are necessary to generate the distribution function for a shifted truncated distribution.
 
@@
 
-`psnorm()` is the distrinìbution function of `snorm` shifted normal distribution.
-
-So for computing the distribution function of a trucated shifted normal distribution we have to pass `snorm` distribution as `ptruncate()` argument:
+`psnorm()` is the distribution function of `snorm`, a shifted normal distribution.
+The distribution function of a truncated shifted normal distribution is generated passing `snorm` distribution as `ptruncate()` argument:
 

### 2016-09-14 docs/article-shift+trunc.Rmd
articolo distr trunc+shift

--- 2016-09-14 docs/article-shift+trunc.Rmd
@@
 
+# Introduction
+
 R programming functional nature allows us programmer to implement and manipulate functions as we need.
@@
 It has a great power and lots of properties that combined together can allows us to produce results with the most efficient computational method.
-In this article we will consider the functional factory method proposed by Andrea Spanò in "R package qdist: A functional programming approach to truncated probability ditributions", and used by Veronica Giro in "R package qshift: A functional programming approach to shifted probability distributions", that produces functions for truncated and shifted distributions, respectively.
-[`qdist`](https://github.com/andreaspano/qdist) and [`qshift`](https://github.com/veronicagiro/qshift) packages are available on Github.
-We will demonstrate that it is possible to produce functions for truncated and shifted or shifted and truncated distributions only combining the two packages.
+In this article we will consider the functional factory method proposed by Andrea Spanò in "R package qdist: A functional programming approach to truncated probability ditributions", and that I used in "R package qshift: A functional programming approach to shifted probability distributions", that produces functions for truncated and shifted distributions, respectively.
+[`qdist`](https://github.com/andreaspano/qdist) and [`qshift`](https://github.com/veronicagiro/qshift) packages are available on Github:
+
+```{r, eval=FALSE}
+devtools::install_github("andreaspano/qdist")
+devtools::install_github("veronicagiro/qshift")
+```
+
+We will demonstrate that it is possible to produce functions for truncated and shifted or for shifted and truncated distributions only combining the functions of the two packages.
 
 # Functional programming as a new approach for computing distribution functions
+
+Let us sum up the alternative functional programming approach used for developing `qdist` and `qshift`.
 
@@
 
-Starting from this well structured syntax, Andrea Spanò proposed a functional programming method that computes truncated distributions functions for each distribution chosen without changing the programming style and syntax of previous functions.
+Starting from this well structured syntax, Andrea Spanò proposed in `qdist` a functional programming method that computes truncated distributions functions for each distribution chosen without changing the programming style and syntax of previous functions.
 
-The methodology is based in two steps:
+The methodology is based on two R key concepts:
 
-* define the required functions suitable for computing from specific truncated probability models by using the tools available within its package:
+* the environment of a function can be used as a placeholder for other objects;
+* function formals can be manipulated.
+
+and develops itself in two steps:
+
+* defines the required functions suitable for computing from specific truncated probability distribution by using the functions available within its package.
+These functions take as input a probability distribution as a character string, in this case the normal distribution, and return the equivalent truncated distribution function as object:
+
+```{r, message=FALSE}
+require(qdist)
+dtnorm <- dtruncate("norm") # density function
+ptnorm <- ptruncate("norm") # probability function 
+qtnorm <- qtruncate("norm") # quantile function
+rtnorm <- rtruncate("norm") # random generation function
+```
+
+* uses the newly created functions (`tpnorm()`, `tdnorm()`, `tqnorm()`, `trnorm()`) for computing from truncated probability distributions.
+
+Let us consider `dtnorm()`.
+It has the same formals as the original `dnorm()`, plus two extra parameters: `L` and `U`, respectively for lower and upper truncation thresholds set by default to `-Inf` and `+Inf`:
 
 ```{r}
-#devtools::install_github("andreaspano/qdist", build=FALSE)
-require(qdist)
-tpnorm <- ptruncate("norm") 
-tdnorm <- dtruncate("norm")
-tqnorm <- qtruncate("norm")
-trnorm <- rtruncate("norm")
+args(dnorm)
+args(dtnorm)
 ```
-    These functions take as input a probability distribution as a character string, in this case the normal distribution, and
-return the equivalent truncated distribution as a function object.
 
-* use the newly created functions for computing from truncated probability models.
+Let us see graphically the differences between `dnorm()` to compute the density function of a normal distribution with mean 10 and standard deviation 1 and `tdnorm()` to compute the density function of a normal distribution with mean 10 and standard deviation 1 lower truncated at 9 and upper truncated at 11:
 
 ```{r}
-q <- seq(6, 14, len = 100)
-ptnorm(q, mean = 10, sd = 1, L = 9, U = 11)
+x <- seq(6, 14, len = 100)
+std_density <- dnorm(x = x, mean = 10, sd = 1)
+trunc_density <- dtnorm(x = x, mean = 10, sd = 1, L = 9, U = 11)
+
+df <- data.frame(x=x, 
+                 type = factor(c(rep("std_density", times=length(x)),
+                    rep("trunc_density", times=length(x))),
+                    levels = c("std_density", "trunc_density")), 
+                  value = c(std_density, trunc_density))
+
+require(ggplot2)
+pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
+  geom_line(size=1) +
+  scale_color_manual(values = c("std_density" = "cyan3", 
+                                "trunc_density" = "#89FF99")) +
+  xlab("Value") + ylab("Density") +
+  ggtitle("Density plot for non-truncated and truncated Normal distributions") +
+  theme(legend.title=element_blank(), legend.text= element_text(size=10),
+  plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
 ```
-    we use function tpnorm() to compute the distribution function of a normal distribution with mean 10 and standard deviation 1, truncated in 9 and 11.
 
-The newly generated function `ptnorm()` has the same formals as the original `pnorm()`, plus two extra parameters: `L` and `U`, respectively for lower and upper truncation thresholds set by default to `-Inf` and `+Inf`:
+`qshift` package shares the same logic and the same methodology of `qdist`. The difference is that the functions are designed to compute functions for shifted distributions.
+
+The methodology is made up of the previously descripted two steps. So it:
+
+* defines the required functions suitable for computing from specific shifted probability models by using the `qshift` functions, that takes as input a probability distribution as a character string, in this case the normal distribution, and returns the equivalent shifted distribution function as object.
+
+```{r, message=FALSE}
+require(qshift)
+dsnorm <- dshift("norm") # density function
+psnorm <- pshift("norm") # probability function
+qsnorm <- qshift("norm") # quantile function
+rsnorm <- rshift("norm") # random generation function
+```
+
+* use the newly created functions (`dsnorm()`, `psnorm()`, `qsnorm()`, `rsnorm()`) for computing from shifted probability distributions.
+
+Let us consider `dsnorm()`. Also this function has the same formals as the original `dnorm()`, plus one extra parameter: `shift`, set by default to 0:
 
 ```{r}
-args(pnorm)
-args(tpnorm)
+args(dnorm)
+args(dtnorm)
 ```
 
-This article will explain an alternative computational method based on function factory and on two R key concepts:
+Let us see graphically the differences between `dnorm()` to compute the density function of a normal distribution with mean 10 and standard deviation 1 and `sdnorm()` to compute the density function of a normal distribution with mean 10 and standard deviation 1 shifted of one unit:
 
-• the environment of a function can be used as a placeholder for other objects;
-• function formals can be manipulated.
+```{r}
+x <- seq(6, 14, len = 100)
+std_density <- dnorm(x = x, mean = 10, sd = 1)
+shift_density <- dsnorm(x = x, mean = 10, sd = 1, shift = 1)
+
+df <- data.frame(x=x, 
+                 type = factor(c(rep("std_density", times=length(x)),
+                    rep("shift_density", times=length(x))),
+                    levels = c("std_density", "shift_density")), 
+                  value = c(std_density, shift_density))
+
+pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
+  geom_line(size=1) +
+  scale_color_manual(values = c("std_density" = "cyan3", 
+                                "shift_density" = "#FF89B4")) +
+  xlab("Value") + ylab("Density") +
+  ggtitle("Density plot for non-shifted and shifted Normal distributions") +
+  theme(legend.title=element_blank(), legend.text= element_text(size=10),
+  plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+```
+
+For more details and examples about how `qdist` and `qshift` works, see articles "R package qdist: A functional programming approach to truncated probability ditributions" and "R package qshift: A functional programming approach to shifted probability distributions" available on [Quantide blog](http://www.quantide.com/blog/).
+
+# Distribution function for truncated and shifted or for shifted and truncated distributions
+
+The functional programming approach whereby `qdist` and `qshift` packages are built is the key that allows to obtain distribution functions for truncated and shifted or for shifted and truncated distributions.
+Let us see how!
+
+Suppose we want to generate the density functions for a truncated shifted normal distribution.  
+As `dnorm()` is the function that computes the density function for a normal distributions, conventionally named `norm`, `dtnorm()`, defined as: 
+```{r}
+dtnorm <- dtruncate("norm")
+```
+
+is the function that computes the density function for a truncated normal distributions, named `tnorm`.
+
+We have to define also the others functions for truncated distribution, as they are necessary for the further step:
+
+```{r}
+ptnorm <- ptruncate("norm")
+qtnorm <- qtruncate("norm")
+rtnorm <- rtruncate("norm")
+```
+
+So for computing the density function of a shifted trucated normal distribution we have to pass `tnorm` distribution as `dshift()` argument:
+
+```{r}
+dstnorm <- dshift("tnorm")
+```
+
+So `dstnorm()` is the function that computes the density function of `stnorm`, which is the shifted truncated normal distribution.
+
+```{r}
+args(dstnorm)
+```
+
+Looking at the function arguments we see that the sintax style is that of `dnorm()` function plus `L` and `U` arguments for truncated distribution and `shift` argument for shifted distribution.
+
+The inverse procedure generates the density function for a shifted truncated distribution.
+
+Starting from `dshift`, we define `dsnorm()`, the density function of `snorm`, which is the shifted normal distribution:
+```{r}
+dsnorm <- dshift("norm")
+```
+
+We have to define also the others functions for shifted distribution, as they are necessary for the further step:
+
+```{r}
+psnorm <- pshift("norm")
+qsnorm <- qshift("norm")
+rsnorm <- rshift("norm")
+```
+
+The density function of a truncated shifted normal distribution is generated passing `snorm` distribution as `ptruncate()` argument:
+```{r}
+dtsnorm <- dtruncate("snorm")
+```
+
+So `ptsnorm()` is the function that computes the density function of `stnorm`, which is the truncated shifted normal distribution.
+
+Let us see graphically the differences between the density function of normal distribution, of a shifted truncated normal distribution and of a truncated shifted normal distribution.
+
+```{r}
+std_density <- dnorm(x = x, mean = 10, sd = 1)
+st_density <- dstnorm(x = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
+ts_density <- dtsnorm(x = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
+```
+
+```{r}
+df <- data.frame(x=x, 
+                 type = factor(c(rep("std_density", times=length(x)),
+                    rep("st_density", times=length(x)),
+                    rep("ts_density", times=length(x))),
+                    levels = c("std_density", "st_density", "ts_density")), 
+                  value = c(std_density, st_density, ts_density))
+
+pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
+  geom_line(size=1) +
+  scale_color_manual(values = c("std_density" = "cyan3", 
+                                "st_density" = "#89FF99", 
+                                "ts_density" = "#FF89B4")) +
+  xlab("Value") + ylab("Density") +
+  ggtitle("Density plot for transformed and not-transformed Normal distributions") +
+  theme(legend.title=element_blank(), legend.text= element_text(size=10),
+  plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+```
+
+We can conclude that the previous functiona are completely different.

### 2016-09-14 docs/article-shift+trunc.Rmd
articolo distr trunc + shift + prove

@@
 
-* probability distribution functions p<dist>(q) with q vector of quantiles;
-* density functions d<dist>(x) with x vector of quantiles;
-* quantile functions q<dist>(p) with p vector of probabilities;
-* random number generation r<dist>(n) with n number of observations.
+* probability distribution functions p< dist >(q) with q vector of quantiles;
+* density functions d< dist >(x) with x vector of quantiles;
+* quantile functions q< dist >(p) with p vector of probabilities;
+* random number generation r< dist >(n) with n number of observations.
 
-where <dist> indicates the distribution family.
+where < dist > indicates the distribution family.
 
@@
 
-```{r}
+```{r, message=FALSE}
 x <- seq(6, 14, len = 100)
@@
 args(dnorm)
-args(dtnorm)
+args(dsnorm)
 ```
@@
 
-# Distribution function for truncated and shifted or for shifted and truncated distributions
+# Distribution functions for truncated and shifted or for shifted and truncated distributions
 
@@
 
-We have to define also the others functions for truncated distribution, as they are necessary for the further step:
+We have to define also the others functions for truncated distribution, as they are necessary implicitly for the further step:
 
@@
 
-We have to define also the others functions for shifted distribution, as they are necessary for the further step:
+We have to define also the others functions for shifted distribution, as they are implicitly necessary for the further step:
 
@@
 
-So `ptsnorm()` is the function that computes the density function of `stnorm`, which is the truncated shifted normal distribution.
+So `dtsnorm()` is the function that computes the density function of `stnorm`, which is the truncated shifted normal distribution.
 
@@
 
-We can conclude that the previous functiona are completely different.
+We can conclude that the previous functions are completely different.
+
+Let us analyze the other distributions functions:
+
+* probability distribution function of truncated-shifted normal distribution and of shifted-truncated normal distribution:
+
+```{r}
+pstnorm <- pshift("tnorm")
+ptsnorm <- ptruncate("snorm")
+std_distr <- pnorm(q = x, mean = 10, sd = 1)
+st_distr <- pstnorm(q = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
+ts_distr <- ptsnorm(q = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
+```
+
+```{r}
+df <- data.frame(x=x, 
+                 type = factor(c(rep("std_distr", times=length(x)),
+                    rep("st_distr", times=length(x)),
+                    rep("ts_distr", times=length(x))),
+                    levels = c("std_distr", "st_distr", "ts_distr")), 
+                  value = c(std_distr, st_distr, ts_distr))
+
+pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
+  geom_line(size=1) +
+  scale_color_manual(values = c("std_distr" = "cyan3", 
+                                "st_distr" = "#89FF99", 
+                                "ts_distr" = "#FF89B4")) +
+  xlab("Value") + ylab("Density") +
+  ggtitle("Distribution plot for transformed and not-transformed Normal distributions") +
+  theme(legend.title=element_blank(), legend.text= element_text(size=10),
+  plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+```
+
+* quantile function of truncated-shifted normal distribution and of shifted-truncated normal distribution:
+
+```{r}
+qstnorm <- qshift("tnorm")
+qtsnorm <- qtruncate("snorm")
+x <- 1:999/1000
+std_quant <- qnorm(p = x, mean = 10, sd = 1)
+st_quant <- qstnorm(p = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
+ts_quant <- qtsnorm(p = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
+```
+
+```{r}
+df <- data.frame(x=x, 
+                 type = factor(c(rep("std_quant", times=length(x)),
+                    rep("st_quant", times=length(x)),
+                    rep("ts_quant", times=length(x))),
+                    levels = c("std_quant", "st_quant", "ts_quant")), 
+                  value = c(std_quant, st_quant, ts_quant))
+
+pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
+  geom_line(size=1) +
+  scale_color_manual(values = c("std_quant" = "cyan3", 
+                                "st_quant" = "#89FF99", 
+                                "ts_quant" = "#FF89B4")) +
+  xlab("Value") + ylab("Density") +
+  ggtitle("Distribution plot for transformed and not-transformed Normal distributions") +
+  theme(legend.title=element_blank(), legend.text= element_text(size=10),
+  plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+```
+
+* random generation function of truncated-shifted normal distribution and of shifted-truncated normal distribution:
+
+```{r}
+rstnorm <- rshift("tnorm")
+rtsnorm <- rtruncate("snorm")
+std_rand <- rnorm(n=1000, mean = 10, sd = 1)
+st_rand <- rstnorm(n=1000, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
+ts_rand <- rtsnorm(n=1000, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
+```
+
+```{r}
+df <- data.frame(x=1:1000, 
+                 type = factor(c(rep("std_rand", times=1000),
+                    rep("st_rand", times=1000),
+                    rep("ts_rand", times=1000)),
+                    levels = c("std_rand", "st_rand", "ts_rand")), 
+                  value = c(std_rand, st_rand, ts_rand))
+
+pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
+  geom_line(size=1) +
+  scale_color_manual(values = c("std_rand" = "cyan3", 
+                                "st_rand" = "#89FF99", 
+                                "ts_rand" = "#FF89B4")) +
+  xlab("Value") + ylab("Density") +
+  ggtitle("Distribution plot for transformed and not-transformed Normal distributions") +
+  theme(legend.title=element_blank(), legend.text= element_text(size=10),
+  plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+```
+
+<!--
+pensare ad un modo diverso per spiegare la cosa perchè da questi grafici nn si capisce niente
+-->
+## Other Examples
+
+Let us see how it the other distribution functions works with other distributions.
+
+Let us consider Poisson distribution and distribution function.
+
+First of all functions for truncated distribution and for shifted distributio have to be defined:
+
+```{r}
+dtpois <- dtruncate("pois")
+ptpois <- ptruncate("pois")
+qtpois <- qtruncate("pois")
+rtpois <- rtruncate("pois")
+
+dspois <- dshift("pois")
+pspois <- pshift("pois")
+qspois <- qshift("pois")
+rspois <- rshift("pois")
+```

### 2016-09-14 docs/prove-qdist+qshift.R

@@ -46,6 +46,7 @@ dsnorm <- dshift("norm")
 b <- dsnorm(x = 1:10, mean = 5, sd = 3, shift=2)
 
 dtsnorm <- dtruncate("snorm")
+ptsnorm <- ptruncate("snorm")
 c <- dtsnorm(x = 1:10, mean = 5, sd = 3, shift=2, L = 2, U = 8)
 
 df <- data.frame(a,b,c,d =1:10, stringsAsFactors = F)
@@ -70,6 +71,7 @@ b <- ptnorm(q = 1:10, mean = 5, sd = 3, L = 2, U = 8)
 ptsnorm <- pshift("tnorm")
 
 c <- ptsnorm(q = 1:10, mean = 5, sd = 3, shift=2, L = 2, U = 8)
+c <- qtsnorm(p = c(0.0000000, 0.0000000, 0.0000000, 0.0000000, 0.1374524, 0.3087584, 0.5000000, 0.6912416, 0.8625476, 1.0000000), mean = 5, sd = 3, shift=2, L = 2, U = 8)
 
 df <- data.frame(a,b,c,d =1:10, stringsAsFactors = F)
 
@@ -121,6 +123,29 @@ pl <- ggplot() +
   geom_line(data = df, mapping = aes(y=c, x=d), col="green")
 
 pl
+
+###############################
+# prove da p a q e viceversa  #
+###############################
+
+# tronco poi shifto
+a <- qtsnorm(p = (1:9)/10, mean = 5, sd = 3, shift=2, L = 2, U = 8)
+ptsnorm(q = a, mean = 5, sd = 3, shift=2, L = 2, U = 8)
+# giusto
+
+a <- ptsnorm(q = 1:10, mean = 5, sd = 3, shift=2, L = 2, U = 8)
+qtsnorm(p = a, mean = 5, sd = 3, shift=2, L = 2, U = 8)
+
+# shifto poi tronco
+c <- qtsnorm(p = 1:9/10, mean = 5, sd = 3, shift=2, L = 2, U = 8)
+ptsnorm(q = c, mean = 5, sd = 3, shift=2, L = 2, U = 8)
+
+
+c <- ptsnorm(q = 1:10, mean = 5, sd = 3, shift=2, L = 2, U = 8)
+qtsnorm(p = c, mean = 5, sd = 3, shift=2, L = 2, U = 8)
+
+
+# da q a p è sempre giusto, da p a q no
 
 #############################
[…]

### 2016-09-15 docs/article-shift+trunc.Rmd
articolo distr shiftate e troncate

@@
 R programming functional nature allows us programmer to implement and manipulate functions as we need.
-Functional programming is a style of programming that emphasizes the evaluation of expressions rather than the execution of commands.
-It has a great power and lots of properties that combined together can allows us to produce results with the most efficient computational method.
-In this article we will consider the functional factory method proposed by Andrea Spanò in "R package qdist: A functional programming approach to truncated probability ditributions", and that I used in "R package qshift: A functional programming approach to shifted probability distributions", that produces functions for truncated and shifted distributions, respectively.
+Functional programming is a style of programming that emphasizes the evaluation of expressions rather than the execution of commands, so these functions produce results that depend only on their inputs and not on the program state.
+This style has a great power and includes lots of tools that combined together can allows us to produce results with the most efficient computational method.
+One of them is the functional factory, which is a factory for making new functions.
+It allows us to use closures, which are functions together with a referencing environment, to write specific functions that, in turn, can be used to generate new functions.
+So we can have a double layer of development: a first layer that is used to do all the complex work in common to all functions and a second layer that defines the details of each function.
+This fundamentals was used to develop the computational method proposed by Andrea Spanò in "R package qdist: A functional programming approach to truncated probability ditributions", and that I used in "R package qshift: A functional programming approach to shifted probability distributions", that produces functions for truncated and shifted distributions, respectively.
 [`qdist`](https://github.com/andreaspano/qdist) and [`qshift`](https://github.com/veronicagiro/qshift) packages are available on Github:
@@
 
-We will demonstrate that it is possible to produce functions for truncated and shifted or for shifted and truncated distributions only combining the functions of the two packages.
+In the previous citated articles, we dealt separately shifted and truncated distributions.
+But in the real life, we sometimes work with random variables arising from more than one transformation.
+So in this article, we will demonstrate that it is possible to produce functions for truncated and shifted or for shifted and truncated distributions only combining the functions of the two packages thanks to functional programming approach used to built them.
+
+<!--
 
@@
 
-<!--
 pensare ad un modo diverso per spiegare la cosa perchè da questi grafici nn si capisce niente
 -->
-## Other Examples
 
-Let us see how it the other distribution functions works with other distributions.
+# Distribution functions for truncated and shifted or for shifted and truncated distributions
 
-Let us consider Poisson distribution and distribution function.
+R provides probability, density, quantile and random number generation functions in `stats` package.
+The functions, that we refer to, share a stable naming convention both for the names of the functions and in the first argument of these functions.
+In particular:
 
-First of all functions for truncated distribution and for shifted distributio have to be defined:
+* `p<dist>(q)` with `q` vector of quantiles, identifies probability distribution functions
+* `d<dist>(x)` with `x` vector of quantiles, identifies density functions
+* `q<dist>(p)` with `p` vector of probabilities, identifies quantile functions
+* `r<dist>(n)` with `n` number of observations, identifies random number generation
+
+where `<dist>` indicates the distribution family.
+
+As a result we have: `pnorm(q)`, `dnorm(x)`, `qnorm(p)` and `rnorm(n)` for normal distribution and similarly for all other distributions.
+
+Andrea Spanò proposed in `qdist` a functional programming method that computes truncated distributions functions for each distribution chosen without changing the well structured programming style and syntax of previous functions.
+
+The computational methodology proposed in `qdist` is based on two R key concepts:
+
+* the environment of a function can be used as a placeholder for other objects;
+* function formals can be manipulated.
+
+and develops itself in two steps:
+
+* defines the required functions suitable for computing from specific truncated probability distribution by using the functions available within its package.
+These functions take as input a probability distribution as a character string, in this case the normal distribution, and return the equivalent truncated distribution function as object:
+
+```{r, message=FALSE}
+require(qdist)
+dtnorm <- dtruncate("norm") # density function
+ptnorm <- ptruncate("norm") # probability function 
+qtnorm <- qtruncate("norm") # quantile function
+rtnorm <- rtruncate("norm") # random generation function
+```
+
+* uses the newly created functions (`tpnorm()`, `tdnorm()`, `tqnorm()`, `trnorm()`) for computing from truncated probability distributions.
+
+Let us consider `dtnorm()`.
+It has the same formals as the original `dnorm()`, plus two extra parameters: `L` and `U`, respectively for lower and upper truncation thresholds set by default to `-Inf` and `+Inf`:
 
 ```{r}
+args(dnorm)
+args(dtnorm)
+```
+
+As `dnorm()` is the function that computes the density function for a normal distribution, conventionally named `norm`, `dtnorm()` is the function that computes the density function for a truncated normal distributions, named `tnorm`.
+
+Suppose we want to generate the density functions for a **shifted truncated** normal distribution.
+We have to entrust to `qshift` package.
+`qshift` package shares the same logic and the same methodology of `qdist`.
+The difference is that the functions are designed to compute functions for shifted distributions.
+In particular, `qshift` package includes the following four functions:
+
+* `dshift()`, for density function
+* `pshift()`, for distribution function
+* `qshift()`, for quantile function
+* `rshift()`, for random generator function
+
+So for computing the density function of a shifted trucated normal distribution we have to pass `tnorm` distribution as `dshift()` argument:
+
+```{r, message=FALSE}
+require(qshift)
+dstnorm <- dshift("tnorm")
+```
+
+So `dstnorm()` is the function that computes the density function of `stnorm`, which is the shifted truncated normal distribution.
+
+```{r}
+args(dstnorm)
+```
+
+Looking at the function arguments we see that the sintax style is that of `dnorm()` function plus `L` and `U` arguments for truncated distribution and `shift` argument for shifted distribution.
+
+Let us graphically compare the results of `dnorm()`, `dtnorm()` and `dstnorm()` functions:
+
+```{r}
+x <- seq(6, 14, len = 100)
+std_density <- dnorm(x = x, mean = 10, sd = 1)
+trunc_density <- dtnorm(x = x, mean = 10, sd = 1, L = 9, U = 11)
+shift_trunc_density <- dstnorm(x = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
+```
+
+```{r}
+df <- data.frame(x=x, 
+                 type = factor(c(rep("std_density", times=length(x)),
+                    rep("trunc_density", times=length(x)),
+                    rep("shift_trunc_density", times=length(x))),
+                    levels = c("std_density", "trunc_density", "shift_trunc_density")), 
+                  value = c(std_density, trunc_density, shift_trunc_density))
+require(ggplot2)
+pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
+  geom_line(size=1) +
+  scale_color_manual(values = c("std_density" = "cyan3", 
+                                "trunc_density" = "#89FF99", 
+                                "shift_trunc_density" = "#FF89B4")) +
+  xlab("Value") + ylab("Density") +
+  ggtitle(paste("Density plot for non-transformed, truncated", "and shifted truncated Normal distributions", sep = "\n")) +
+  theme(legend.title=element_blank(), legend.text= element_text(size=10),
+  plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+```
+
+The inverse procedure generates the density function for a **truncated shifted** normal distribution.
+
+Starting from `dshift()` function of `qshift` package, we define `dsnorm()`, the density function of `snorm`, which is the shifted normal distribution:
+```{r}
+dsnorm <- dshift("norm")
+```
+
+We have to define also the others functions for shifted distribution, as they are implicitly necessary for the further step:
+
+```{r}
+psnorm <- pshift("norm")
+qsnorm <- qshift("norm")
+rsnorm <- rshift("norm")
+```
+
+The density function of a truncated shifted normal distribution is generated passing `snorm` distribution as `ptruncate()` argument:
+```{r}
+dtsnorm <- dtruncate("snorm")
+```
+
+So `dtsnorm()` is the function that computes the density function of `stnorm`, which is the truncated shifted normal distribution.
+
+Let us graphically represent the output of the three functions:
+
+```{r}
+std_density <- dnorm(x = x, mean = 10, sd = 1)
+shift_density <- dsnorm(x = x, mean = 10, sd = 1, shift = 1)
+trunc_shift_density <- dtsnorm(x = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
+```
+
+```{r}
+df <- data.frame(x=x, 
+                 type = factor(c(rep("std_density", times=length(x)),
+                    rep("shift_density", times=length(x)),
+                    rep("trunc_shift_density", times=length(x))),
+                    levels = c("std_density", "shift_density", "trunc_shift_density")), 
+                  value = c(std_density, shift_density, trunc_shift_density))
+
+pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
+  geom_line(size=1) +
+  scale_color_manual(values = c("std_density" = "cyan3", 
+                                "shift_density" = "#89FF99", 
+                                "trunc_shift_density" = "#FF89B4")) +
+  xlab("Value") + ylab("Density") +
+  ggtitle(paste("Density plot for non-transformed, shifted and","truncated shifted Normal distributions", sep = "\n")) +
+  theme(legend.title=element_blank(), legend.text= element_text(size=10),
+  plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+```
+
+Let us see graphically the differences between the density function of normal distribution, of a shifted truncated normal distribution and of a truncated shifted normal distribution.
+
+```{r}
+df <- data.frame(x=x, 
+                 type = factor(c(rep("std_density", times=length(x)),
+                    rep("trunc_shift_density", times=length(x)),
+                    rep("shift_trunc_density", times=length(x))),
+                    levels = c("std_density", "trunc_shift_density", "shift_trunc_density")), 
+                  value = c(std_density, trunc_shift_density, shift_trunc_density))
+
+pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
+  geom_line(size=1) +
+  scale_color_manual(values = c("std_density" = "cyan3", 
+                                "trunc_shift_density" = "#89FF99", 
+                                "shift_trunc_density" = "#FF89B4")) +
+  xlab("Value") + ylab("Density") +
+  ggtitle(paste("Density plot for non-transformed, truncated-shifted", "and shifted-truncated Normal distribution", sep="\n")) +
+  theme(legend.title=element_blank(), legend.text= element_text(size=10),
+  plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+```
+
+Looking at the figure we can see that a shifted truncated distribution is different from a truncated shifted distribution.
+
+Let us analyze the other distributions functions:
+
+* probability distribution function of truncated-shifted normal distribution and of shifted-truncated normal distribution:
+
+```{r}
+pstnorm <- pshift("tnorm")
+ptsnorm <- ptruncate("snorm")
+std_distr <- pnorm(q = x, mean = 10, sd = 1)
+st_distr <- pstnorm(q = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
+ts_distr <- ptsnorm(q = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
+```
+
+```{r}
+df <- data.frame(x=x, 
+                 type = factor(c(rep("std_distr", times=length(x)),
+                    rep("st_distr", times=length(x)),
+                    rep("ts_distr", times=length(x))),
+                    levels = c("std_distr", "st_distr", "ts_distr")), 
+                  value = c(std_distr, st_distr, ts_distr))
+
+pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
+  geom_line(size=1) +
+  scale_color_manual(values = c("std_distr" = "cyan3", 
+                                "st_distr" = "#89FF99", 
+                                "ts_distr" = "#FF89B4")) +
+  xlab("Value") + ylab("Density") +
+  ggtitle("Distribution plot for transformed and not-transformed Normal distributions") +
+  theme(legend.title=element_blank(), legend.text= element_text(size=10),
+  plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+```
+
+* quantile function of truncated-shifted normal distribution and of shifted-truncated normal distribution:
+
+```{r}
+qstnorm <- qshift("tnorm")
+qtsnorm <- qtruncate("snorm")
+x <- 1:999/1000
+std_quant <- qnorm(p = x, mean = 10, sd = 1)
+st_quant <- qstnorm(p = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
+ts_quant <- qtsnorm(p = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
+```
+
+```{r}
+df <- data.frame(x=x, 
+                 type = factor(c(rep("std_quant", times=length(x)),
+                    rep("st_quant", times=length(x)),
+                    rep("ts_quant", times=length(x))),
+                    levels = c("std_quant", "st_quant", "ts_quant")), 
+                  value = c(std_quant, st_quant, ts_quant))
+
+pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
+  geom_line(size=1) +
+  scale_color_manual(values = c("std_quant" = "cyan3", 
+                                "st_quant" = "#89FF99", 
+                                "ts_quant" = "#FF89B4")) +
+  xlab("Value") + ylab("Density") +
+  ggtitle("Distribution plot for transformed and not-transformed Normal distributions") +
+  theme(legend.title=element_blank(), legend.text= element_text(size=10),
+  plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+```
+
+* random generation function of truncated-shifted normal distribution and of shifted-truncated normal distribution:
+
+```{r}
+rstnorm <- rshift("tnorm")
+rtsnorm <- rtruncate("snorm")
+std_rand <- rnorm(n=1000, mean = 10, sd = 1)
+st_rand <- rstnorm(n=1000, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
+ts_rand <- rtsnorm(n=1000, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
+```
+
+```{r}
+df <- data.frame(x=1:1000, 
+                 type = factor(c(rep("std_rand", times=1000),
+                    rep("st_rand", times=1000),
+                    rep("ts_rand", times=1000)),
+                    levels = c("std_rand", "st_rand", "ts_rand")), 
+                  value = c(std_rand, st_rand, ts_rand))
+
+pl <- ggplot(data = df, mapping = aes(x=value, group=type, fill=type)) +
+        geom_density(alpha=0.6) +
+        scale_fill_manual(values = c("std_rand" = "cyan3",
+                              "st_rand" = "#89FF99",
+                              "ts_rand" = "#FF89B4")) +
+        ggtitle("Density plot for shifted and not-shifted Normal distributions") +
+        theme(legend.title=element_blank(), legend.text= element_text(size=10),
+        plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+```
+
+## Other examples
+
+Let us consider Poisson distribution and define the distribution function for shifted and truncated Poisson distribution.
+
+```{r}
+dspois <- dshift("pois")
+pspois <- pshift("pois")
+qspois <- qshift("pois")
+rspois <- rshift("pois")
+
 dtpois <- dtruncate("pois")
@@
 rtpois <- rtruncate("pois")
+```
 
-dspois <- dshift("pois")
-pspois <- pshift("pois")
-qspois <- qshift("pois")
-rspois <- rshift("pois")
+* Density function of shifted truncated and truncated shifted Poisson distribution.
+
+```{r}
+# truncated shifted
+dtspois <- dtruncate("spois")
+# shifted truncated 
+dstpois <- dshift("tpois")
 ```
+
+```{r}
+x <- 1:10
+std_density <- dpois(x = x, lambda = 3)
+shift_trunc_density <- dstpois(x = x, lambda = 3, shift = 1, L = 2, U = 7)
+trunc_shift_density <- dtspois(x = x, lambda = 3, L = 2, U = 7, shift = 1)
+```
+
+```{r}
+df <- data.frame(x=x, 
+                 type = factor(c(rep("std_density", times=length(x)),
+                    rep("shift_trunc_density", times=length(x)),
+                    rep("trunc_shift_density", times=length(x))),
+                    levels = c("std_density", "shift_trunc_density", "trunc_shift_density")), 
+                  value = c(std_density, shift_trunc_density, trunc_shift_density))
+
+pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
+  geom_line(size=1) +
+  scale_color_manual(values = c("std_density" = "cyan3", 
+                                "shift_trunc_density" = "#89FF99", 
+                                "trunc_shift_density" = "#FF89B4")) +
+  xlab("Value") + ylab("Density") +
+  ggtitle(paste("Density plot for non-transformed, shifted and","truncated shifted Normal distributions", sep = "\n")) +
+  theme(legend.title=element_blank(), legend.text= element_text(size=10),
+  plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+```
+
+* Probability function of shifted truncated and truncated shifted Poisson distribution.
+
+```{r}
+# truncated shifted
+ptspois <- ptruncate("spois")
+# shifted truncated 
+pstpois <- pshift("tpois")
+```
+
+```{r}
+std_distr <- ppois(q = x, lambda = 3)
+st_distr <- pstpois(q = x, lambda = 3, L = 2, U = 7, shift = 1)
+ts_distr <- ptspois(q = x, lambda = 3, L = 2, U = 7, shift = 1)
+```
+
+```{r}
+df <- data.frame(x=x, 
+                 type = factor(c(rep("std_distr", times=length(x)),
+                    rep("st_distr", times=length(x)),
+                    rep("ts_distr", times=length(x))),
+                    levels = c("std_distr", "st_distr", "ts_distr")), 
+                  value = c(std_distr, st_distr, ts_distr))
+
+pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
+  geom_line(size=1) +
+  scale_color_manual(values = c("std_distr" = "cyan3", 
+                                "st_distr" = "#89FF99", 
+                                "ts_distr" = "#FF89B4")) +
+  xlab("Value") + ylab("Density") +
+  ggtitle("Distribution plot for transformed and not-transformed Normal distributions") +
+  theme(legend.title=element_blank(), legend.text= element_text(size=10),
+  plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+```
+
+* Quantile function of shifted truncated and truncated shifted Poisson distribution.
+
+```{r}
+# truncated shifted
+qtspois <- qtruncate("spois")
+# shifted truncated 
+qstpois <- qshift("tpois")
+```
+
+```{r}
+x <- 1:999/1000
+std_distr <- qpois(p = x, lambda = 3)
+st_distr <- qstpois(p = x, lambda = 3, L = 2, U = 7, shift = 1)
+ts_distr <- qtspois(p = x, lambda = 3, L = 2, U = 7, shift = 1)
+```
+
+```{r}
+df <- data.frame(x=x, 
+                 type = factor(c(rep("std_distr", times=length(x)),
+                    rep("st_distr", times=length(x)),
+                    rep("ts_distr", times=length(x))),
+                    levels = c("std_distr", "st_distr", "ts_distr")), 
+                  value = c(std_distr, st_distr, ts_distr))
+
+pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
+  geom_line(size=1) +
+  scale_color_manual(values = c("std_distr" = "cyan3", 
+                                "st_distr" = "#89FF99", 
+                                "ts_distr" = "#FF89B4")) +
+  xlab("Value") + ylab("Density") +
+  ggtitle("Distribution plot for transformed and not-transformed Normal distributions") +
+  theme(legend.title=element_blank(), legend.text= element_text(size=10),
+  plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+```
+
+* random generation function of truncated-shifted normal distribution and of shifted-truncated poisson distribution:
+
+```{r}
+# truncated shifted
+rtspois <- rtruncate("spois")
+# shifted truncated 
+rstpois <- rshift("tpois")
+```
+
+```{r}
+std_rand <- rpois(n=1000, lambda = 3)
+st_rand <- rstpois(n=1000, lambda = 3, L = 2, U = 7, shift = 1)
+ts_rand <- rtspois(n=1000, lambda = 3, L = 2, U = 7, shift = 1)
+```
+
+```{r}
+df <- data.frame(x=1:1000, 
+                 type = factor(c(rep("std_rand", times=1000),
+                    rep("st_rand", times=1000),
+                    rep("ts_rand", times=1000)),
+                    levels = c("std_rand", "st_rand", "ts_rand")), 
+                  value = c(std_rand, st_rand, ts_rand))
+
+pl <- ggplot(data = df, mapping = aes(x=value, group=type, fill=type)) +
+        geom_density(alpha=0.6) +
+        scale_fill_manual(values = c("std_rand" = "cyan3",
+                              "st_rand" = "#89FF99",
+                              "ts_rand" = "#FF89B4")) +
+        ggtitle("Density plot for shifted and not-shifted Normal distributions") +
+        theme(legend.title=element_blank(), legend.text= element_text(size=10),
+        plot.title = element_text(size=14), axis.title=element_text(size=12))
+
+print(pl)
+```

### 2016-09-15 docs/article-shift+trunc.Rmd
articolo trunc+shift rivisto

@@
 So in this article, we will demonstrate that it is possible to produce functions for truncated and shifted or for shifted and truncated distributions only combining the functions of the two packages thanks to functional programming approach used to built them.
-
-<!--
-
-# Functional programming as a new approach for computing distribution functions
-
-Let us sum up the alternative functional programming approach used for developing `qdist` and `qshift`.
-
-R provides probability, density, quantile and random number generator functions in `stats` package, using a stable naming convention both for the names of the functions and in the first argument of these functions:
-
-* probability distribution functions p< dist >(q) with q vector of quantiles;
-* density functions d< dist >(x) with x vector of quantiles;
-* quantile functions q< dist >(p) with p vector of probabilities;
-* random number generation r< dist >(n) with n number of observations.
-
-where < dist > indicates the distribution family.
-
-As a result we have: pnorm(q), dnorm(x), qnorm(p) and rnorm(n) for normal distribution and similarly for all other distributions.
-
-Starting from this well structured syntax, Andrea Spanò proposed in `qdist` a functional programming method that computes truncated distributions functions for each distribution chosen without changing the programming style and syntax of previous functions.
-
-The methodology is based on two R key concepts:
-
-* the environment of a function can be used as a placeholder for other objects;
-* function formals can be manipulated.
-
-and develops itself in two steps:
-
-* defines the required functions suitable for computing from specific truncated probability distribution by using the functions available within its package.
-These functions take as input a probability distribution as a character string, in this case the normal distribution, and return the equivalent truncated distribution function as object:
-
-```{r, message=FALSE}
-require(qdist)
-dtnorm <- dtruncate("norm") # density function
-ptnorm <- ptruncate("norm") # probability function 
-qtnorm <- qtruncate("norm") # quantile function
-rtnorm <- rtruncate("norm") # random generation function
-```
-
-* uses the newly created functions (`tpnorm()`, `tdnorm()`, `tqnorm()`, `trnorm()`) for computing from truncated probability distributions.
-
-Let us consider `dtnorm()`.
-It has the same formals as the original `dnorm()`, plus two extra parameters: `L` and `U`, respectively for lower and upper truncation thresholds set by default to `-Inf` and `+Inf`:
-
-```{r}
-args(dnorm)
-args(dtnorm)
-```
-
-Let us see graphically the differences between `dnorm()` to compute the density function of a normal distribution with mean 10 and standard deviation 1 and `tdnorm()` to compute the density function of a normal distribution with mean 10 and standard deviation 1 lower truncated at 9 and upper truncated at 11:
-
-```{r, message=FALSE}
-x <- seq(6, 14, len = 100)
-std_density <- dnorm(x = x, mean = 10, sd = 1)
-trunc_density <- dtnorm(x = x, mean = 10, sd = 1, L = 9, U = 11)
-
-df <- data.frame(x=x, 
-                 type = factor(c(rep("std_density", times=length(x)),
-                    rep("trunc_density", times=length(x))),
-                    levels = c("std_density", "trunc_density")), 
-                  value = c(std_density, trunc_density))
-
-require(ggplot2)
-pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
-  geom_line(size=1) +
-  scale_color_manual(values = c("std_density" = "cyan3", 
-                                "trunc_density" = "#89FF99")) +
-  xlab("Value") + ylab("Density") +
-  ggtitle("Density plot for non-truncated and truncated Normal distributions") +
-  theme(legend.title=element_blank(), legend.text= element_text(size=10),
-  plot.title = element_text(size=14), axis.title=element_text(size=12))
-
-print(pl)
-```
-
-`qshift` package shares the same logic and the same methodology of `qdist`. The difference is that the functions are designed to compute functions for shifted distributions.
-
-The methodology is made up of the previously descripted two steps. So it:
-
-* defines the required functions suitable for computing from specific shifted probability models by using the `qshift` functions, that takes as input a probability distribution as a character string, in this case the normal distribution, and returns the equivalent shifted distribution function as object.
-
-```{r, message=FALSE}
-require(qshift)
-dsnorm <- dshift("norm") # density function
-psnorm <- pshift("norm") # probability function
-qsnorm <- qshift("norm") # quantile function
-rsnorm <- rshift("norm") # random generation function
-```
-
-* use the newly created functions (`dsnorm()`, `psnorm()`, `qsnorm()`, `rsnorm()`) for computing from shifted probability distributions.
-
-Let us consider `dsnorm()`. Also this function has the same formals as the original `dnorm()`, plus one extra parameter: `shift`, set by default to 0:
-
-```{r}
-args(dnorm)
-args(dsnorm)
-```
-
-Let us see graphically the differences between `dnorm()` to compute the density function of a normal distribution with mean 10 and standard deviation 1 and `sdnorm()` to compute the density function of a normal distribution with mean 10 and standard deviation 1 shifted of one unit:
-
-```{r}
-x <- seq(6, 14, len = 100)
-std_density <- dnorm(x = x, mean = 10, sd = 1)
-shift_density <- dsnorm(x = x, mean = 10, sd = 1, shift = 1)
-
-df <- data.frame(x=x, 
-                 type = factor(c(rep("std_density", times=length(x)),
-                    rep("shift_density", times=length(x))),
-                    levels = c("std_density", "shift_density")), 
-                  value = c(std_density, shift_density))
-
-pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
-  geom_line(size=1) +
-  scale_color_manual(values = c("std_density" = "cyan3", 
-                                "shift_density" = "#FF89B4")) +
-  xlab("Value") + ylab("Density") +
-  ggtitle("Density plot for non-shifted and shifted Normal distributions") +
-  theme(legend.title=element_blank(), legend.text= element_text(size=10),
-  plot.title = element_text(size=14), axis.title=element_text(size=12))
-
-print(pl)
-```
-
-For more details and examples about how `qdist` and `qshift` works, see articles "R package qdist: A functional programming approach to truncated probability ditributions" and "R package qshift: A functional programming approach to shifted probability distributions" available on [Quantide blog](http://www.quantide.com/blog/).
-
-# Distribution functions for truncated and shifted or for shifted and truncated distributions
-
-The functional programming approach whereby `qdist` and `qshift` packages are built is the key that allows to obtain distribution functions for truncated and shifted or for shifted and truncated distributions.
-Let us see how!
-
-Suppose we want to generate the density functions for a truncated shifted normal distribution.  
-As `dnorm()` is the function that computes the density function for a normal distributions, conventionally named `norm`, `dtnorm()`, defined as: 
-```{r}
-dtnorm <- dtruncate("norm")
-```
-
-is the function that computes the density function for a truncated normal distributions, named `tnorm`.
-
-We have to define also the others functions for truncated distribution, as they are necessary implicitly for the further step:
-
-```{r}
-ptnorm <- ptruncate("norm")
-qtnorm <- qtruncate("norm")
-rtnorm <- rtruncate("norm")
-```
-
-So for computing the density function of a shifted trucated normal distribution we have to pass `tnorm` distribution as `dshift()` argument:
-
-```{r}
-dstnorm <- dshift("tnorm")
-```
-
-So `dstnorm()` is the function that computes the density function of `stnorm`, which is the shifted truncated normal distribution.
-
-```{r}
-args(dstnorm)
-```
-
-Looking at the function arguments we see that the sintax style is that of `dnorm()` function plus `L` and `U` arguments for truncated distribution and `shift` argument for shifted distribution.
-
-The inverse procedure generates the density function for a shifted truncated distribution.
-
-Starting from `dshift`, we define `dsnorm()`, the density function of `snorm`, which is the shifted normal distribution:
-```{r}
-dsnorm <- dshift("norm")
-```
-
-We have to define also the others functions for shifted distribution, as they are implicitly necessary for the further step:
-
-```{r}
-psnorm <- pshift("norm")
-qsnorm <- qshift("norm")
-rsnorm <- rshift("norm")
-```
-
-The density function of a truncated shifted normal distribution is generated passing `snorm` distribution as `ptruncate()` argument:
-```{r}
-dtsnorm <- dtruncate("snorm")
-```
-
-So `dtsnorm()` is the function that computes the density function of `stnorm`, which is the truncated shifted normal distribution.
-
-Let us see graphically the differences between the density function of normal distribution, of a shifted truncated normal distribution and of a truncated shifted normal distribution.
-
-```{r}
-std_density <- dnorm(x = x, mean = 10, sd = 1)
-st_density <- dstnorm(x = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
-ts_density <- dtsnorm(x = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
-```
-
-```{r}
-df <- data.frame(x=x, 
-                 type = factor(c(rep("std_density", times=length(x)),
-                    rep("st_density", times=length(x)),
-                    rep("ts_density", times=length(x))),
-                    levels = c("std_density", "st_density", "ts_density")), 
-                  value = c(std_density, st_density, ts_density))
-
-pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
-  geom_line(size=1) +
-  scale_color_manual(values = c("std_density" = "cyan3", 
-                                "st_density" = "#89FF99", 
-                                "ts_density" = "#FF89B4")) +
-  xlab("Value") + ylab("Density") +
-  ggtitle("Density plot for transformed and not-transformed Normal distributions") +
-  theme(legend.title=element_blank(), legend.text= element_text(size=10),
-  plot.title = element_text(size=14), axis.title=element_text(size=12))
-
-print(pl)
-```
-
-We can conclude that the previous functions are completely different.
-
-Let us analyze the other distributions functions:
-
-* probability distribution function of truncated-shifted normal distribution and of shifted-truncated normal distribution:
-
-```{r}
-pstnorm <- pshift("tnorm")
-ptsnorm <- ptruncate("snorm")
-std_distr <- pnorm(q = x, mean = 10, sd = 1)
-st_distr <- pstnorm(q = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
-ts_distr <- ptsnorm(q = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
-```
-
-```{r}
-df <- data.frame(x=x, 
-                 type = factor(c(rep("std_distr", times=length(x)),
-                    rep("st_distr", times=length(x)),
-                    rep("ts_distr", times=length(x))),
-                    levels = c("std_distr", "st_distr", "ts_distr")), 
-                  value = c(std_distr, st_distr, ts_distr))
-
-pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
-  geom_line(size=1) +
-  scale_color_manual(values = c("std_distr" = "cyan3", 
-                                "st_distr" = "#89FF99", 
-                                "ts_distr" = "#FF89B4")) +
-  xlab("Value") + ylab("Density") +
-  ggtitle("Distribution plot for transformed and not-transformed Normal distributions") +
-  theme(legend.title=element_blank(), legend.text= element_text(size=10),
-  plot.title = element_text(size=14), axis.title=element_text(size=12))
-
-print(pl)
-```
-
-* quantile function of truncated-shifted normal distribution and of shifted-truncated normal distribution:
-
-```{r}
-qstnorm <- qshift("tnorm")
-qtsnorm <- qtruncate("snorm")
-x <- 1:999/1000
-std_quant <- qnorm(p = x, mean = 10, sd = 1)
-st_quant <- qstnorm(p = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
-ts_quant <- qtsnorm(p = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
-```
-
-```{r}
-df <- data.frame(x=x, 
-                 type = factor(c(rep("std_quant", times=length(x)),
-                    rep("st_quant", times=length(x)),
-                    rep("ts_quant", times=length(x))),
-                    levels = c("std_quant", "st_quant", "ts_quant")), 
-                  value = c(std_quant, st_quant, ts_quant))
-
-pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
-  geom_line(size=1) +
-  scale_color_manual(values = c("std_quant" = "cyan3", 
-                                "st_quant" = "#89FF99", 
-                                "ts_quant" = "#FF89B4")) +
-  xlab("Value") + ylab("Density") +
-  ggtitle("Distribution plot for transformed and not-transformed Normal distributions") +
-  theme(legend.title=element_blank(), legend.text= element_text(size=10),
-  plot.title = element_text(size=14), axis.title=element_text(size=12))
-
-print(pl)
-```
-
-* random generation function of truncated-shifted normal distribution and of shifted-truncated normal distribution:
-
-```{r}
-rstnorm <- rshift("tnorm")
-rtsnorm <- rtruncate("snorm")
-std_rand <- rnorm(n=1000, mean = 10, sd = 1)
-st_rand <- rstnorm(n=1000, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
-ts_rand <- rtsnorm(n=1000, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
-```
-
-```{r}
-df <- data.frame(x=1:1000, 
-                 type = factor(c(rep("std_rand", times=1000),
-                    rep("st_rand", times=1000),
-                    rep("ts_rand", times=1000)),
-                    levels = c("std_rand", "st_rand", "ts_rand")), 
-                  value = c(std_rand, st_rand, ts_rand))
-
-pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
-  geom_line(size=1) +
-  scale_color_manual(values = c("std_rand" = "cyan3", 
-                                "st_rand" = "#89FF99", 
-                                "ts_rand" = "#FF89B4")) +
-  xlab("Value") + ylab("Density") +
-  ggtitle("Distribution plot for transformed and not-transformed Normal distributions") +
-  theme(legend.title=element_blank(), legend.text= element_text(size=10),
-  plot.title = element_text(size=14), axis.title=element_text(size=12))
-
-print(pl)
-```
-
-pensare ad un modo diverso per spiegare la cosa perchè da questi grafici nn si capisce niente
--->
 

### 2016-09-15 docs/prove-qdist+qshift.R

@@ -1175,7 +1175,7 @@ ltweibull <- function(x, shift = 0 ,L = -Inf, U = Inf){
     shape <- theta[1]
     scale <- theta[2]
     #shift <- theta[3]
-    ld <- dstweibull(x = x, shape = shape, scale = scale, shift = shift, L = L, U = U, log = TRUE)
+    ld <- log(dstweibull(x = x, shape = shape, scale = scale, shift = shift, L = L, U = U))
     -sum(ld)
   }
   # maximum likelihood estimation
@@ -1186,18 +1186,30 @@ ltweibull <- function(x, shift = 0 ,L = -Inf, U = Inf){
 rtweibull <- rtruncate("weibull")
 rstweibull <- rshift("tweibull")
 x <- rstweibull(n = 100000, shape = 1, scale = 5, shift = 1, L = 1, U = 10000)
-x <- rstweibull(n = 100000, shape = 1, scale = 5, L = 1)
+x <- rstweibull(n = 100000, shape = 1, scale = 5, shift=2, L = 1, U = 100)
 # maximum likelihood estimate for the shifted weibull distribution
-ltweibull(x = x, L = 1)
+ltweibull(x = x, shift=2, L = 1, U = 100)
 
 
 
+trnorm <- rtruncate("norm")
+x <- trnorm(n = 1000, mean = 5, sd = 2, L = 3, U = 6)
 
+sd(x)
 
+ltnorm <- function(x, L = -Inf, U = Inf) {
+  theta <- c(5, 2)
+  ll <- function(theta, x, L = -Inf, U = Inf) {
+    mean <- theta[1]
+    sd <- theta[2]
+    ld <- log(tdnorm(x = x, mean = mean, sd = sd, L = L , U = U))
+    -sum(ld)
+  }
+  optim(par = theta, fn = ll, x = x, L = L, U = U, method = "Nelder-Mead")[["par"]]
+}
 
-
-
-
+ltnorm( x = x, L = 3, U = 6)
+tdnorm <- dtruncate("norm")
 
 
 

### 2016-09-15 docs/prove-qdist+qshift.R
articolo distr tronc+shift rivisto

@@ -1138,7 +1138,7 @@ ltweibull <- function(x, L = -Inf, U = Inf){
     shape <- theta[1]
     scale <- theta[2]
     shift <- theta[3]
-    ld <- dtsweibull(x = x, shape = shape, scale = scale, shift = shift, L = L, U = U, log = TRUE)
+    ld <- log(dtsweibull(x = x, shape = shape, scale = scale, shift = shift, L = L, U = U))
     -sum(ld)
   }
   # maximum likelihood estimation
@@ -1148,7 +1148,7 @@ ltweibull <- function(x, L = -Inf, U = Inf){
 
 rsweibull <- rshift("weibull")
 rtsweibull <- rtruncate("sweibull")
-x <- rtsweibull(n = 100000, shape = 1, scale = 2, shift = 1, L = 5, U = 10)
+x <- rtsweibull(n = 100000, shape = 1, scale = 1.5, shift = 1, L = 5, U = 10)
 
 # maximum likelihood estimate for the shifted weibull distribution
 ltweibull(x = x, L = 5, U = 10)

### 2016-09-19 docs/qdist-article/truncdist.Rmd
riorganizzazione cartella docs

---
title: "R package qdist: A functional programming approach to truncated probability ditributions"
output:
  html_document:
    theme: null
    highlight: null
---

## Introduction

Truncated probability distribution are widely used in many applied areas of statistics. Loss severity
models in the finance or insurance industries are good cases in point. Package `stats` provides a wide
set of functions for distribution models. Other packages, such as package `evd`, provide extra distributional models.
Truncated distribution models can be found in package `truncdist` ([Novomestky and Nadarajah, 2012](#[p1-2]) ; [Nadarajah and Kotz, 2006](#[p1-1])). Package `truncdist` provides an efficient method for computing from
truncated distribution but requires a change in our programming style when compared with the
approach we use for non-truncated models. This article aims to provide an alternative method for
computing from truncated probability distribution. This method is made of two steps:

* define the required functions suitable for computing from specific truncated probability models
by using the tools available within this package;
* use the newly created functions for computing from truncated probability models.

## Truncated probability function in R

R provides probability, density, quantile and random number generator functions using a stable
naming convention both for the names of the functions and in the first argument of these functions:

* probability distribution functions `p<dist>(q)` with `q` vector of quantiles;
* density functions `d<dist>(x)` with `x` vector of quantiles;
* quantile functions `q<dist>(p)` with `p` vector of probabilities;
* random number generation `r<dist>(n)` with `n` number of observations.

where `<dist>` indicates the distribution family.

As a result we have: `pnorm(q)`, `dnorm(x)`, `qnorm(p)` and `rnorm(n)` for normal distribution and
similarly for all other distributions.

Therefore, we use to write:

```{r}
pnorm(q = 8:12, mean = 10, sd = 1)
```

to get probability values at `8:12` from a normal distribution with parameters `mean=10` and `sd=1`.

In case we need values from a truncated distribution, as far as we know, we need to load an extra
package such as `truncdist`.

```{r, message=FALSE}
require("truncdist")
```

The package itself works perfectly. In fact, assuming that function `pnorm()` exists, we can get
probability values from a normal distribution left truncated at `a` and right truncated at `b`, with
parameters `mean = 10` and `sd = 1` by simply writing:

```{r}
ptrunc(q = 8:12, spec = "norm", mean = 10, sd = 1, a = 9, b = 11)
```

where `a = 9` and `b = 11` respectively represent the left and the right thresholds for truncation.

Nevertheless, the above command requires a change in our programming style. We are used to
have a single R function for each distribution: `pnorm()`, `pweibull()` etc. . .

We could easily write function `tpnorm()`, a truncated version for `pnorm()`, as:

```{r}
tpnorm <- function (q, mean = 0, sd = 1, L = -Inf, U = Inf, ...){
  q <- pmax(pmin(q,U),L)
  pq <- stats::pnorm(q = q, mean = mean, sd = sd, ...)
  pL <- stats::pnorm(q = L, mean = mean, sd = sd, ...)
  pU <- stats::pnorm(q = U, mean = mean, sd = sd, ...)
  p <- (pq-pL)/(pU-pL)
  p
}
```

This function clearly works:

```{r}
tpnorm(q = 8:12, mean = 10, sd = 1, L = 9, U = 11)
```

but it is limited to the normal distribution. Following this approach, we would need to write
a different function for each probability distribution and, we’ll have to admit that all of this could
become quite time consuming and not really efficient.

As an alternative approach, we decided to develop package `qdist` available at [https://github.com/andreaspano/qdist](https://github.com/andreaspano/qdist) .

This package is made of four functions:

* `ptruncate()`: truncated probability distributions;
* `dtruncate()`: truncated density functions;
* `qtruncate()`: truncated quantile functions;
* `rtruncate()`: truncated random numbers generator functions.

These functions share the same logic: they take as input a probability distribution as a character
string and return the equivalent truncated distribution as a function object so that we can write:

```{r, message=FALSE}
devtools::install_github("andreaspano/qdist", build=FALSE)
require(qdist)
tpnorm <- ptruncate("norm")
tdnorm <- dtruncate("norm")
tqnorm <- qtruncate("norm")
trnorm <- rtruncate("norm")
```

As a first detailed example, let’s consider `ptruncate()` in more details applied to the normal case.

The newly generated function `tpnorm()` has the same formals as the original `pnorm()`, plus two
extra parameters: `L` and `U`, respectively for lower and upper truncation thresholds set by default to
`-Inf` and `+Inf`:

```{r}
args(pnorm)
args(tpnorm)
```

Once we have defined `tpnorm()` we can use it for generating probability values from a non-
truncated normal distribution by leaving parameters `L` and `U` set to their defaults:

```{r}
q <- seq(6, 14, len = 100)
p_L_U <- tpnorm(q, mean = 10, sd = 1)
```

or from any truncated normal distribution by setting values for `L` and `U`:

```{r}
p_L9_U11 <- tpnorm(q, mean = 10, sd = 1, L = 9, U = 11)
```

We can visualize these results by plotting them:

```{r, fig.width=8, warning=FALSE, message=FALSE}
df <- data.frame(
  q=q,
  tpnorm_type = factor(c(rep("p_L_U", times=length(q)), rep("p_L9_U11", times=length(q))), levels = c("p_L_U", "p_L9_U11")),
  value = c(p_L_U, p_L9_U11)
)

require(ggplot2)

pl <- ggplot(data = df, mapping = aes(x=q, y=value, col=tpnorm_type)) +
  geom_line(size=1) + scale_color_manual(values = c("p_L_U" = "darkgray", "p_L9_U11" ="darkred")) +   
  xlab("Quantile") + ylab("Probability") + 
  ggtitle("Probability plot for truncated and non-truncated normal distribution") +
  theme(legend.title=element_blank(), legend.text= element_text(size=12), plot.title = element_text(size=17), axis.title=element_text(size=14))

print(pl)
```

As a second example, we consider a density function for a left truncated Weibull distribution. We
first generate function `tdweibull()` by using `dtruncate()`:

```{r}
tdweibull <- dtruncate("weibull")
```

and afterward, we can use it as:

```{r}
x <- qweibull(ppoints(1000), shape = 2, scale = 7)
d_L0 <- tdweibull(x, shape = 2, scale = 7)
d_L3 <- tdweibull(x, shape = 2, scale = 7, L= 3)
d_L5 <- tdweibull(x, shape = 2, scale = 7, L= 5)
d_L7 <- tdweibull(x, shape = 2, scale = 7, L= 7)
```

with the following results:

```{r, warning=FALSE, message=FALSE}
df <- data.frame(
  x=x,
  tdweibull_type = factor(c(rep("d_L0", times=length(x)), rep("d_L3", times=length(x)), rep("d_L5", times=length(x)), rep("d_L7", times=length(x))), levels = c("d_L0", "d_L3", "d_L5", "d_L7")),
  value = c(d_L0, d_L3, d_L5, d_L7)
)

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=tdweibull_type)) +
  geom_line(size=1) + 
  scale_color_manual(values = c("d_L0" = "darkgray", "d_L3"="darkred", "d_L5" = "darkblue", "d_L7"="darkgreen")) +   
  xlab("Quantile") + ylab("Density") + ggtitle("Density plot for truncated Weibull distributions") +
  theme(legend.title=element_blank(), legend.text= element_text(size=12), plot.title = element_text(size=18), axis.title=element_text(size=14))
print(pl)
```

As an example of random number generator function we consider the Gumbel distribution from package `evd`:

```{r, warning=FALSE, message=FALSE}
require(evd)
trgumbel <- rtruncate("gumbel")
args(trgumbel)
rg <- trgumbel(10^4, loc = 3, scale = 2, L = 5)
df <- data.frame(rg=rg)

pl <- ggplot(data = df, mapping = aes(x=rg)) +
  geom_histogram(fill="grey", color="black", breaks=0:25) + xlim(c(0,25)) +
  ggtitle("Random data from truncated Gumbel distribution") +
  theme(plot.title = element_text(size=18), axis.title=element_text(size=14))

print(pl)
```

This approach works for both continuous and discrete probability functions. Let’s consider the
case of main quantiles generated from a Poisson distribution. We first generate the truncated quantile
function:

```{r}
tqpois <- qtruncate("pois")
```

We use the newly created function for computing percentiles from four Poisson distributions
truncated at different tresholds:

```{r}
p <- ppoints(100)
# No truncation
q <- tqpois(p, lambda = 10)
# Lower truncation L = 9
q_L9 <- tqpois(p, lambda = 10, L = 9)
# Upper truncation H = 14
q_U14 <- tqpois(p, lambda = 10, U = 14)
# Both sides truncation L = 9, H = 14
q_L9_U14 <- tqpois(p, lambda = 10, L = 9, U = 14)
```

Finally, we plot the computed percentiles:

```{r}
df <- data.frame(p=p,
  tqpois_type=factor(c(rep("No truncation", times=length(p)), rep("Lower truncation L = 9", times=length(p)), 
    rep("Upper truncation H = 14", times=length(p)), rep("Both sides truncation L = 9, H = 14", times=length(p))), 
    levels= c("No truncation", "Lower truncation L = 9", "Upper truncation H = 14", "Both sides truncation L = 9, H = 14")),
  value=c(q, q_L9, q_U14, q_L9_U14))

pl <- ggplot(data = df, mapping = aes(x=value, y=p, col=tqpois_type)) +
  geom_line(size=1, show.legend = FALSE) + 
  ylab("Probability") + xlab("Quantiles") + ggtitle("Quantiles from truncated Poisson distribution") +
  facet_wrap(~tqpois_type, nrow = 2) +
  theme(strip.text = element_text(size=12), plot.title = element_text(size=18), axis.title=element_text(size=14))

print(pl)  
```

## The `gamma` case

The gamma case may result a bit faulty because of the check on input parameters implemented within
the body of the gamma set of functions. As a consequence, suppose we define a truncated quantile of the
gamma distribution as:

```{r}
tqgamma <- qtruncate("gamma")
```

When we use `tqgamma()` with parameter rate it works but may return a set of `warnings`:

```{r}
tqgamma(.25, shape = 1, rate = .3)
```

when we use the same function with parameter `scale`, it returns an error:

```{r, error=TRUE}
tqgamma(ppoints(10), shape = 1, scale = 3)
```

Note that, if we simply redefine `qgamma()` and `pgamma()` as:

```{r, eval=FALSE, echo=FALSE}
# Andrea
qgamma <- function (p, shape, rate = 1, lower.tail = TRUE, log.p = FALSE) {
  scale <- 1/rate
  .External(stats:::C_qgamma, p, shape, scale, lower.tail, log.p)
}

pgamma <- function (q, shape, rate = 1, lower.tail = TRUE, log.p = FALSE) {
  scale <- 1/rate
  .External(stats:::C_pgamma, q, shape, scale, lower.tail, log.p)
}
```

```{r}
qgamma <- function (p, shape, rate = 1, lower.tail = TRUE, log.p = FALSE) {
  scale <- 1/rate
  .Call(stats:::C_qgamma, p, shape, scale, lower.tail, log.p)
}

pgamma <- function (q, shape, rate = 1, lower.tail = TRUE, log.p = FALSE) {
  scale <- 1/rate
  .Call(stats:::C_pgamma, q, shape, scale, lower.tail, log.p)
}

```

now everything should work fine:

```{r, error=TRUE}
tqgamma <- qtruncate("gamma")
tqgamma(p = .25, shape = 1, rate = 3)
```

## Extending the computation

Once a truncated distribution is defined, we can use it as a building block for further implementations.
Suppose we want to define a function for maximum likelihood estimate for the truncated normal
distribution, we can achieve this by first defining the truncated density function for the normal
distribution:

```{r}
tdnorm <- dtruncate("norm")
```

and, subsequently, by using `tdnorm()` within an estimator function:

```{r}
ltnorm <- function(x, L = -Inf, U = Inf) {
  theta <- c(mean(x), sd(x))
  ll <- function(theta, x, L = -Inf, U = Inf) {
    mean <- theta[1]
    sd <- theta[2]
    ld <- tdnorm(x = x, mean = mean, sd = sd, L = L , U = U, log = TRUE)
    -sum(ld)
  }
  optim(par = theta, fn = ll, x = x, L = L, U = U)[["par"]]
}
```

As a result:

```{r}
trnorm <- rtruncate("norm")
x <- trnorm(n = 1000, mean = 5, sd = 2, L = 3, U = 6)
ltnorm( x = x, L = 3, U = 6)
```

## Computational details

Package `qdist` can be thought as a function factory ([Wickham](#[p7])) for truncated distribution functions.

Function `ptruncate()`, for example, takes as input the distribution name: say _norm_ and proceeds as
follows:

* gets the corresponding function `pnorm()`;
* uses `pnorm()` to create a function, say `probability()` that computes truncated probability for normal distributions;
* modify the formals of `probability()` so that it has the same formals as `pnorm()` plus `L` and `U` corresponding to the lower and upper threshold for truncation;
* returns `probability()`.

Package `qdist` is based on two key concepts being part of the functional nature of the R programming language:

* the environment of a function can be used as a placeholder for other objects;
* function formals can be manipulated.

### Environment of a function

The _R Language Definition_ manual defines environments as: _“consisting of two things. A frame, consisting of a set of symbol-value pairs, and an enclosure, a pointer to an enclosing environment”_

This idea of a pointer to an enclosing environment is at the core of the R mechanism when looking for objects:

_“When R looks up the value for a symbol the frame is examined and if a matching symbol is found its value will be returned. If not, the enclosing environment is then accessed and the process repeated”._

Environments in R play a crucial role as they just work in the background of any R functionality.
Any R session has an associated environment as returned by:

```{r}
environment()
```

and, very important for our purposes, any function has an associated environment as stated by the
_R Language Definition_ manual: _“functions have three basic components: a formal argument list, a body and an environment”_.

Specifically, the _environment of a function_ is the environment that was active at the time that the
function was created. Generally, for user defined function, the _Global environment_:

```{r}
f <- function() 0
environment(f)
```

or, when a function is defined within a package, the environment associated to that package:

```{r}
environment(mean)
```

Along with the environment where the function was created, functions interact with several other
environments. The _evaluation environment_ of a function is one of them.

The _evaluation environment_ of a function is created any time a function is called and is used to host the computation of the function.

The evaluation environment is destroyed when the function exits.

As any environment, the evaluation environment of a function has a parent: the environment of the function.

When we define a function, the function itself knows about its environment and, as a consequence,
the function has access to all symbols belonging to that environment.

As an example we may consider a function defined in a dedicated environment along with some
other objects defined in the same environment.

```{r}
env <- new.env()
with(env,{y <- 99; g <- function(x){x+y}})
with(env, g(1))
```

As we can see, clearly `g()` knows that `x=1` as it was passed to the function as an argument but, `g()`
also remembers that `y=99` as y belongs the the environment `env`: the environment of `g()`.

Playing with the environment of a function and the execution environment of a function, we can
create a function `g()` that returns a function `f()`:

```{r}
g <- function() {
  f <- function() 0
  f
}
```

create `f()` as a result of a call to `g()`

```{r}
f <- g()
```

and normally run `f()` as:

```{r}
f()
```

in this case, the _evaluation environment_ of `g()` corresponds to the _environment_ of `f()` as `f()` is created
within the _evaluation environment_ of `g()`.

As a result, when `g()` exits, its _evaluation environment_ is not destroyed as it became the _environment_ of `f()`.

Following this line, we can define `g(y)` so that the _evaluation environment_ of `g(y)` is used to pass any argument `y` to `f(x)`:

```{r}
g <- function(y) {
  f <- function(x) {x+y}
}
```

We can now define and run `f(x)` as:

```{r}
f <- g(1)
f(x = 2)
```

This mechanism allows us to define a _functions factory_: a function `g(y)` that, by varying the values assigned to `y`, allows many `fi(x)` to be defined with very little effort:

```{r}
f1 <- g(1)
f2 <- g(2)
f3 <- g(3)
```

and use them straighfortly:

```{r}
f1(x = 100)
f2(x = 100)
f3(x = 100)
```

### Formals argument list

The argument list of a function, as stated in _R Language Definition_ manual is: _a comma-separated list of arguments. An argument can be a symbol, or a symbol = default construct_.

Function `formals()` returns the formal arguments of a function as an object of class `pairlist`.

```{r}
formals_sd <- formals(sd)
formals_sd
class(formals_sd)
```

As a replacement method exists for function formals :

```{r}
exists("formals<-")
```

formals of a function can manipulated by using function `alist()`: a `list()` type function that
handles unevaluated arguments

```{r}
f <- function(x, y=0) x+y
f(1)
```

```{r}
formals(f) <- alist(x=, y=1)
f(1)
```

As an example of practical use of `formals()` we may decide to re-define function `mean()` that
defaults `na.rm` to `TRUE` by simply:

```{r}
formals(mean.default)[["na.rm"]] <- TRUE
mean(c(1,2,NA))
```

## Bibliography

Nadarajah S. and Kotz S. . R programs for truncated distributions. _Journal of Statistical Software, Code Snippets_, 16(2):1–8, 8 2006. ISSN 1548-7660. URL [http://www.jstatsoft.org/v16/c02](http://www.jstatsoft.org/v16/c02). [p1] <a name="[p1-1]"></a>

Novomestky F. and Nadarajah S. . _truncdist: Truncated Random Variables_, 2012. URL [http://CRAN.R-project.org/package=truncdist](http://CRAN.R-project.org/package=truncdist) . R package version 1.0-1. [p1] <a name="[p1-2]"></a>

Wickham H. . Advanced r. [http://adv-r.had.co.nz/](http://adv-r.had.co.nz/). Accessed: 2014-07-31. [p7]

<a name="[p7]"></a>

### 2016-09-19 docs/qshift-article/article-shift+trunc.Rmd
riorganizzazione cartella docs

--- 2016-09-15 docs/article-shift+trunc.Rmd
@@
 
-R programming functional nature allows us programmer to implement and manipulate functions as we need.
+R programming functional nature allows programmer to implement and manipulate functions as we need.
 Functional programming is a style of programming that emphasizes the evaluation of expressions rather than the execution of commands, so these functions produce results that depend only on their inputs and not on the program state.
@@
 
-In the previous citated articles, we dealt separately shifted and truncated distributions.
+In the previous citated articles, we dealt with separately shifted and truncated distributions.
 But in the real life, we sometimes work with random variables arising from more than one transformation.
@@
 R provides probability, density, quantile and random number generation functions in `stats` package.
-The functions, that we refer to, share a stable naming convention both for the names of the functions and in the first argument of these functions.
+These functions share a stable naming convention both for the names of the functions and in the first argument of these functions.
 In particular:
@@
 
-As a result we have: `pnorm(q)`, `dnorm(x)`, `qnorm(p)` and `rnorm(n)` for normal distribution and similarly for all other distributions.
+As a result we have: `pweibull(q)`, `dweibull(x)`, `qweibull(p)` and `rweibull(n)` for Weibull distribution and similarly for all other distributions.
 
@@
 
-* the environment of a function can be used as a placeholder for other objects;
-* function formals can be manipulated.
+* the environment of a function can be used as a placeholder for other objects (for detail see Andrea Spanò article "R package qdist: A functional programming approach to truncated probability ditributions"")
+* function formals can be manipulated
 
-and develops itself in two steps:
+and it is made of two steps:
 
-* defines the required functions suitable for computing from specific truncated probability distribution by using the functions available within its package.
-These functions take as input a probability distribution as a character string, in this case the normal distribution, and return the equivalent truncated distribution function as object:
+1.
+defines the required functions suitable for computing from specific truncated probability distribution by using the functions available within its package.
+These functions take as input a probability distribution as a character string, in this case the Weibull distribution, and return the equivalent truncated distribution function as object:
 
@@
 require(qdist)
-dtnorm <- dtruncate("norm") # density function
-ptnorm <- ptruncate("norm") # probability function 
-qtnorm <- qtruncate("norm") # quantile function
-rtnorm <- rtruncate("norm") # random generation function
+dtweibull <- dtruncate("weibull") # density function
+ptweibull <- ptruncate("weibull") # probability function 
+qtweibull <- qtruncate("weibull") # quantile function
+rtweibull <- rtruncate("weibull") # random generation function
 ```
 
-* uses the newly created functions (`tpnorm()`, `tdnorm()`, `tqnorm()`, `trnorm()`) for computing from truncated probability distributions.
+2. uses the newly created functions (`tpweibull()`, `tdweibull()`, `tqweibull()`, `trweibull()`) for computing from truncated probability distributions.
 
-Let us consider `dtnorm()`.
-It has the same formals as the original `dnorm()`, plus two extra parameters: `L` and `U`, respectively for lower and upper truncation thresholds set by default to `-Inf` and `+Inf`:
+Let us consider `dtweibull()`.
+It has the same formals as the original `dweibull()`, plus two extra parameters: `L` and `U`, respectively for lower and upper truncation thresholds set by default to `-Inf` and `+Inf`:
 
 ```{r}
-args(dnorm)
-args(dtnorm)
+args(dweibull)
+args(dtweibull)
 ```
 
-As `dnorm()` is the function that computes the density function for a normal distribution, conventionally named `norm`, `dtnorm()` is the function that computes the density function for a truncated normal distributions, named `tnorm`.
+As `dweibull()` is the function that computes the density function for a Weibull distribution, conventionally named `weibull`, `dtweibull()` is the function that computes the density function for a truncated Weibull distributions, named `tweibull`.
 
-Suppose we want to generate the density functions for a **shifted truncated** normal distribution.
+Suppose we want to generate the density function for a **shifted truncated** Weibull distribution.
 We have to entrust to `qshift` package.
@@
 
-So for computing the density function of a shifted trucated normal distribution we have to pass `tnorm` distribution as `dshift()` argument:
+So for computing the density function of a shifted trucated Weibull distribution we have to pass `tweibull` distribution as `dshift()` argument:
 
@@
 require(qshift)
-dstnorm <- dshift("tnorm")
+dstweibull <- dshift("tweibull")
 ```
 
-So `dstnorm()` is the function that computes the density function of `stnorm`, which is the shifted truncated normal distribution.
+So `dstweibull()` is the function that computes the density function of `stweibull`, which is the shifted truncated Weibull distribution.
 
 ```{r}
-args(dstnorm)
+args(dstweibull)
 ```
 
-Looking at the function arguments we see that the sintax style is that of `dnorm()` function plus `L` and `U` arguments for truncated distribution and `shift` argument for shifted distribution.
+Looking at the function arguments we see that the sintax style is that of `dweibull()` function plus `L` and `U` arguments for truncated distribution and `shift` argument for shifted distribution.
 
-Let us graphically compare the results of `dnorm()`, `dtnorm()` and `dstnorm()` functions:
+Let us graphically compare the results of `dweibull()`, `dtweibull()` and `dstweibull()` functions:
 
 ```{r}
-x <- seq(6, 14, len = 100)
-std_density <- dnorm(x = x, mean = 10, sd = 1)
-trunc_density <- dtnorm(x = x, mean = 10, sd = 1, L = 9, U = 11)
-shift_trunc_density <- dstnorm(x = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
+x <- seq(0, 6, len = 100)
+orig_density <- dweibull(x = x, shape = 2, scale = 1.5)
+trunc_density <- dtweibull(x = x, shape = 2, scale = 1.5, L = 1, U = 4)
+shift_trunc_density <- dstweibull(x = x, shape = 2, scale = 1.5, L = 1, U = 4, shift = 1)
 ```
@@
 df <- data.frame(x=x, 
-                 type = factor(c(rep("std_density", times=length(x)),
+                 type = factor(c(rep("orig_density", times=length(x)),
                     rep("trunc_density", times=length(x)),
                     rep("shift_trunc_density", times=length(x))),
-                    levels = c("std_density", "trunc_density", "shift_trunc_density")), 
-                  value = c(std_density, trunc_density, shift_trunc_density))
+                    levels = c("orig_density", "trunc_density", "shift_trunc_density")), 
+                  value = c(orig_density, trunc_density, shift_trunc_density))
 require(ggplot2)
@@
   geom_line(size=1) +
-  scale_color_manual(values = c("std_density" = "cyan3", 
+  scale_color_manual(values = c("orig_density" = "cyan3", 
                                 "trunc_density" = "#89FF99", 
@@
 
-The inverse procedure generates the density function for a **truncated shifted** normal distribution.
+The inverse procedure generates the density function for a **truncated shifted** weibull distribution.
 
-Starting from `dshift()` function of `qshift` package, we define `dsnorm()`, the density function of `snorm`, which is the shifted normal distribution:
+Starting from `dshift()` function of `qshift` package, we define `dsweibull()`, the density function of `sweibull`, which is the shifted weibull distribution:
 ```{r}
-dsnorm <- dshift("norm")
+dsweibull <- dshift("weibull")
 ```
@@
 ```{r}
-psnorm <- pshift("norm")
-qsnorm <- qshift("norm")
-rsnorm <- rshift("norm")
+psweibull <- pshift("weibull")
+qsweibull <- qshift("weibull")
+rsweibull <- rshift("weibull")
 ```
 
-The density function of a truncated shifted normal distribution is generated passing `snorm` distribution as `ptruncate()` argument:
+The density function of a truncated shifted Weibull distribution is generated passing `sweibull` distribution as `ptruncate()` argument:
 ```{r}
-dtsnorm <- dtruncate("snorm")
+dtsweibull <- dtruncate("sweibull")
 ```
 
-So `dtsnorm()` is the function that computes the density function of `stnorm`, which is the truncated shifted normal distribution.
+So `dtsweibull()` is the function that computes the density function of `stweibull`, which is the truncated shifted Weibull distribution.
 
@@
 ```{r}
-std_density <- dnorm(x = x, mean = 10, sd = 1)
-shift_density <- dsnorm(x = x, mean = 10, sd = 1, shift = 1)
-trunc_shift_density <- dtsnorm(x = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
+orig_density <- dweibull(x = x, shape = 3, scale = 1)
+shift_density <- dsweibull(x = x, shape = 3, scale = 1, shift = 1)
+trunc_shift_density <- dtsweibull(x = x, shape = 3, scale = 1, L = 3, U = 11, shift = 1)
 ```
@@
 df <- data.frame(x=x, 
-                 type = factor(c(rep("std_density", times=length(x)),
+                 type = factor(c(rep("orig_density", times=length(x)),
                     rep("shift_density", times=length(x)),
                     rep("trunc_shift_density", times=length(x))),
-                    levels = c("std_density", "shift_density", "trunc_shift_density")), 
-                  value = c(std_density, shift_density, trunc_shift_density))
+                    levels = c("orig_density", "shift_density", "trunc_shift_density")), 
+                  value = c(orig_density, shift_density, trunc_shift_density))
 
@@
   geom_line(size=1) +
-  scale_color_manual(values = c("std_density" = "cyan3", 
+  scale_color_manual(values = c("orig_density" = "cyan3", 
                                 "shift_density" = "#89FF99", 
@@
 df <- data.frame(x=x, 
-                 type = factor(c(rep("std_density", times=length(x)),
+                 type = factor(c(rep("orig_density", times=length(x)),
                     rep("trunc_shift_density", times=length(x)),
                     rep("shift_trunc_density", times=length(x))),
-                    levels = c("std_density", "trunc_shift_density", "shift_trunc_density")), 
-                  value = c(std_density, trunc_shift_density, shift_trunc_density))
+                    levels = c("orig_density", "trunc_shift_density", "shift_trunc_density")), 
+                  value = c(orig_density, trunc_shift_density, shift_trunc_density))
 
@@
   geom_line(size=1) +
-  scale_color_manual(values = c("std_density" = "cyan3", 
+  scale_color_manual(values = c("orig_density" = "cyan3", 
                                 "trunc_shift_density" = "#89FF99", 
@@
 
-* probability distribution function of truncated-shifted normal distribution and of shifted-truncated normal distribution:
+* probability distribution function of truncated-shifted Weibull distribution and of shifted-truncated Weibull distribution:
 
 ```{r}
-pstnorm <- pshift("tnorm")
-ptsnorm <- ptruncate("snorm")
-std_distr <- pnorm(q = x, mean = 10, sd = 1)
-st_distr <- pstnorm(q = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
-ts_distr <- ptsnorm(q = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
+pstweibull <- pshift("tweibull")
+ptsweibull <- ptruncate("sweibull")
+orig_distr <- pweibull(q = x, shape = 3, scale = 1)
+st_distr <- pstweibull(q = x, shape = 3, scale = 1, L = 3, U = 11, shift = 1)
+ts_distr <- ptsweibull(q = x, shape = 3, scale = 1, L = 3, U = 11, shift = 1)
 ```
@@
 df <- data.frame(x=x, 
-                 type = factor(c(rep("std_distr", times=length(x)),
+                 type = factor(c(rep("orig_distr", times=length(x)),
                     rep("st_distr", times=length(x)),
                     rep("ts_distr", times=length(x))),
-                    levels = c("std_distr", "st_distr", "ts_distr")), 
-                  value = c(std_distr, st_distr, ts_distr))
+                    levels = c("orig_distr", "st_distr", "ts_distr")), 
+                  value = c(orig_distr, st_distr, ts_distr))
 
@@
   geom_line(size=1) +
-  scale_color_manual(values = c("std_distr" = "cyan3", 
+  scale_color_manual(values = c("orig_distr" = "cyan3", 
                                 "st_distr" = "#89FF99", 
@@
 
-* quantile function of truncated-shifted normal distribution and of shifted-truncated normal distribution:
+* quantile function of truncated-shifted Weibull distribution and of shifted-truncated Weibull distribution:
 
 ```{r}
-qstnorm <- qshift("tnorm")
-qtsnorm <- qtruncate("snorm")
+qstweibull <- qshift("tweibull")
+qtsweibull <- qtruncate("sweibull")
 x <- 1:999/1000
-std_quant <- qnorm(p = x, mean = 10, sd = 1)
-st_quant <- qstnorm(p = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
-ts_quant <- qtsnorm(p = x, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
+orig_quant <- qweibull(p = x, shape = 3, scale = 1)
+st_quant <- qstweibull(p = x, shape = 3, scale = 1, L = 3, U = 11, shift = 1)
+ts_quant <- qtsweibull(p = x, shape = 3, scale = 1, L = 3, U = 11, shift = 1)
 ```
@@
 df <- data.frame(x=x, 
-                 type = factor(c(rep("std_quant", times=length(x)),
+                 type = factor(c(rep("orig_quant", times=length(x)),
                     rep("st_quant", times=length(x)),
                     rep("ts_quant", times=length(x))),
-                    levels = c("std_quant", "st_quant", "ts_quant")), 
-                  value = c(std_quant, st_quant, ts_quant))
+                    levels = c("orig_quant", "st_quant", "ts_quant")), 
+                  value = c(orig_quant, st_quant, ts_quant))
 
@@
   geom_line(size=1) +
-  scale_color_manual(values = c("std_quant" = "cyan3", 
+  scale_color_manual(values = c("orig_quant" = "cyan3", 
                                 "st_quant" = "#89FF99", 
@@
 
-* random generation function of truncated-shifted normal distribution and of shifted-truncated normal distribution:
+* random generation function of truncated-shifted Weibull distribution and of shifted-truncated Weibull distribution:
 
 ```{r}
-rstnorm <- rshift("tnorm")
-rtsnorm <- rtruncate("snorm")
-std_rand <- rnorm(n=1000, mean = 10, sd = 1)
-st_rand <- rstnorm(n=1000, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
-ts_rand <- rtsnorm(n=1000, mean = 10, sd = 1, L = 9, U = 11, shift = 1)
+rstweibull <- rshift("tweibull")
+rtsweibull <- rtruncate("sweibull")
+orig_rand <- rweibull(n=1000, shape = 1.5, scale = 1)
+st_rand <- rstweibull(n=1000, shape = 1.5, scale = 1, L = 3, U = 11, shift = 1)
+ts_rand <- rtsweibull(n=1000, shape = 1.5, scale = 1, L = 3, U = 11, shift = 1)
 ```
@@
 df <- data.frame(x=1:1000, 
-                 type = factor(c(rep("std_rand", times=1000),
+                 type = factor(c(rep("orig_rand", times=1000),
                     rep("st_rand", times=1000),
                     rep("ts_rand", times=1000)),
-                    levels = c("std_rand", "st_rand", "ts_rand")), 
-                  value = c(std_rand, st_rand, ts_rand))
+                    levels = c("orig_rand", "st_rand", "ts_rand")), 
+                  value = c(orig_rand, st_rand, ts_rand))
 
@@
         geom_density(alpha=0.6) +
-        scale_fill_manual(values = c("std_rand" = "cyan3",
+        scale_fill_manual(values = c("orig_rand" = "cyan3",
                               "st_rand" = "#89FF99",
@@
 
+<!--
 ## Other examples
@@
 ```
+
+-->

### 2016-09-19 docs/qdist-master/R/dtruncate.R

@@ -1,0 +1,90 @@
+dtruncate <-
+  function (dist){ 
+    
+    #dist <- as.character(substitute(dist))
+    #envir <- as.environment(paste("package", as.character(substitute(pkg)), sep = ":"))
+    
+    ddist=paste("d", dist, sep = "") 
+    pdist=paste("p", dist, sep = "")
+    
+    #check parameters
+    #if(!missing(dist) & (!missing(ddist) | !missing(pdist))) 
+    #stop("Error! Either dist or ddist and pdist can be provided")
+    
+    
+    #gets density function                    
+    ddist <-  get(ddist, mode = "function")
+    #ddist <-  get(ddist, mode = "function", envir = envir)
+    #gets argument of density function
+    dargs <- formals(ddist)
+   
+    # Check first argument of density function is named 'x'
+    #if(names(dargs[1]) != "x") stop("Error: first argument of ddist must be named 'x'")
+      
+    
+    #gets probability function                    
+    #pdist <- get(pdist, mode = "function", envir = envir)
+    pdist <- get(pdist, mode = "function")
+    #gets argument of probability function
+    pargs <- formals(pdist)
+    # Checks first argument of density function is named 'q'
+    #if(names(pargs[1]) != "q") stop("Error: first argument of ddist must be named 'q'")
+    
+    
+    #Output function starts here
+    density <- function () 
+    {
+      #check L U 
+      if (L > U) stop("U must be greater than or equal to L")
+      
+      #gets density arguments
+      
+      call <- as.list(match.call())[-1]
+      
+      #call[is.element(names(call), names(dargs))] -> 
+      #arguments passed to density and belonging to dargs 
+      #i.e. arguments belonging to density function ddist
+      
+      #c(dargs[!is.element(names(dargs), names(call))]
+      #arguments belonging to dargs i.e. arguments belonging to density function ddist
+      # and not being part of arguments passed to density 
+      
+      #as a result, the whole string gets all unique arguments belonging to density function and ddist 
+      #dargs <- c(dargs[!is.element(names(dargs), names(call))], call[is.element(names(call), names(dargs))])
+      dargs <- intersect_args(x = dargs, y = call)
+      
+      #all unique arguments belonging to probability and pdist 
+      #pargs <- c(pargs[!is.element(names(pargs), names(call))], call[is.element(names(call), names(pargs))])
[…]

### 2016-09-19 R/dshift.R
qshift sistemato

@@ -65,8 +65,8 @@ dshift <- function(dist){
       dargs <- intersect_args(x = dargs, y = call)
 
       # method for computing density values for shifted distributions
-      dargs$x <- eval(dargs$x)
-      dargs$x <- dargs$x - shift
+      dargs$x <- x - shift
+
       density <- do.call("ddist", as.list(dargs))
 
       # returns density values for shifted distributions

### 2016-09-19 R/pshift.R

@@ -65,8 +65,7 @@ pshift <- function(dist){
     pargs <- intersect_args(x = pargs, y = call)
 
     # method for computing distribution values for shifted distribution
-    pargs$q <- eval(pargs$q)
-    pargs$q <- pargs$q - shift
+    pargs$q <- q - shift
     probability <-  do.call("pdist", as.list(pargs))
 
     # returns distribution values for shifted distributions

### 2016-09-19 test.R

@@ -1,0 +1,44 @@
+require(qshift)
+
+sdnorm <- dshift("norm")
+sdnorm(x = 1:10, mean = 2, sd = 3, shift = 4, log = T)
+sdnorm(x = 1:10, mean = 2, sd = 3, log = T)
+dnorm(x = 1:10, mean = 2, sd = 3, log = T)
+
+
+spnorm <- pshift("norm")
+
+spnorm(q = 1:10, mean = 2, sd = 3)
+pnorm(q = 1:10, mean = 2, sd = 3)
+
+spnorm(q = 1:10, mean = 2, sd = 3, shift=1)
+
+spnorm(q = 1:10, mean = 2, sd = 3, log.p = T, shift=1)
+pnorm(q = 1:10, mean = 2, sd = 3, log.p = T)
+
+spnorm(q = 1:10, mean = 2, sd = 3, lower.tail = F, shift=1)
+pnorm(q = 1:10, mean = 2, sd = 3, lower.tail = F)
+
+
+pnorm(q = 1:10, mean = 2, sd = 3, log.p = T, lower.tail = F)
+spnorm(q = 1:10, mean = 2, sd = 3, log.p = T, lower.tail = F, shift = 1)
+spnorm(q = 1:10, mean = 2, sd = 3, log.p = T, lower.tail = F, shift = 1)
+spnorm(q = 1:10, mean = 2, sd = 3, log.p = F, lower.tail = F, shift = 1)
+
+qsnorm <- qshift("norm")
+a <- qsnorm(p = log(1:9/10), mean = 2, sd = 3, shift = 4, log = T, lower.tail = F)
+spnorm(q = a, mean = 2, sd = 3, shift=4, log.p = T, lower.tail = F)
+
+a <- qsnorm(p = 1:9/10, mean = 2, sd = 3, shift = 4, log = F, lower.tail = F)
+spnorm(q = a, mean = 2, sd = 3, shift=4, log.p = F, lower.tail = F)
+
+
+
+qsnorm(p = 1:9/10, mean = 2, sd = 3, log = F, lower.tail = F)
+qnorm(p = 1:9/10, mean = 2, sd = 3, log = F, lower.tail = F)
+
+qsnorm(p = 1:9/10, mean = 2, sd = 3, shift = 5, log = F, lower.tail = F)
+
+
+
+

### 2016-09-19 test.R
qshift

@@ -33,6 +33,27 @@ a <- qsnorm(p = 1:9/10, mean = 2, sd = 3, shift = 4, log = F, lower.tail = F)
 spnorm(q = a, mean = 2, sd = 3, shift=4, log.p = F, lower.tail = F)
 
 
+spweibull <- pshift("weibull")
+qsweibull <- qshift("weibull")
+
+a <- qsweibull(p = log(1:9/10), shape = 2, scale = 3, shift = 4, log = T, lower.tail = F)
+spweibull(q = a, shape = 2, scale = 3, shift=4, log.p = T, lower.tail = F)
+
+a <- qsweibull(p = log(1:9/10), shape = 2, scale = 3, shift = -4, log = T, lower.tail = F)
+spweibull(q = a, shape = 2, scale = 3, shift=-4, log.p = F, lower.tail = F)
+
+a <- qsweibull(p = log(1:9/10), shape = 2, scale = 3, shift = -4, log = T, lower.tail = T)
+spweibull(q = a, shape = 2, scale = 3, shift=-4, log.p = T, lower.tail = T)
+
+a <- qsweibull(p = 1:9/10, shape = 2, scale = 3, shift = -4, log = F, lower.tail = F)
+spweibull(q = a, shape = 2, scale = 3, shift=-4, log.p = F, lower.tail = F)
+
+
+a <- qsweibull(p = 1:9/10, shape = 2, scale = 3, shift = -4, log = F, lower.tail = T)
+spweibull(q = a, shape = 2, scale = 3, shift=-4, log.p = F, lower.tail = T)
+
+
+
 
 qsnorm(p = 1:9/10, mean = 2, sd = 3, log = F, lower.tail = F)
 qnorm(p = 1:9/10, mean = 2, sd = 3, log = F, lower.tail = F)

### 2016-09-20 R/qshift.R
nuova versione funzioni

@@ -46,7 +46,7 @@
 #' two_shift <- qc(p = c(0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1), df = 2, shift = 2)
 #' two_shift
 #'
-qshift <- function (dist){
+qshift <- function(dist){
 
     qdist=paste("q", dist, sep = "")
 
@@ -56,7 +56,7 @@ qshift <- function (dist){
     qargs <-  formals(qdist)
 
     # Output function starts here
-    quantile <- function() {
+    quantile <- function(){
 
       # gets quantile arguments
       call <- as.list(match.call())[-1]

### 2016-09-20 R/dshift.R
qshift: reorder functions

@@ -1,5 +1,6 @@
 #' Compute density function of a shifted distribution
-#'
+#' @description This function generates the density function of a shifted specified distribution.
+#' Passing the distribution conventional name as argument, it returns the density function of the shifted specified distribution.
 #' @param dist character string indicating the distribution name. It can be set as:
 #' \itemize{
 #'   \item beta Beta Distribution
@@ -23,7 +24,7 @@
 #'   \item signrank Wilcoxon Signed Rank Statistic Distribution
 #' }
 #'
-#' @return numeric vector of the density function of a specified distribution
+#' @return density function of the shifted specified distribution
 #' @export
 #'
 #' @examples
@@ -46,36 +47,29 @@
 #' two_shift
 #'
 dshift <- function(dist){
-
-    ddist=paste("d", dist, sep = "")
-
-    # gets density function
-    ddist <-  get(ddist, mode = "function")
-
-    # gets argument of density function
-    dargs <- formals(ddist)
-
-    # Output function starts here
-    density <- function(){
-
-      # gets density arguments
-      call <- as.list(match.call())[-1]
-
-      # as a result, the whole string gets all unique arguments belonging to density function and ddist
-      dargs <- intersect_args(x = dargs, y = call)
-
-      # method for computing density values for shifted distributions
-      dargs$x <- x - shift
-
-      density <- do.call("ddist", as.list(dargs))
-
-      # returns density values for shifted distributions
-      return(density)
-
-    }
-
-    # add to density function formals shift with values as passed with dshift
-    formals(density) <-  c(formals(ddist), eval(substitute(alist(shift=0))))
-    # return density function
+  # generate density function name of the specified distribution
+  ddist=paste("d", dist, sep = "")
+  # gets density function and its arguments
+  ddist <-  get(ddist, mode = "function")
+  dargs <- formals(ddist)
+  
+  # Output function starts here
+  density <- function(){
+    # gets density arguments
+    call <- as.list(match.call())[-1]
+    # intersect density function and the call (ddist), giving to density function arguments the values set to the corresponding arguments of the call (ddist)
+    dargs <- intersect_args(x = dargs, y = call)
+    # method for computing density values for shifted distributions
+    dargs$x <- x - shift
+    density <- do.call("ddist", as.list(dargs))
+    
+    # returns density values for shifted distributions
     return(density)
+    
   }
+  
+  # add to density function formals shift with values as passed with dshift
[…]

### 2016-09-20 R/pshift.R

@@ -1,5 +1,6 @@
 #' Compute distribution function of a shifted distribution
-#'
+#' @description This function generates the distribution function of a shifted specified distribution.
+#' Passing the distribution conventional name as argument, it returns the distribution function of the shifted specified distribution.
 #' @param dist character string indicating distribution name. It can be set as:
 #' \itemize{
 #'   \item beta Beta Distribution
@@ -23,7 +24,7 @@
 #'   \item signrank Wilcoxon Signed Rank Statistic Distribution
 #' }
 #'
-#' @return numeric vector of the distribution function of a specified distribution
+#' @return distribution function of the truncated specified distribution
 #' @export
 #'
 #' @examples
@@ -46,37 +47,33 @@
 #' two_shift
 #'
 pshift <- function(dist){
-
-    pdist <- paste("p", dist, sep = "")
-
-    # gets distribution function
-    pdist <-  get(pdist, mode = "function")
-
-    # gets argument of distribution function
-    pargs <-  formals(pdist)
-
-    # Output function starts here
-    probability <- function(){
-
+  # generate distribution function name of the specified distribution
+  pdist <- paste("p", dist, sep = "")
+  # get distribution function and its arguments
+  pdist <-  get(pdist, mode = "function")
+  pargs <-  formals(pdist)
+  
+  # Output function starts here
+  probability <- function(){
+    
     # gets distribution arguments
     call <-  as.list(match.call())[-1]
-
-    # as a result, the whole string gets all unique arguments belonging to distribution function and pdist
+    # intersect distribution function and the call (pdist), giving to distribution function arguments the values set to the corresponding arguments of the call (pdist)
     pargs <- intersect_args(x = pargs, y = call)
-
+    
     # method for computing distribution values for shifted distribution
     pargs$q <- q - shift
     probability <-  do.call("pdist", as.list(pargs))
-
+    
     # returns distribution values for shifted distributions
     return(probability)
-
-    }
-
-    # add to distribution function formals shift with values as passed with pshift
-    formals(probability) <- c(pargs, eval(substitute(alist(shift=0))))
-    # return distribution function
-    return(probability)
-
+    
   }
+  
+  # add to distribution function formals shift with values as passed with pshift
+  formals(probability) <- c(pargs, eval(substitute(alist(shift=0))))
+  # return distribution function
+  return(probability)
+  
+}
 

### 2016-09-20 R/qshift.R

@@ -1,5 +1,6 @@
 #' Compute quantile function of a shifted distribution
-#'
+#' @description This function generates the quantile function of a shifted specified distribution.
+#' Passing the distribution conventional name as argument, it returns the quantile function of the shifted specified distribution.
 #' @param dist dist character string indicating distribution name. It can be set as:
 #' \itemize{
 #'   \item beta Beta Distribution
@@ -23,7 +24,7 @@
 #'   \item signrank Wilcoxon Signed Rank Statistic Distribution
 #' }
 #'
-#' @return numeric vector of the quantile function of a specified distribution
+#' @return quantile function of a shifted specified distribution
 #' @export
 #'
 #' @examples
@@ -47,35 +48,34 @@
 #' two_shift
 #'
 qshift <- function(dist){
-
-    qdist=paste("q", dist, sep = "")
-
-    # gets quantile function
-    # gets argument of quantile function
-    qdist <-  get(qdist, mode = "function")
-    qargs <-  formals(qdist)
-
-    # Output function starts here
-    quantile <- function(){
-
-      # gets quantile arguments
-      call <- as.list(match.call())[-1]
-
-      # as a result, the whole string gets all unique arguments belonging to quantile function and qdist
-      qargs <- intersect_args(x = qargs, y = call)
-
-      # method for computing quantiles values for shifted distributions
-       quantile <- do.call("qdist", as.list(qargs))
-       quantile <- quantile + shift
-
-       # returns quantile values for shifted distributions
-      return(quantile)
-
-    }
-    # add to quantile function formals shift with values as passed with qshift
-    formals(quantile) <-  c(formals(qdist), eval(substitute(alist(shift=0))))
-    # return quantile function
+  # generate quantile function name of the specified distribution
+  qdist=paste("q", dist, sep = "")
+  
+  # gets quantile function and its arguments
+  qdist <-  get(qdist, mode = "function")
+  qargs <-  formals(qdist)
+  
+  # Output function starts here
+  quantile <- function(){
+    
+    # gets quantile arguments
+    call <- as.list(match.call())[-1]
+    
+    # intersect quantile function and the call (qdist), giving to quantile function arguments the values set to the corresponding arguments of the call (qdist)
+    qargs <- intersect_args(x = qargs, y = call)
+    
+    # method for computing quantiles values for shifted distributions
+    quantile <- do.call("qdist", as.list(qargs))
+    quantile <- quantile + shift
+    
[…]

### 2016-09-20 R/rshift.R

@@ -1,5 +1,6 @@
 #' Perform Random generation for a shifted distribution
-#'
+#' @description This function generates the random number generator function of a shifted specified distribution.
+#' Passing the distribution conventional name as argument, it returns the random number generator function of the shifted specified distribution.
 #' @param dist character string indicating distribution name. It can be set as:
 #' \itemize{
 #'   \item beta Beta Distribution
@@ -24,7 +25,7 @@
 #' }
 #'
 #'
-#' @return numeric vector of random generated values from a specified distribution
+#' @return random number generator function of a shifted specified distribution
 #' @export
 #'
 #' @examples
@@ -66,13 +67,8 @@ rshift <- function(dist){
     # gets random generation arguments
     call <- as.list(match.call())[-1]
 
-    # as a result, the whole string gets all unique arguments belonging to random generation function and rdist
+    # intersect random number generator function and the call (rdist), giving to random number generator function arguments the values set to the corresponding arguments of the call (rdist)
     rargs <- intersect_args(x = rargs, y = call)
-
-    # Define starting point for random computation or not
-    if(set_seed){
-      set.seed(1)
-    }
 
     # method for computing random generated values for shifted distributions
     random <- do.call(rdist, as.list(rargs))
@@ -82,8 +78,8 @@ rshift <- function(dist){
 
   }
 
-  # add to random generation function formals shift and set_seed with values as passed with rshift
-  formals(random) <-  c(formals(rdist), eval(substitute(alist(shift=0, set_seed = FALSE))))
+  # add to random generation function formal shift with values as passed with rshift
+  formals(random) <-  c(formals(rdist), eval(substitute(alist(shift=0))))
   # return random generation function
   return(random)
 }

### 2016-09-20 docs/qdist-article/truncdist.Rmd
qdist article revised

@@
 * quantile functions `q<dist>(p)` with `p` vector of probabilities;
-* random number generation `r<dist>(n)` with `n` number of observations.
+* random number generator function `r<dist>(n)` with `n` number of observations.
 
@@
 
-We could easily write function `tpnorm()`, a truncated version for `pnorm()`, as:
+We could easily write function `ptnorm()`, a truncated version for `pnorm()`, as:
 
@@
 
-As an alternative approach, we decided to develop package `qdist` available at [https://github.com/andreaspano/qdist](https://github.com/andreaspano/qdist) .
+As an alternative approach, we decided to develop package `qdist` available at [https://github.com/veronicagiro/qtruncate](https://github.com/veronicagiro/qtruncate) .
 
@@
 ```{r, message=FALSE}
-devtools::install_github("andreaspano/qdist", build=FALSE)
+devtools::install_github("veronicagiro/qtruncate", build=FALSE)
 require(qdist)
-tpnorm <- ptruncate("norm")
-tdnorm <- dtruncate("norm")
-tqnorm <- qtruncate("norm")
-trnorm <- rtruncate("norm")
+ptnorm <- ptruncate("norm")
+dtnorm <- dtruncate("norm")
+qtnorm <- qtruncate("norm")
+rtnorm <- rtruncate("norm")
 ```
@@
 
-The newly generated function `tpnorm()` has the same formals as the original `pnorm()`, plus two
+The newly generated function `ptnorm()` has the same formals as the original `pnorm()`, plus two
 extra parameters: `L` and `U`, respectively for lower and upper truncation thresholds set by default to
@@
 args(pnorm)
-args(tpnorm)
+args(ptnorm)
 ```
 
-Once we have defined `tpnorm()` we can use it for generating probability values from a non-
+Once we have defined `ptnorm()` we can use it for generating probability values from a non-
 truncated normal distribution by leaving parameters `L` and `U` set to their defaults:
@@
 q <- seq(6, 14, len = 100)
-p_L_U <- tpnorm(q, mean = 10, sd = 1)
+p_L_U <- ptnorm(q, mean = 10, sd = 1)
 ```
@@
 ```{r}
-p_L9_U11 <- tpnorm(q, mean = 10, sd = 1, L = 9, U = 11)
+p_L9_U11 <- ptnorm(q, mean = 10, sd = 1, L = 9, U = 11)
 ```
@@
   q=q,
-  tpnorm_type = factor(c(rep("p_L_U", times=length(q)), rep("p_L9_U11", times=length(q))), levels = c("p_L_U", "p_L9_U11")),
+  ptnorm_type = factor(c(rep("p_L_U", times=length(q)), rep("p_L9_U11", times=length(q))), levels = c("p_L_U", "p_L9_U11")),
   value = c(p_L_U, p_L9_U11)
@@
 
-pl <- ggplot(data = df, mapping = aes(x=q, y=value, col=tpnorm_type)) +
+pl <- ggplot(data = df, mapping = aes(x=q, y=value, col=ptnorm_type)) +
   geom_line(size=1) + scale_color_manual(values = c("p_L_U" = "darkgray", "p_L9_U11" ="darkred")) +   
@@
 As a second example, we consider a density function for a left truncated Weibull distribution. We
-first generate function `tdweibull()` by using `dtruncate()`:
+first generate function `dtweibull()` by using `dtruncate()`:
 
 ```{r}
-tdweibull <- dtruncate("weibull")
+dtweibull <- dtruncate("weibull")
 ```
@@
 x <- qweibull(ppoints(1000), shape = 2, scale = 7)
-d_L0 <- tdweibull(x, shape = 2, scale = 7)
-d_L3 <- tdweibull(x, shape = 2, scale = 7, L= 3)
-d_L5 <- tdweibull(x, shape = 2, scale = 7, L= 5)
-d_L7 <- tdweibull(x, shape = 2, scale = 7, L= 7)
+d_L0 <- dtweibull(x, shape = 2, scale = 7)
+d_L3 <- dtweibull(x, shape = 2, scale = 7, L= 3)
+d_L5 <- dtweibull(x, shape = 2, scale = 7, L= 5)
+d_L7 <- dtweibull(x, shape = 2, scale = 7, L= 7)
 ```
@@
   x=x,
-  tdweibull_type = factor(c(rep("d_L0", times=length(x)), rep("d_L3", times=length(x)), rep("d_L5", times=length(x)), rep("d_L7", times=length(x))), levels = c("d_L0", "d_L3", "d_L5", "d_L7")),
+  dtweibull_type = factor(c(rep("d_L0", times=length(x)), rep("d_L3", times=length(x)), rep("d_L5", times=length(x)), rep("d_L7", times=length(x))), levels = c("d_L0", "d_L3", "d_L5", "d_L7")),
   value = c(d_L0, d_L3, d_L5, d_L7)
@@
 
-pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=tdweibull_type)) +
+pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=dtweibull_type)) +
   geom_line(size=1) + 
@@
 require(evd)
-trgumbel <- rtruncate("gumbel")
-args(trgumbel)
-rg <- trgumbel(10^4, loc = 3, scale = 2, L = 5)
+rtgumbel <- rtruncate("gumbel")
+args(rtgumbel)
+rg <- rtgumbel(10^4, loc = 3, scale = 2, L = 5)
 df <- data.frame(rg=rg)
@@
 ```{r}
-tqpois <- qtruncate("pois")
+qtpois <- qtruncate("pois")
 ```
@@
 # No truncation
-q <- tqpois(p, lambda = 10)
+q <- qtpois(p, lambda = 10)
 # Lower truncation L = 9
-q_L9 <- tqpois(p, lambda = 10, L = 9)
+q_L9 <- qtpois(p, lambda = 10, L = 9)
 # Upper truncation H = 14
-q_U14 <- tqpois(p, lambda = 10, U = 14)
+q_U14 <- qtpois(p, lambda = 10, U = 14)
 # Both sides truncation L = 9, H = 14
-q_L9_U14 <- tqpois(p, lambda = 10, L = 9, U = 14)
+q_L9_U14 <- qtpois(p, lambda = 10, L = 9, U = 14)
 ```
@@
 df <- data.frame(p=p,
-  tqpois_type=factor(c(rep("No truncation", times=length(p)), rep("Lower truncation L = 9", times=length(p)), 
+  qtpois_type=factor(c(rep("No truncation", times=length(p)), rep("Lower truncation L = 9", times=length(p)), 
     rep("Upper truncation H = 14", times=length(p)), rep("Both sides truncation L = 9, H = 14", times=length(p))), 
@@
 
-pl <- ggplot(data = df, mapping = aes(x=value, y=p, col=tqpois_type)) +
+pl <- ggplot(data = df, mapping = aes(x=value, y=p, col=qtpois_type)) +
   geom_line(size=1, show.legend = FALSE) + 
   ylab("Probability") + xlab("Quantiles") + ggtitle("Quantiles from truncated Poisson distribution") +
-  facet_wrap(~tqpois_type, nrow = 2) +
+  facet_wrap(~qtpois_type, nrow = 2) +
   theme(strip.text = element_text(size=12), plot.title = element_text(size=18), axis.title=element_text(size=14))
@@
 ```{r}
-tqgamma <- qtruncate("gamma")
+qtgamma <- qtruncate("gamma")
 ```
 
-When we use `tqgamma()` with parameter rate it works but may return a set of `warnings`:
+When we use `qtgamma()` with parameter rate it works but may return a set of `warnings`:
 
 ```{r}
-tqgamma(.25, shape = 1, rate = .3)
+qtgamma(.25, shape = 1, rate = .3)
 ```
@@
 ```{r, error=TRUE}
-tqgamma(ppoints(10), shape = 1, scale = 3)
+qtgamma(ppoints(10), shape = 1, scale = 3)
 ```
@@
 ```{r, error=TRUE}
-tqgamma <- qtruncate("gamma")
-tqgamma(p = .25, shape = 1, rate = 3)
+qtgamma <- qtruncate("gamma")
+qtgamma(p = .25, shape = 1, rate = 3)
 ```
@@
 ```{r}
-tdnorm <- dtruncate("norm")
+dtnorm <- dtruncate("norm")
 ```
@@
     sd <- theta[2]
-    ld <- tdnorm(x = x, mean = mean, sd = sd, L = L , U = U, log = TRUE)
+    ld <- dtnorm(x = x, mean = mean, sd = sd, L = L , U = U, log = TRUE)
     -sum(ld)
@@
 ```{r}
-trnorm <- rtruncate("norm")
-x <- trnorm(n = 1000, mean = 5, sd = 2, L = 3, U = 6)
+rtnorm <- rtruncate("norm")
+x <- rtnorm(n = 10000, mean = 5, sd = 2, L = 3, U = 6)
 ltnorm( x = x, L = 3, U = 6)
@@
 
+### log, log.p and lower.tail formals
+
+Unlike `truncdist` package, `qdist` allows us to use also the functionalities of formals `log` for density function, `log.p` and `lower.tail` for distribution and quantile functions.   
+`qdist` functions uses `stats` distribution functions for computing from truncated distributions models, but not for performing the functionalities of `log`, `log.p` and `lower.tail` arguments, as they would have been incorrect applied to a truncated distribution.
+`log`, `log.p` and `lower.tail` functionalities are computed by using an alternative methods that produce results extremely close to the original ones.
+
+<!--
+Considering the Normal distribution as example, `ptruncate()` function uses `pnorm()` function to built probability function for truncated Normal distribution, defined as `ptnorm()`.
+However, `log.p` and `lower.tail` formals of the resulting function `ptnorm()` are not computed with the help of `pnorm()` function because their functionalities would have been incorrect applied to a truncated distribution.
+The functionalities of `log.p` and `lower.tail` are computed by using an alternative method that produces results extremely close to the original ones.
+
+-->
+
 ## Bibliography

### 2016-09-20 docs/qshift-article/article.Rmd
shift article revised

---
title: 'R package qshift: A functional programming approach to shifted probability
  distributions'
output:
  html_document:
    highlight: null
    theme: null
  pdf_document: default
---

## Introduction

Shifted distributions are used daily in many statistical fields.  
Given a random variable $X$ with distribution function $F(x)$, a shifted random variable $Y$ from $X$ is a random variable with a constant, $k$, added, defined as $Y = X+k$, with distribution function $F(x-k)$. $Y$ support changes according to shift value.

The aim of this paper is to describe an R package, `qshift` which generates shifted distribution functions for any distribution chosen. The package was realized using a functional programming approach and inspires its computational method to Andrea Spanò article "R package `qdist`: A functional programming approach to truncated probability distributions", which provides an alternative method for computing from truncated probability distributions using a functional programming approach.

## Shifted probability functions in R

R does not provide a specific package for shifted distribution functions, but it provides package `stats` which contains a wide set of functions for standard distributions: probability, density, quantile and random number generation functions. All them share a stable naming convention both for the names of the functions and in the first argument of these functions. In particular:

* density functions: `d<dist>(x)` with `x` vector of quantiles;
* probability distribution functions: `p<dist>(q)` with `q` vector of quantiles;
* quantile functions: `q<dist>(p)` with `p` vector of probabilities;
* random number generation: `r<dist>(n)` with `n` number of observations.

where `<dist>` indicates the conventional name of distribution family.

Let us see an example.  
Considering the Normal distribution (`norm`), we have: `dnorm(x)`, `pnorm(q)`, `qnorm(p)` and `rnorm(n)`.

So, we use to write:

```{r dnorm}
dnorm(x = 6:10, mean = 8, sd = 1)
```

to get density values at 6:10 from a Normal distribution with parameters `mean = 8` and `sd = 1`.

Going back to shifted distribution, we could easily write a shifted version for `dnorm()`, as:

```{r sdnorm_definition}
shift_dnorm <- function(x, mean = 0, sd = 1, shift = 0, ...){
            x_shifted <- x - shift
            density <- dnorm(x = x_shifted, mean = mean, sd = sd, ...)
            return(density)
}
```

So we replicate the previous example but shifting the Normal distribution of 1 unit:

```{r sdnorm_example_1}
shift_dnorm(x = 6:10, mean = 8, sd = 1, shift = 1)
```

At first view, this approach seems easy and cool but thinking about it, we realize that we would need to write a different function for each probability distribution we would like to work with. Functional programming approach could be a solution, as Andrea Spanò demonstrates in his article (see "R package `qdist`: A functional programming approach to truncated probability distributions" for more details).

`qshift` package includes four general functions which compute shifted distribution functions for any distribution chosen:

* `dshift()` for shifted density functions;
* `pshift()` for shifted probability distributions;
* `qshift()` for shifted quantile functions;
* `rshift()` for shifted random numbers generation functions.

`qshift` package is available at [https://github.com/veronicagiro/qshift](https://github.com/veronicagiro/qshift).

```{r github, eval = FALSE}
devtools::install_github("veronicagiro/qshift")
```

The four functions share the same logic: they take as input a probability distribution name, as a character string, and return the equivalent shifted distribution as a function object.

```{r qshift, message=FALSE}
require(qshift)
dsnorm <- dshift("norm") # density function
psnorm <- pshift("norm") # probability function
qsnorm <- qshift("norm") # quantile function
rsnorm <- rshift("norm") # random generation function
```

Talking about the computational process, package `qshift` is based on two key concepts being part of the functional nature of the R programming language:

* the environment of a function can be used as a placeholder for other objects;
* function formals can be manipulated.

For example, let us consider `dshift()`. As we seen, it takes as input the distribution name (`norm`) and proceeds as follows:

* gets the corresponding function `dnorm()`;
* uses `dnorm()` to create a function, say `density()` that computes shifted density for normal distributions;
* modify the formals of `density()` so that it has the same formals as `dnorm()` plus `shift`, corresponding to the shift value;
* returns `density()`.

So, the function `dsnorm()` has the same formals of the function `dnorm()`, plus the parameter `shift` which indicates the shift value to be applied to distribution; by default it is set as 0.

```{r args}
args(dsnorm)
args(dnorm)
```

Other references are in chapters [Environments](http://www.quantide.com/website/ramarro-chapter-03/) and [Functions](http://www.quantide.com/website/ramarro-chapter-04/) of Andrea Spanò weeb Book [Ramarro](http://www.quantide.com/website/ramarro-r-for-developers/), available on Quantide web site.

## Examples

Let us consider the Normal distribution and the previously defined functions `dsnorm()`, `psnorm()`, `qsnorm()` and `rsnorm()`.

* Density function, `dsnorm()`.  
    This function has a double use:

    * generate probability values from a non-shifted normal distribution by leaving parameter `shift` set to its default (`shift = 0`)
    * generate probability values from any shifted normal distribution by setting `shift` value

    ```{r sdnorm_example}
x <- seq(4, 14, len = 100)
not_shifted <- dsnorm(x = x, mean = 8, sd = 1.5)
positive_shifted <- dsnorm(x = x, mean = 8, sd = 1.5, shift = 1)
negative_shifted <- dsnorm(x = x, mean = 8, sd = 1.5, shift = -1)
    ```
    We can visualize these results by plotting them:

    ```{r sdnorm_plot, message=FALSE}
require(ggplot2)
df <- data.frame(x=x, 
                 dsnorm_type = factor(c(rep("not_shifted", times=length(x)),
                    rep("positive_shifted", times=length(x)),
                    rep("negative_shifted", times=length(x))),
                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
                  value = c(not_shifted, positive_shifted, negative_shifted))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=dsnorm_type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("not_shifted" = "cyan3", 
                                "positive_shifted" = "#89FF99", 
                                "negative_shifted" = "#FF89B4")) +
  xlab("Value") + ylab("Density") +
  ggtitle("Density plot for shifted and not-shifted Normal distributions") +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

print(pl)
    ```

* Probability function, `spnorm()`:

    ```{r spnorm_example}
not_shifted <- psnorm(q = x, mean = 8, sd = 1.5)
positive_shifted <- psnorm(q = x, mean = 8, sd = 1.5, shift = 1)
negative_shifted <- psnorm(q = x, mean = 8, sd = 1.5, shift = -1)
    ```
    We can visualize these results by plotting them:

    ```{r spnorm_plot}
df <- data.frame(x=x, 
                 psnorm_type = factor(c(rep("not_shifted", times=length(x)),
                    rep("positive_shifted", times=length(x)),
                    rep("negative_shifted", times=length(x))),
                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
                  value = c(not_shifted, positive_shifted, negative_shifted))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=psnorm_type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("not_shifted" = "cyan3", 
                                "positive_shifted" = "#89FF99", 
                                "negative_shifted" = "#FF89B4")) +
  xlab("Quantile") + ylab("Probability") +
  ggtitle("Probability plot for shifted and not-shifted Normal distributions") +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

print(pl)
    ```

* Quantile function, `qsnorm()`:

    ```{r sqnorm_example}
x <- (1:999)/1000
not_shifted <- qsnorm(p = x, mean = 8, sd = 1.5)
positive_shifted <- qsnorm(p = x, mean = 8, sd = 1.5, shift = 1)
negative_shifted <- qsnorm(p = x, mean = 8, sd = 1.5, shift = -1)
    ```
    We can visualize these results by plotting them:

    ```{r sqnorm_plot}
df <- data.frame(x=x, 
                 qsnorm_type = factor(c(rep("not_shifted", times=length(x)),
                    rep("positive_shifted", times=length(x)),
                    rep("negative_shifted", times=length(x))),
                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
                  value = c(not_shifted, positive_shifted, negative_shifted))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=qsnorm_type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("not_shifted" = "cyan3", 
                                "positive_shifted" = "#89FF99", 
                                "negative_shifted" = "#FF89B4")) +
  ylab("Quantile") + xlab("Probability") +
  ggtitle("Quantile plot for shifted and not-shifted Normal distributions") +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

print(pl)
    ```

* Random generation function, `rsnorm()`:

    ```{r srnorm_example}
    x <- 100000
not_shifted <- rsnorm(n = x, mean = 8, sd = 1.5)
positive_shifted <- rsnorm(n = x, mean = 8, sd = 1.5, shift = 3)
negative_shifted <- rsnorm(n = x, mean = 8, sd = 1.5, shift = -3)
    ```
    We can visualize these results by plotting them:

    ```{r srnorm_plot}
df <- data.frame(rsnorm_type = factor(c(rep("not_shifted", times=x),
                    rep("positive_shifted", times=x),
                    rep("negative_shifted", times=x)),
                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
                  value = c(not_shifted, positive_shifted, negative_shifted))

pl <- pl <- ggplot(data = df, mapping = aes(x=value, group=rsnorm_type, fill=rsnorm_type)) +
  geom_density(alpha=0.6) +
    scale_fill_manual(values = c("not_shifted" = "cyan3", 
                                "positive_shifted" = "#89FF99", 
                                "negative_shifted" = "#FF89B4")) +
  ggtitle("Density plot for shifted and not-shifted Normal distributions") +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

print(pl)
    ```

As expected, the Normal shifted distribution is a Normal distribution with changed mean. The change in the mean is equal to the shift.

As we know, Normal distribution support goes from $-\infty$ to $+\infty$. What happens considerig a distribution with a limited support?   
Let us consider the Uniform distribution.

First of all we have to define its probability functions:

```{r qshift_unif}
dsunif <- dshift("unif")
psunif <- pshift("unif")
qsunif <- qshift("unif")
rsunif <- rshift("unif")
```

We limited Uniform support between 6 and 12.

* Density function, `dsunif()`:

    ```{r sdunif_example}
x <- seq(4, 14, len = 100)
not_shifted <- dsunif(x = x, min = 6, max = 12)
positive_shifted <- dsunif(x = x, min = 6, max = 12, shift = 1)
negative_shifted <- dsunif(x = x, min = 6, max = 12, shift = -1)
    ```
    We can visualize these results by plotting them:

    ```{r sdunif_plot}
df <- data.frame(x=x, 
                 dsunif_type = factor(c(rep("not_shifted", times=length(x)),
                    rep("positive_shifted", times=length(x)),
                    rep("negative_shifted", times=length(x))),
                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
                  value = c(not_shifted, positive_shifted, negative_shifted))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=dsunif_type)) +
  geom_line(size=1, alpha =0.8) +
  scale_color_manual(values = c("not_shifted" = "cyan3", 
                                "positive_shifted" = "#89FF99", 
                                "negative_shifted" = "#FF89B4")) +
  xlab("Value") + ylab("Density") +
  ggtitle("Density plot for shifted and not-shifted Uniform distributions") +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

print(pl)
    ```

* Probability function, `psunif()`:

    ```{r spunif_example}
not_shifted <- psunif(q = x, min = 6, max = 12)
positive_shifted <- psunif(q = x, min = 6, max = 12, shift = 1)
negative_shifted <- psunif(q = x, min = 6, max = 12, shift = -1)
    ```
    We can visualize these results by plotting them:

    ```{r spunif_plot}
df <- data.frame(x=x, 
                 psunif_type = factor(c(rep("not_shifted", times=length(x)),
                    rep("positive_shifted", times=length(x)),
                    rep("negative_shifted", times=length(x))),
                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
                  value = c(not_shifted, positive_shifted, negative_shifted))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=psunif_type)) +
  geom_line(size=1, alpha =0.8) +
  scale_color_manual(values = c("not_shifted" = "cyan3", 
                                "positive_shifted" = "#89FF99", 
                                "negative_shifted" = "#FF89B4")) +
  xlab("Quantile") + ylab("Probability") +
  ggtitle("Probability plot for shifted and not-shifted Uniform distributions") +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

print(pl)
    ```

* Quantile function, `qsunif()`:

    ```{r squnif_example}
x <- (1:999)/1000
not_shifted <- qsunif(p = x, min = 6, max = 12)
positive_shifted <- qsunif(p = x, min = 6, max = 12, shift = 1)
negative_shifted <- qsunif(p = x, min = 6, max = 12, shift = -1)
    ```
    We can visualize these results by plotting them:

    ```{r squnif_plot}
df <- data.frame(x=x, 
                 qsunif_type = factor(c(rep("not_shifted", times=length(x)),
                    rep("positive_shifted", times=length(x)),
                    rep("negative_shifted", times=length(x))),
                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
                  value = c(not_shifted, positive_shifted, negative_shifted))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=qsunif_type)) +
  geom_line(size=1, alpha =0.8) +
  scale_color_manual(values = c("not_shifted" = "cyan3",  
                                "positive_shifted" = "#89FF99", 
                                "negative_shifted" = "#FF89B4")) +
  ylab("Quantile") + xlab("Probability") +
  ggtitle("Quantile plot for shifted and not-shifted Uniform distributions") +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

print(pl)
    ```

* Random generation function, `rsunif()`:

    ```{r srunif_example}
    x <- 1000
not_shifted <- rsunif(n = x, min = 6, max = 12)
positive_shifted <- rsunif(n = x, min = 6, max = 12, shift = 3)
negative_shifted <- rsunif(n = x, min = 6, max = 12, shift = -3)
    ```
    We can visualize these results by plotting them:

    ```{r srunif_plot, message = FALSE}
df <- data.frame(rsunif_type = factor(c(rep("not_shifted", times=x),
                    rep("positive_shifted", times=x),
                    rep("negative_shifted", times=x)),
                    levels = c("not_shifted", "positive_shifted", "negative_shifted")), 
                  value = c(not_shifted, positive_shifted, negative_shifted))

pl <- ggplot(data = df, mapping = aes(x=value, fill=rsunif_type, group=rsunif_type)) +
  geom_histogram(alpha=0.5, bins = 30, position="identity") +
    scale_color_manual(values = c("not_shifted" = "cyan3", 
                                "positive_shifted" = "#89FF99", 
                                "negative_shifted" = "#FF89B4")) +
  xlab("Value") + ylab("Density") +
  ggtitle("Histogram for shifted and not-shifted Uniform distributions") +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

print(pl)
    ```

As we see, Uniform support changes according to the transformation (the shift) applied to the distribution.

## Extending the computation

Further implementations are possible starting from the definition of a shifted distribution.  
Suppose we want to define a function for maximum likelihood estimate for the shifted Weibull distribution.

Firstly let us define the shifted density function for the Weibull distribution:

```{r sdweibull_def}
dsweibull <- dshift("weibull")
```

and, subsequently, by using `dsweibull()` within an estimator function:

```{r max_likelihood_est}
ltweibull <- function(x){
  # starting values for parameters (scale, shape) and shift
  shift <- min(x) - 0.01
  x1 <- x - shift
  shape <- (sd(x1)/mean(x1))^(-1.086)
  scale <- mean(x1)/gamma(1+1/shape)
  # parameters vector definition
  theta <- c(shape, scale, shift)
  # likelihood function
  ll <- function(theta, x){
    shape <- theta[1]
    scale <- theta[2]
    shift <- theta[3]
    ld <- dsweibull(x = x, shape = shape, scale = scale, shift = shift, log = TRUE)
    -sum(ld)
  } 
  # maximum likelihood estimation
  optim(par = theta, fn = ll, x = x)[["par"]]
}
```

Let us see how it works on a random generated sample from a shifted Weibull distribution:

```{r max_likelihood_est_example}
# generate a random sample from a shifted weibull distribution
rsweibull <- rshift("weibull")
x <- rsweibull(n = 100000, shape = 1, scale = 5, shift = 1)

# maximum likelihood estimate for the shifted weibull distribution
ltweibull(x = x)
```

## Bibliography

Pace L., Salvan A. . _Introduzione alla Statistica_. CEDAM. 2010

Spanò A. .  _R package qdist: A functional programming approach to truncated probability ditributions_. url

Wickham H. . _Advanced r_. http://adv-r.had.co.nz/. Accessed: 2016-08-24.

### 2016-09-20 docs/shift+trunc-article/prove-qdist+qshift.R
articolo distr trunc e shift: prove maz loglikhood ok!!!

@@ -1138,7 +1138,7 @@ ltweibull <- function(x, L = -Inf, U = Inf){
     shape <- theta[1]
     scale <- theta[2]
     shift <- theta[3]
-    ld <- log(dtsweibull(x = x, shape = shape, scale = scale, shift = shift, L = L, U = U))
+    ld <- dtsweibull(x = x, shape = shape, scale = scale, shift = shift, L = L, U = U, log=T)
     -sum(ld)
   }
   # maximum likelihood estimation
@@ -1153,9 +1153,11 @@ x <- rtsweibull(n = 100000, shape = 1, scale = 1.5, shift = 1, L = 5, U = 10)
 # maximum likelihood estimate for the shifted weibull distribution
 ltweibull(x = x, L = 5, U = 10)
 
+## funzionaaaaaaaaaaaa
+
 ##############################################################################
 
-### troncata shiftata -> non funziona
+### troncata shiftata -> funzionaaaaaaaaaaa
 
 dtweibull <- dtruncate("weibull")
 dstweibull <- dshift("tweibull")
@@ -1165,6 +1167,7 @@ ltweibull <- function(x, shift = 0 ,L = -Inf, U = Inf){
   # starting values for parameters (scale, shape) and shift
   #shift <- min(x) - 0.01
   x1 <- x - shift
+  #x1 <- x
   shape <- (sd(x1)/mean(x1))^(-1.086)
   scale <- mean(x1)/gamma(1+1/shape)
   # parameters vector definition
@@ -1175,21 +1178,21 @@ ltweibull <- function(x, shift = 0 ,L = -Inf, U = Inf){
     shape <- theta[1]
     scale <- theta[2]
     #shift <- theta[3]
-    ld <- log(dstweibull(x = x, shape = shape, scale = scale, shift = shift, L = L, U = U))
+    ld <- dstweibull(x = x, shape = shape, scale = scale, shift = shift, L = L, U = U, log=TRUE)
     -sum(ld)
   }
   # maximum likelihood estimation
-  optim(par = theta, fn = ll, x = x, L = L, U = U, method = "Nelder-Mead")[["par"]]
+  optim(par = theta, fn = ll, x = x, shift=shift, L = L, U = U, method = "Nelder-Mead")[["par"]]
 }
 
 
 rtweibull <- rtruncate("weibull")
 rstweibull <- rshift("tweibull")
[…]

### 2016-09-21 docs/shift+trunc-article/article-shift+trunc.Rmd
truncated and shifted article

---
title: "R functional programming approach for truncated and shifted and for shifted and truncated probability distribution functions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

R programming functional nature allows programmer to implement and manipulate functions as needed. Functional programming is a style of programming that emphasizes the evaluation of expressions rather than the execution of commands, so these functions produce results that depend only on their inputs and not on the program state. This style has a great power and includes lots of tools that combined together can allows us to produce results with the most efficient computational method. One of them is the functional factory, which is a factory for making new functions. It allows us to use closures, which are functions together with a referencing environment, to write specific functions that, in turn, can be used to generate new functions. So we can have a double layer of development: a first layer that is used to do all the complex work in common to all functions and a second layer that defines the details of each function. This fundamentals was used to develop the computational method proposed by Andrea Spanò in "R package qdist: A functional programming approach to truncated probability distributions", and that I used in "R package qshift: A functional programming approach to shifted probability distributions", that produces functions for truncated and shifted distributions, respectively.  [`qdist`](https://github.com/veronicagiro/qdist) and [`qshift`](https://github.com/veronicagiro/qshift) packages are available on Github:

```{r, eval=FALSE}
devtools::install_github("veronicagiro/qtruncate")
devtools::install_github("veronicagiro/qshift")
```

In the previous citated articles, we dealt separately with shifted and truncated distributions. But in real life, we sometimes work with random variables arising from more than one transformation. So in this article, we will demonstrate that we can produce functions for truncated and shifted or for shifted and truncated distributions only combining the functions of the two packages. This is possible because of functional programming approach used to build them.

# Distribution functions for truncated and shifted or for shifted and truncated distributions

R provides density, probability, quantile and random number generator functions in `stats` package. These functions share a stable naming convention both for the names of the functions and in the first argument of these functions. In particular:

* `d<dist>(x)` with `x` vector of quantiles, identifies density functions
* `p<dist>(q)` with `q` vector of quantiles, identifies probability distribution functions
* `q<dist>(p)` with `p` vector of probabilities, identifies quantile functions
* `r<dist>(n)` with `n` number of observations, identifies random number generation

where `<dist>` indicates the distribution family.

As a result we have: `dweibull(x)`, `pweibull(q)`, `qweibull(p)` and `rweibull(n)` for Weibull distribution and similarly for all other distributions.

Andrea Spanò proposed in `qdist` a functional programming method that computes truncated distributions functions for each distribution chosen without changing the well structured programming style and syntax of previous functions.

The computational method proposed in `qdist` is based on two R key concepts:

* the environment of a function can be used as a placeholder for other objects (for detail see Andrea Spanò article "R package qdist: A functional programming approach to truncated probability ditributions"")
* function formals can be manipulated

and it is made up of two steps:

1. defines the required functions suitable for computing from specific truncated probability distribution by using the functions available within its package. These functions take as input a probability distribution as a character string, in this case the Weibull distribution, and return the equivalent truncated distribution function as object:

```{r, message=FALSE}
require(qdist)
dtweibull <- dtruncate("weibull") # density function
ptweibull <- ptruncate("weibull") # probability function 
qtweibull <- qtruncate("weibull") # quantile function
rtweibull <- rtruncate("weibull") # random generator function
```

2. uses the newly created functions (`dtweibull()`, `ptweibull()`, `qtweibull()`, `rtweibull()`) for computing from truncated probability distributions.

Let us consider `dtweibull()`. It has the same formals as the original `dweibull()`, plus two extra parameters: `L` and `U`, respectively for lower and upper truncation thresholds set by default to `-Inf` and `+Inf`:

```{r}
args(dweibull)
args(dtweibull)
```

As `dweibull()` is the function that computes the density function for a Weibull distribution, conventionally named `weibull`, `dtweibull()` is the function that computes the density function for a truncated Weibull distributions, named `tweibull`.

Suppose we want to generate the density function for a **truncated and shifted** Weibull distribution. We need `qshift` package. `qshift` package shares the same logic and the same methodology of `qdist`. The difference is that the functions are designed to compute functions for shifted distributions.
In particular, `qshift` package includes the following four functions:

* `dshift()`, for density function
* `pshift()`, for distribution function
* `qshift()`, for quantile function
* `rshift()`, for random generator function

So for computing the density function of a trucated and shifted Weibull distribution we have to set `tweibull` distribution as `dshift()` argument:

```{r, message=FALSE}
require(qshift)
dstweibull <- dshift("tweibull")
```

So `dstweibull()` is the function that computes the density function of `stweibull`, which is the truncated and shifted Weibull distribution.

```{r}
args(dstweibull)
```

Looking at the function arguments, we see that the syntax style is that of `dweibull()` function plus `L` and `U` arguments for truncated component of the distribution and `shift` argument for shifted component of the distribution.

<!--
va bene dire componente troncata e componente shiftata della distribuzione?
-->

Let us graphically compare the results of `dweibull()`, `dtweibull()` and `dstweibull()` functions:

```{r}
x <- seq(0, 5, len = 100)
orig_density <- dweibull(x = x, shape = 1.5, scale = 1)
trunc_density <- dtweibull(x = x, shape = 1.5, scale = 1, L = 1.5, U = 3)
trunc_shift_density <- dstweibull(x = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
```

```{r, message=FALSE}
df <- data.frame(x=x, 
                 type = factor(c(rep("orig_density", times=length(x)),
                    rep("trunc_density", times=length(x)),
                    rep("trunc_shift_density", times=length(x))),
                    levels = c("orig_density", "trunc_density", "trunc_shift_density")), 
                  value = c(orig_density, trunc_density, trunc_shift_density))
require(ggplot2)
pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("orig_density" = "cyan3", 
                                "trunc_density" = "#89FF99", 
                                "trunc_shift_density" = "#FF89B4")) +
  xlab("Value") + ylab("Density") + ylim(c(0,1.9)) +
  ggtitle(paste("Density plot for non-transformed, truncated", "and truncated and shifted Weibull distributions", sep = "\n")) +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

print(pl)
```

The inverse procedure generates the density function for a **shifted and truncated** Weibull distribution.

Starting from `dshift()` function of `qshift` package, we define `dsweibull()`, the density function of `sweibull`, which is the shifted Weibull distribution:
```{r}
dsweibull <- dshift("weibull") # density function
```

We have to define also the others functions for shifted distribution, as they are implicitly necessary for the further step:

```{r}
psweibull <- pshift("weibull") # probability function 
qsweibull <- qshift("weibull") # quantile function
rsweibull <- rshift("weibull") # random generator function
```

The density function of a shifted and truncated Weibull distribution is generated setting `sweibull` distribution as `ptruncate()` argument:
```{r}
dtsweibull <- dtruncate("sweibull")
```

So `dtsweibull()` is the function that computes the density function of `tsweibull`, which is the shifted and truncated Weibull distribution.

Let us graphically represent the output of the three functions:

```{r}
orig_density <- dweibull(x = x, shape = 1.5, scale = 1)
shift_density <- dsweibull(x = x, shape = 1.5, scale = 1, shift = 1)
shift_trunc_density <- dtsweibull(x = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
```

```{r}
df <- data.frame(x=x, 
                 type = factor(c(rep("orig_density", times=length(x)),
                    rep("shift_density", times=length(x)),
                    rep("shift_trunc_density", times=length(x))),
                    levels = c("orig_density", "shift_density", "shift_trunc_density")), 
                  value = c(orig_density, shift_density, shift_trunc_density))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("orig_density" = "cyan3", 
                                "shift_density" = "#89FF99", 
                                "shift_trunc_density" = "#FF89B4")) +
  xlab("Value") + ylab("Density") + ylim(c(0,1.9)) +
  ggtitle(paste("Density plot for non-transformed, shifted and","shifted and truncated Normal distributions", sep = "\n")) +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

print(pl)
```

Let us see graphically the differences between the density function of Weibull distribution, of a truncated and shifted Weibull distribution and of a shifted and truncated Weibull distribution.

```{r}
df <- data.frame(x=x, 
                 type = factor(c(rep("orig_density", times=length(x)),
                    rep("trunc_shift_density", times=length(x)),
                    rep("shift_trunc_density", times=length(x))),
                    levels = c("orig_density", "trunc_shift_density", "shift_trunc_density")), 
                  value = c(orig_density, trunc_shift_density, shift_trunc_density))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("orig_density" = "cyan3", 
                                "trunc_shift_density" = "#89FF99", 
                                "shift_trunc_density" = "#FF89B4")) +
  xlab("Value") + ylab("Density") + ylim(c(0,1.9)) +
  ggtitle(paste("Density plot for non-transformed, truncated-shifted", 
                "and shifted-truncated Weibull distribution", sep="\n")) +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

print(pl)
```

Looking at the figure we can see that a truncated and shifted distribution is different from a shifted amd truncated distribution.

Let us analyze the other distributions functions:

* probability distribution function of a truncated and shifted Weibull distribution and of a shifted and truncated Weibull distribution:

```{r}
pstweibull <- pshift("tweibull")
ptsweibull <- ptruncate("sweibull")
orig_distr <- pweibull(q = x, shape = 1.5, scale = 1)
shift_trunc_distr <- pstweibull(q = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
trunc_shift_distr <- ptsweibull(q = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
```

```{r}
df <- data.frame(x=x, 
                 type = factor(c(rep("orig_distr", times=length(x)),
                    rep("shift_trunc_distr", times=length(x)),
                    rep("trunc_shift_distr", times=length(x))),
                    levels = c("orig_distr", "shift_trunc_distr", "trunc_shift_distr")), 
                  value = c(orig_distr, shift_trunc_distr, trunc_shift_distr))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("orig_distr" = "cyan3", 
                                "shift_trunc_distr" = "#89FF99", 
                                "trunc_shift_distr" = "#FF89B4")) +
  xlab("Value") + ylab("Density") +
  ggtitle(paste("Distribution plot for non-transformed, truncated-shifted", 
                "and shifted-truncated Weibull distribution", sep="\n")) +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

print(pl)
```

* quantile function of a truncated and shifted Weibull distribution and of a shifted and truncated Weibull distribution:

```{r}
qstweibull <- qshift("tweibull")
qtsweibull <- qtruncate("sweibull")
x <- 1:999/1000
orig_quant <- qweibull(p = x, shape = 1.5, scale = 1)
shift_trunc_quant <- qstweibull(p = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
trunc_shift_quant <- qtsweibull(p = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
```

```{r}
df <- data.frame(x=x, 
                 type = factor(c(rep("orig_quant", times=length(x)),
                    rep("shift_trunc_quant", times=length(x)),
                    rep("trunc_shift_quant", times=length(x))),
                    levels = c("orig_quant", "shift_trunc_quant", "trunc_shift_quant")), 
                  value = c(orig_quant, shift_trunc_quant, trunc_shift_quant))

pl <- ggplot(data = df, mapping = aes(x=x, y=value, col=type)) +
  geom_line(size=1) +
  scale_color_manual(values = c("orig_quant" = "cyan3", 
                                "shift_trunc_quant" = "#89FF99", 
                                "trunc_shift_quant" = "#FF89B4")) +
  xlab("Value") + ylab("Density") +
  ggtitle(paste("Quantile plot for non-transformed, truncated-shifted", 
                "and shifted-truncated Weibull distribution", sep="\n")) +
  theme(legend.title=element_blank(), legend.text= element_text(size=10),
  plot.title = element_text(size=14), axis.title=element_text(size=12))

print(pl)
```

* random generation function of a truncated and shifted Weibull distribution and of a shifted and truncated Weibull distribution:

```{r}
rstweibull <- rshift("tweibull")
rtsweibull <- rtruncate("sweibull")
orig_rand <- rweibull(n=1000, shape = 1.5, scale = 1)
shift_trunc_rand <- rstweibull(n=1000, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
trunc_shift_rand <- rtsweibull(n=1000, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
```

```{r}
df <- data.frame(x=1:1000, 
                 type = factor(c(rep("orig_rand", times=1000),
                    rep("shift_trunc_rand", times=1000),
                    rep("trunc_shift_rand", times=1000)),
                    levels = c("orig_rand", "shift_trunc_rand", "trunc_shift_rand")), 
                  value = c(orig_rand, shift_trunc_rand, trunc_shift_rand))

pl <- ggplot(data = df, mapping = aes(x=value, group=type, fill=type)) +
        geom_density(alpha=0.6) +
        scale_fill_manual(values = c("orig_rand" = "cyan3",
                              "shift_trunc_rand" = "#89FF99",
                              "trunc_shift_rand" = "#FF89B4")) +
        ggtitle(paste("Density plot for non-transformed, truncated-shifted", 
                      "and shifted-truncated Weibull distribution", sep="\n")) +
        theme(legend.title=element_blank(), legend.text= element_text(size=10),
        plot.title = element_text(size=14), axis.title=element_text(size=12))

print(pl)
```

# Extending the computation

Once we demonstrated that we can generate distribution functions for truncated and shifted and for shifted and truncated distributions, we can use them for further implementations. Suppose we want to define a function for maximum likelihood estimate for truncated and shifted distributions and one for shifted and truncated distributions. In the following examples, we will continue to consider the Weibull distribution.

Let us start from the function for maximum likelihood estimate for truncated and shifted  Weibull distribution.
We consider the previously defined density function, `dstweibull()`, for a truncated and shifted Weibull distribution and we use it inside the estimator function:

```{r}
ltweibull <- function(x, shift = 0 ,L = -Inf, U = Inf){
  x1 <- x - shift
  # starting values for parameters (scale, shape)
  shape <- (sd(x1)/mean(x1))^(-1.086)
  scale <- mean(x1)/gamma(1+1/shape)
  # parameters vector definition
  theta <- c(shape, scale)
  # likelihood function
  ll <- function(theta, x, shift = 0, L = -Inf, U = Inf){
    shape <- theta[1]
    scale <- theta[2]
    ld <- dstweibull(x = x, shape = shape, scale = scale, shift = shift, L = L, U = U, log=TRUE)
    -sum(ld)
  }
  # maximum likelihood estimation
  optim(par = theta, fn = ll, x = x, shift=shift, L = L, U = U, method = "Nelder-Mead")[["par"]]
}
```

Let's test it to see how it works.  
We consider the previously defined random generator function `rstweibull()` for a truncated and shifted Weibull distribution to create the sample for testing the maximum likelihood estimate function:

```{r}
x <- rstweibull(n = 100000, shape = 1.6, scale = 1, shift = 1, L = 1, U = 2.5)
# maximum likelihood estimate for the shifted and truncated weibull distribution
ltweibull(x = x, shift=1, L = 1, U = 2.5)
```

The maximum likelihood estimate function produces a good estimate of the scale and shape parameters of truncated and shifted Weibull distribution.

Let us continue with the function for maximum likelihood estimate for shifted and truncated Weibull distribution.

We will consider the previously defined density function, `dtsweibull()`, for a shifted and truncated Weibull distribution. We use `dtsweibull()` inside the estimator function:

```{r}
ltweibull <- function(x, shift = 0, L = -Inf, U = Inf){
  x1 <- x - shift
  # starting values for parameters (scale, shape)
  shape <- (sd(x1)/mean(x1))^(-1.086)
  scale <- mean(x1)/gamma(1+1/shape)
  # parameters vector definition
  theta <- c(shape, scale)
  # likelihood function
  ll <- function(theta, x, L = -Inf, U = Inf){
    shape <- theta[1]
    scale <- theta[2]
    ld <- dtsweibull(x = x, shape = shape, scale = scale, shift = shift, L = L, U = U, log=T)
    -sum(ld)
  }
  # maximum likelihood estimation
  optim(par = theta, fn = ll, x = x, L = L, U = U, method = "Nelder-Mead")[["par"]]
}
```

Let's test it to see how it works.  
We consider the previously defined random generator function `rtsweibull()` for a shifted and truncated Weibull distribution to create the sample for testing the maximum likelihood estimate function:

```{r}
x <- rtsweibull(n = 100000, shape = 1.6, scale = 1, shift = 1, L = 0.5, U = 3)
# maximum likelihood estimate for the shifted weibull distribution
ltweibull(x = x, shift = 1, L = 0.5, U = 3)
```

Also in this case, the maximum likelihood estimate function produces a good estimate of the scale and shape parameters of shifted and truncated Weibull distribution.

# Bibliography
Giro V. . R package qshift: A functional programming approach to shifted probability ditributions. url  
Spanò A. . R package qdist: A functional programming approach to truncated probability ditributions. url  
Spanò A. . Ramarro. R for Developers. [http://www.quantide.com/ramarro-r-for-developers/](http://www.quantide.com/ramarro-r-for-developers/). Accessed: 2016-09-20   
Wickham H. . Advanced r. [http://adv-r.had.co.nz/](http://adv-r.had.co.nz/). Accessed: 2016-09-20.

### 2016-09-21 docs/shift+trunc-article/article-shift+trunc.Rmd
s

@@
 ---
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
 
@@
 So we can have a double layer of development: a first layer that is used to do all the complex work in common to all functions and a second layer that defines the details of each function.
-This fundamentals was used to develop the computational method proposed by Andrea Spanò in "R package qdist: A functional programming approach to truncated probability distributions", and that I used in "R package qshift: A functional programming approach to shifted probability distributions", that produces functions for truncated and shifted distributions, respectively.
+This concept was used to develop the computational method proposed by Andrea Spanò in "R package qdist: A functional programming approach to truncated probability distributions", and that I used in "R package qshift: A functional programming approach to shifted probability distributions", that produces functions for truncated and shifted distributions, respectively.
 [`qdist`](https://github.com/veronicagiro/qdist) and [`qshift`](https://github.com/veronicagiro/qshift) packages are available on Github:
 
-```{r, eval=FALSE}
+```{r repos, eval=FALSE}
 devtools::install_github("veronicagiro/qtruncate")
@@
 
-In the previous citated articles, we dealt separately with shifted and truncated distributions.
+In the previously cited articles, we dealt separately with shifted and truncated distributions.
 But in real life, we sometimes work with random variables arising from more than one transformation.
-So in this article, we will demonstrate that we can produce functions for truncated and shifted or for shifted and truncated distributions only combining the functions of the two packages.
+So in this article, we will demonstrate that we can produce functions for truncated and shifted or for shifted and truncated distributions only combining the functions of the `qdist` and `qshift` packages.
 This is possible because of functional programming approach used to build them.
@@
 * `q<dist>(p)` with `p` vector of probabilities, identifies quantile functions
-* `r<dist>(n)` with `n` number of observations, identifies random number generation
+* `r<dist>(n)` with `n` number of observations, identifies random number generator functions
 
@@
 
-Andrea Spanò proposed in `qdist` a functional programming method that computes truncated distributions functions for each distribution chosen without changing the well structured programming style and syntax of previous functions.
+Andrea Spanò proposed in `qdist` a functional programming method that computes truncated distribution functions for each distribution chosen without changing the well structured programming style and syntax of the previous functions.
 
@@
 
-* the environment of a function can be used as a placeholder for other objects (for detail see Andrea Spanò article "R package qdist: A functional programming approach to truncated probability ditributions"")
+* the environment of a function can be used as a placeholder for other objects (for details see Andrea Spanò article "R package qdist: A functional programming approach to truncated probability distributions")
 * function formals can be manipulated
@@
 1.
-defines the required functions suitable for computing from specific truncated probability distribution by using the functions available within its package.
+defines the required functions suitable for computing from a specific truncated probability distribution by using the functions available within its package.
 These functions take as input a probability distribution as a character string, in this case the Weibull distribution, and return the equivalent truncated distribution function as object:
 
-```{r, message=FALSE}
+```{r qdist_funs, message=FALSE}
 require(qdist)
@@
 
-```{r}
+```{r compare_args}
 args(dweibull)
@@
 
-As `dweibull()` is the function that computes the density function for a Weibull distribution, conventionally named `weibull`, `dtweibull()` is the function that computes the density function for a truncated Weibull distributions, named `tweibull`.
+As `dweibull()` is the function that computes the density function for a Weibull distribution, conventionally named `weibull`, `dtweibull()` is the function that computes the density function for a truncated Weibull distribution, named `tweibull`.
 
@@
 `qshift` package shares the same logic and the same methodology of `qdist`.
-The difference is that the functions are designed to compute functions for shifted distributions.
+The difference is that its functions are designed to generate functions for shifted distributions.
 In particular, `qshift` package includes the following four functions:
@@
 
-```{r, message=FALSE}
+```{r trunc_shift, message=FALSE}
 require(qshift)
@@
 
-```{r}
+```{r trunc_shift_args}
 args(dstweibull)
@@
 
-```{r}
+```{r trunc_shift_ex}
 x <- seq(0, 5, len = 100)
@@
 
-```{r, message=FALSE}
+```{r trunc_shift_ex_pl, message=FALSE}
 df <- data.frame(x=x, 
-                 type = factor(c(rep("orig_density", times=length(x)),
-                    rep("trunc_density", times=length(x)),
-                    rep("trunc_shift_density", times=length(x))),
-                    levels = c("orig_density", "trunc_density", "trunc_shift_density")), 
+                 type = factor(c(rep("original density", times=length(x)),
+                    rep("trunc density", times=length(x)),
+                    rep("trunc-shift density", times=length(x))),
+                    levels = c("original density", "trunc density", "trunc-shift density")), 
                   value = c(orig_density, trunc_density, trunc_shift_density))
+
 require(ggplot2)
@@
   geom_line(size=1) +
-  scale_color_manual(values = c("orig_density" = "cyan3", 
-                                "trunc_density" = "#89FF99", 
-                                "trunc_shift_density" = "#FF89B4")) +
+  scale_color_manual(values = c("original density" = "cyan3", 
+                                "trunc density" = "#c8a9e3", 
+                                "trunc-shift density" = "#47e583")) +
   xlab("Value") + ylab("Density") + ylim(c(0,1.9)) +
-  ggtitle(paste("Density plot for non-transformed, truncated", "and truncated and shifted Weibull distributions", sep = "\n")) +
+  ggtitle(paste("Density plot for original, truncated", 
+                "and truncated and shifted Weibull distributions", sep = "\n")) +
   theme(legend.title=element_blank(), legend.text= element_text(size=10),
@@
 Starting from `dshift()` function of `qshift` package, we define `dsweibull()`, the density function of `sweibull`, which is the shifted Weibull distribution:
-```{r}
+```{r shift_d}
 dsweibull <- dshift("weibull") # density function
@@
 
-We have to define also the others functions for shifted distribution, as they are implicitly necessary for the further step:
+We have to define also the others functions for shifted distribution, as they are implicitly necessary for further step:
 
-```{r}
+```{r shift_funs}
 psweibull <- pshift("weibull") # probability function 
@@
 The density function of a shifted and truncated Weibull distribution is generated setting `sweibull` distribution as `ptruncate()` argument:
-```{r}
+```{r shift_trunc_d}
 dtsweibull <- dtruncate("sweibull")
@@
 
-```{r}
+```{r shift_trunc_d_ex}
 orig_density <- dweibull(x = x, shape = 1.5, scale = 1)
@@
 
-```{r}
+```{r shift_trunc_ex_pl}
 df <- data.frame(x=x, 
-                 type = factor(c(rep("orig_density", times=length(x)),
-                    rep("shift_density", times=length(x)),
-                    rep("shift_trunc_density", times=length(x))),
-                    levels = c("orig_density", "shift_density", "shift_trunc_density")), 
+                 type = factor(c(rep("original density", times=length(x)),
+                    rep("shift density", times=length(x)),
+                    rep("shift-trunc density", times=length(x))),
+                    levels = c("original density", "shift density", "shift-trunc density")), 
                   value = c(orig_density, shift_density, shift_trunc_density))
@@
   geom_line(size=1) +
-  scale_color_manual(values = c("orig_density" = "cyan3", 
-                                "shift_density" = "#89FF99", 
-                                "shift_trunc_density" = "#FF89B4")) +
+  scale_color_manual(values = c("original density" = "cyan3", 
+                                "shift density" = "#39375b", 
+                                "shift-trunc density" = "#FF89B4")) +
   xlab("Value") + ylab("Density") + ylim(c(0,1.9)) +
-  ggtitle(paste("Density plot for non-transformed, shifted and","shifted and truncated Normal distributions", sep = "\n")) +
+  ggtitle(paste("Density plot for original, shifted and",
+                "shifted and truncated Normal distributions", sep = "\n")) +
   theme(legend.title=element_blank(), legend.text= element_text(size=10),
@@
 
-Let us see graphically the differences between the density function of Weibull distribution, of a truncated and shifted Weibull distribution and of a shifted and truncated Weibull distribution.
+Let us see graphically the differences between the density function of original Weibull distribution, of truncated and shifted Weibull distribution and of shifted and truncated Weibull distribution.
+We refers to values computed in the previous examples.
 
-```{r}
+```{r ts_st_ex_pl}
 df <- data.frame(x=x, 
-                 type = factor(c(rep("orig_density", times=length(x)),
-                    rep("trunc_shift_density", times=length(x)),
-                    rep("shift_trunc_density", times=length(x))),
-                    levels = c("orig_density", "trunc_shift_density", "shift_trunc_density")), 
+                 type = factor(c(rep("original density", times=length(x)),
+                    rep("trunc-shift density", times=length(x)),
+                    rep("shift-trunc density", times=length(x))),
+                    levels = c("original density", "trunc-shift density", "shift-trunc density")), 
                   value = c(orig_density, trunc_shift_density, shift_trunc_density))
@@
   geom_line(size=1) +
-  scale_color_manual(values = c("orig_density" = "cyan3", 
-                                "trunc_shift_density" = "#89FF99", 
-                                "shift_trunc_density" = "#FF89B4")) +
+  scale_color_manual(values = c("original density" = "cyan3", 
+                                "trunc-shift density" = "#47e583", 
+                                "shift-trunc density" = "#FF89B4")) +
   xlab("Value") + ylab("Density") + ylim(c(0,1.9)) +
-  ggtitle(paste("Density plot for non-transformed, truncated-shifted", 
+  ggtitle(paste("Density plot for original, truncated-shifted", 
                 "and shifted-truncated Weibull distribution", sep="\n")) +
@@
 
-Looking at the figure we can see that a truncated and shifted distribution is different from a shifted amd truncated distribution.
+Looking at the figure we can see that a truncated and shifted distribution is different from a shifted and truncated one.
 
-Let us analyze the other distributions functions:
+Let us analyze the other distribution functions:
 
@@
 
-```{r}
+```{r shift_trunc_p}
+# probability distribution function of a truncated and shifted Weibull distribution
 pstweibull <- pshift("tweibull")
+# probability distribution function of a shifted and truncated Weibull distribution
 ptsweibull <- ptruncate("sweibull")
-orig_distr <- pweibull(q = x, shape = 1.5, scale = 1)
-shift_trunc_distr <- pstweibull(q = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
-trunc_shift_distr <- ptsweibull(q = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
 ```
 
-```{r}
+```{r shift_trunc_p_ex}
+orig_distr <- pweibull(q = x, shape = 1.5, scale = 1)
+trunc_shift_distr <- pstweibull(q = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
+shift_trunc_distr <- ptsweibull(q = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
+```
+
+```{r shift_trunc_p_ex_pl}
 df <- data.frame(x=x, 
-                 type = factor(c(rep("orig_distr", times=length(x)),
-                    rep("shift_trunc_distr", times=length(x)),
-                    rep("trunc_shift_distr", times=length(x))),
-                    levels = c("orig_distr", "shift_trunc_distr", "trunc_shift_distr")), 
+                 type = factor(c(rep("original distr", times=length(x)),
+                    rep("shift-trunc distr", times=length(x)),
+                    rep("trunc-shift distr", times=length(x))),
+                    levels = c("original distr", "shift-trunc distr", "trunc-shift distr")), 
                   value = c(orig_distr, shift_trunc_distr, trunc_shift_distr))
@@
   geom_line(size=1) +
-  scale_color_manual(values = c("orig_distr" = "cyan3", 
-                                "shift_trunc_distr" = "#89FF99", 
-                                "trunc_shift_distr" = "#FF89B4")) +
-  xlab("Value") + ylab("Density") +
-  ggtitle(paste("Distribution plot for non-transformed, truncated-shifted", 
+  scale_color_manual(values = c("original distr" = "cyan3", 
+                                "shift-trunc distr" = "#FF89B4", 
+                                "trunc-shift distr" = "#47e583")) +
+  xlab("Quantile") + ylab("Probability") +
+  ggtitle(paste("Distribution plot for original, truncated-shifted", 
                 "and shifted-truncated Weibull distribution", sep="\n")) +
@@
 
-```{r}
+```{r shift_trunc_q}
+# quantile function of a truncated and shifted Weibull distribution
 qstweibull <- qshift("tweibull")
+# quantile function of a shifted and truncated Weibull distribution
 qtsweibull <- qtruncate("sweibull")
+```
+
+```{r shift_trunc_q_ex}
 x <- 1:999/1000
 orig_quant <- qweibull(p = x, shape = 1.5, scale = 1)
-shift_trunc_quant <- qstweibull(p = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
-trunc_shift_quant <- qtsweibull(p = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
+trunc_shift_quant <- qstweibull(p = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
+shift_trunc_quant <- qtsweibull(p = x, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
 ```
 
-```{r}
+```{r shift_trunc_q_ex_pl}
 df <- data.frame(x=x, 
-                 type = factor(c(rep("orig_quant", times=length(x)),
-                    rep("shift_trunc_quant", times=length(x)),
-                    rep("trunc_shift_quant", times=length(x))),
-                    levels = c("orig_quant", "shift_trunc_quant", "trunc_shift_quant")), 
+                 type = factor(c(rep("original quant", times=length(x)),
+                    rep("shift-trunc quant", times=length(x)),
+                    rep("trunc-shift quant", times=length(x))),
+                    levels = c("original quant", "shift-trunc quant", "trunc-shift quant")), 
                   value = c(orig_quant, shift_trunc_quant, trunc_shift_quant))
@@
   geom_line(size=1) +
-  scale_color_manual(values = c("orig_quant" = "cyan3", 
-                                "shift_trunc_quant" = "#89FF99", 
-                                "trunc_shift_quant" = "#FF89B4")) +
-  xlab("Value") + ylab("Density") +
-  ggtitle(paste("Quantile plot for non-transformed, truncated-shifted", 
+  scale_color_manual(values = c("original quant" = "cyan3", 
+                                "shift-trunc quant" = "#FF89B4", 
+                                "trunc-shift quant" = "#47e583")) +
+  xlab("Probability") + ylab("Quantile") +
+  ggtitle(paste("Quantile plot for original, truncated-shifted", 
                 "and shifted-truncated Weibull distribution", sep="\n")) +
@@
 
-```{r}
+```{r shift_trunc_r}
+# random generation function of a truncated and shifted Weibull distribution
 rstweibull <- rshift("tweibull")
+# random generation function of a shifted and truncated Weibull distribution
 rtsweibull <- rtruncate("sweibull")
-orig_rand <- rweibull(n=1000, shape = 1.5, scale = 1)
-shift_trunc_rand <- rstweibull(n=1000, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
-trunc_shift_rand <- rtsweibull(n=1000, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
 ```
 
-```{r}
+```{r shift_trunc_r_ex}
+orig_rand <- rweibull(n=1000, shape = 1.5, scale = 1)
+trunc_shift_rand <- rstweibull(n=1000, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
+shift_trunc_rand <- rtsweibull(n=1000, shape = 1.5, scale = 1, L = 1.5, U = 3, shift = 1)
+```
+
+```{r shift_trunc_r_ex_pl}
 df <- data.frame(x=1:1000, 
-                 type = factor(c(rep("orig_rand", times=1000),
-                    rep("shift_trunc_rand", times=1000),
-                    rep("trunc_shift_rand", times=1000)),
-                    levels = c("orig_rand", "shift_trunc_rand", "trunc_shift_rand")), 
+                 type = factor(c(rep("original rand", times=1000),
+                    rep("shift-trunc rand", times=1000),
+                    rep("trunc-shift rand", times=1000)),
+                    levels = c("original rand", "shift-trunc rand", "trunc-shift rand")), 
                   value = c(orig_rand, shift_trunc_rand, trunc_shift_rand))
@@
   geom_density(alpha=0.6) +
-        scale_fill_manual(values = c("orig_rand" = "cyan3",
-                              "shift_trunc_rand" = "#89FF99",
-                              "trunc_shift_rand" = "#FF89B4")) +
-        ggtitle(paste("Density plot for non-transformed, truncated-shifted", 
+  scale_fill_manual(values = c("original rand" = "cyan3",
+                              "shift-trunc rand" = "#FF89B4",
+                              "trunc-shift rand" = "#47e583")) +
+  ggtitle(paste("Density plot for original, truncated-shifted", 
                 "and shifted-truncated Weibull distribution", sep="\n")) +
+  xlab("Quantile") + ylab("Density") +
   theme(legend.title=element_blank(), legend.text= element_text(size=10),
@@
 
-```{r}
+```{r l_trunc_shift}
 ltweibull <- function(x, shift = 0 ,L = -Inf, U = Inf){
@@
 
-```{r}
+```{r l_trunc_shift_ex}
 x <- rstweibull(n = 100000, shape = 1.6, scale = 1, shift = 1, L = 1, U = 2.5)
@@
 
-We will consider the previously defined density function, `dtsweibull()`, for a shifted and truncated Weibull distribution. We use `dtsweibull()` inside the estimator function:
+We will consider the previously defined density function, `dtsweibull()`, for a shifted and truncated Weibull distribution and we use it inside the estimator function:
 
-```{r}
+```{r l_shift_trunc}
 ltweibull <- function(x, shift = 0, L = -Inf, U = Inf){
@@
 
-```{r}
+```{r l_shift_trunc_ex}
 x <- rtsweibull(n = 100000, shape = 1.6, scale = 1, shift = 1, L = 0.5, U = 3)

### 2016-09-21 docs/shift+trunc-article/article-shift+trunc.Rmd
article shift+trunc in pdf + metadescription

@@
 ---
-title: "R functional programming approach for truncated and shifted and for shifted and truncated probability distribution functions"
-output: html_document
+title: R functional programming approach for truncated and shifted and for shifted
+  and truncated probability distribution functions
+output:
+  pdf_document: default
+  html_document:
+    highlight: null
+    theme: null
 ---

### 2016-09-21 docs/shift+trunc-article/article-shift+trunc.Rmd
fevisione nomenclatura articolo

--- 2016-09-21 docs/shift+trunc-article/article-shift+trunc.Rmd
@@
 ---
-title: "R functional programming approach for truncated and shifted and for shifted and truncated probability distribution functions"
-output: html_document
+title: R functional programming approach for truncated and shifted and for shifted
+  and truncated probability distribution functions
+output:
+  html_document:
+    highlight: null
+    theme: null
 ---
@@
 ```{r l_trunc_shift}
-ltweibull <- function(x, shift = 0 ,L = -Inf, U = Inf){
+lstweibull <- function(x, shift = 0 ,L = -Inf, U = Inf){
   x1 <- x - shift
@@
 # maximum likelihood estimate for the shifted and truncated weibull distribution
-ltweibull(x = x, shift=1, L = 1, U = 2.5)
+lstweibull(x = x, shift=1, L = 1, U = 2.5)
 ```
@@
 ```{r l_shift_trunc}
-ltweibull <- function(x, shift = 0, L = -Inf, U = Inf){
+ltsweibull <- function(x, shift = 0, L = -Inf, U = Inf){
   x1 <- x - shift
@@
 # maximum likelihood estimate for the shifted weibull distribution
-ltweibull(x = x, shift = 1, L = 0.5, U = 3)
+ltsweibull(x = x, shift = 1, L = 0.5, U = 3)
 ```

### 2016-09-21 docs/shift+trunc-article/prove-qdist+qshift.R

@@ -1148,10 +1148,10 @@ ltweibull <- function(x, L = -Inf, U = Inf){
 
 rsweibull <- rshift("weibull")
 rtsweibull <- rtruncate("sweibull")
-x <- rtsweibull(n = 100000, shape = 1, scale = 1.5, shift = 1, L = 5, U = 10)
+x <- rtsweibull(n = 10000000, shape = 1, scale = 1.5, shift = 1, L = 1, U = 3)
 
 # maximum likelihood estimate for the shifted weibull distribution
-ltweibull(x = x, L = 5, U = 10)
+ltweibull(x = x, L = 1, U = 3)
 
 ## funzionaaaaaaaaaaaa
 
@@ -1165,33 +1165,33 @@ dstweibull <- dshift("tweibull")
 
 ltweibull <- function(x, shift = 0 ,L = -Inf, U = Inf){
   # starting values for parameters (scale, shape) and shift
-  #shift <- min(x) - 0.01
+  shift <- min(x) - 0.01
   x1 <- x - shift
   #x1 <- x
   shape <- (sd(x1)/mean(x1))^(-1.086)
   scale <- mean(x1)/gamma(1+1/shape)
   # parameters vector definition
-  #theta <- c(shape, scale, shift)
-  theta <- c(shape, scale)
+  theta <- c(shape, scale, shift)
+  #theta <- c(shape, scale)
   # likelihood function
-  ll <- function(theta, x, shift = 0, L = -Inf, U = Inf){
+  ll <- function(theta, x, L = -Inf, U = Inf){
     shape <- theta[1]
     scale <- theta[2]
-    #shift <- theta[3]
+    shift <- theta[3]
     ld <- dstweibull(x = x, shape = shape, scale = scale, shift = shift, L = L, U = U, log=TRUE)
     -sum(ld)
   }
   # maximum likelihood estimation
-  optim(par = theta, fn = ll, x = x, shift=shift, L = L, U = U, method = "Nelder-Mead")[["par"]]
+  optim(par = theta, fn = ll, x = x, L = L, U = U, method = "Nelder-Mead")[["par"]]
 }
 
 
 rtweibull <- rtruncate("weibull")
 rstweibull <- rshift("tweibull")
-x <- rstweibull(n = 100000, shape = 3, scale = 1.5, shift = 1, L = 1, U = 5)
[…]

### 2016-09-21 docs/qshift-article/article.Rmd
revisione nomenclatura articolo shift

@@
 ```{r max_likelihood_est}
-ltweibull <- function(x){
+lsweibull <- function(x){
   # starting values for parameters (scale, shape) and shift
@@
 # maximum likelihood estimate for the shifted weibull distribution
-ltweibull(x = x)
+lsweibull(x = x)
 ```
